### 2016-11-07 comp_sota.tex
first commit

\subsection{Comparison with the state of the art}
\label{sec:sota}
\begin{table}[!t]
  \centering
  \caption{Comparison to state-of-the-art.}
  \label{tbl:sota}
  %\begin{tabular}{|c|c|c|c|c|}
  \setlength{\tabcolsep}{.5em}
  \begin{tabular}{ccccc}
    \toprule
    Dataset & dim &\textbf{Holidays} & \textbf{Oxford 5k} & \textbf{Oxford 105k}\\
    \midrule
    SPoC+GaussianSLEM                           & 1e3  &  81.4           &  \textbf{64.9}  &   -  \\
    SPoC+PCA+PolySLEM                           & 1e3  &  82    &   64.5  &  \textbf{62.3} \\
    Zepeda \textit{et al.} \cite{ZePe15}        & 8e3 &   78.3   &  57.5&  - \\ 
    Babenko \textit{et al.} \cite{babenko15}     & 2e2  &  80.2           &   58.9  &  57.8 \\
    J\'egou \textit{et al.} \cite{sota1}         & -    &  81.3           &    61.5    &   -   \\
    Jain \textit{et al. } \cite{JaJeGro11}       &   -  &  81.9           &    -    &   -   \\
    \midrule
    Tolias \textit{et al.} \cite{Tolias13}      & 1e7  &  \textbf{88}     &  \textbf{82}    &  \textbf{75}   \\
    \bottomrule
  \end{tabular}
  \label{tab:sota}
\end{table}

In Table \ref{tbl:sota} we compare our proposed SLEM architectures to various state-of-the-art methods using a global image representation, as listed in \cite{holidaysSOTA}. 
Note that our method performs comparably to all methods except the very-high dimensional method of Toilas \textit{et al.} \cite{Tolias13} which uses features of size $1e7$.

### 2016-11-07 conclusion.tex
first commit

\section{Conclusion and future work}
\label{conclusion}
In this paper, we address the problem of large scale image retrieval with global image representation. 
We contribute to the domain by presenting a simple idea, the kernelized square loss exemplar machine, and its efficient implementation.
As a result, we obtain significant improvements over the image representations we tested and outperform similar encoders on different datasets and on large scale image search. 
As future work, we must test our model on state of the art representations and see how it impacts their results.
The use of other kernel functions is also worth investigating. The polynomial kernel outperforms the Gaussian kernel, even though the Hilbert space obtained from the Gaussian kernel has infinite dimensions and the Hilbert space obtained from the polynomial kernel does not.
The spatial pyramid kernel can be used to improve representations based on local descriptors and it offers the better boosts when compared with its baseline.
Many other performing matching kernel are yet to be tested \cite{Tolias13}, and this work introduces a model that we hope can be reproduced for all kernel functions.

### 2016-11-07 esvm.tex
first commit

\documentclass[12pt,letterpaper]{article}

%\usepackage{cvpr}
%\documentclass[11pt,letterpaper]{article}
\setlength{\textheight}{8.875in}
\setlength{\textwidth}{6.875in}
\setlength{\columnsep}{0.3125in}
\setlength{\topmargin}{0in}
\setlength{\headheight}{0in}
\setlength{\headsep}{0in}
\setlength{\parindent}{1pc}
\setlength{\oddsidemargin}{-.304in}
\setlength{\evensidemargin}{-.304in}

\usepackage{times}
\usepackage{bbm}
\usepackage{epsfig}
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{amsthm}
\usepackage{mathrsfs}
\usepackage{caption}
\usepackage{subcaption}
%\usepackage{caption}
\DeclareMathOperator*{\argmin}{\arg\!\min}
\newtheorem{lemma}{Lemma}
\newtheorem{proposition}{Proposition}
\newtheorem{theorem}{Theorem}
\newtheorem{coro}{Corollary}
\newtheorem{prop}{Property}
\newtheorem{definition}{Definition}

% If you comment hyperref and then uncomment it, you should delete
% egpaper.aux before re-running latex. (Or just hit 'q' on the first latex
% run, let it finish, and you should be clear).

\def\texte#1{#1}

\begin{document}
\def\RR{\mathbb{R}}
\def\PP{\mathbb{P}}
\def\AA{\mathbb{A}}
\def\LL{\mathbb{L}}
\def\SS{\mathbb{S}}
\def\barr{\bar{\mathbb{R}}}
\def\mat#1{{\mathcal{#1}}}
\def\vect#1{\mbox{\boldmath $#1$}}
\def\PPi{\mbox{\boldmath$\Pi$}}
\def\squig{\rightsquigarrow}
%\def\vect#1{#1}
%\def\mat#1{#1}
\def\comment#1{{}}
%\def\qmatrix#1{\left[\begin{array}{l}#1\end{array}\right]}
\def\qmatrix#1{\left[\begin{matrix}#1\end{matrix}\right]}

\title{Kernel Square-Loss Examplar Machines}
\author{Jean Ponce}
%\date{}
\maketitle

\begin{abstract}
Zepeda and P\'erez show in~\cite{ZePe15} that the weight vectors
associated with examplar SVMs can be used as global features
representing images in tasks such as retrieval or classification. A
drawback is the cost of the procedure, which requires a (convex)
optimization process to be run for each positive image. This note
shows that a more efficient approach can be obtained using a square
loss instead of the hinge loss. Building on~\cite{BaJo05,FiSc01}, it
also shows that this approach is easily kernelized, and, assuming
low-rank kernels, that the corresponding computational cost is linear
in the number of negative examples for each positive example. The
presentation concludes with a brief discussion of the relationship
between the proposed approach with linear discriminant analysis and
its use in region matching in~\cite{ARS14}.
\end{abstract}

\section{Square-loss examplar machines\label{sec:sqesvm}}
Given some loss function $L:[-1,+1]\times \RR\rightarrow \RR^+$, some
positive training example $x_0$ in $\RR^p$ and $n$ negative training
examples $x_i$ in $\RR^p$ ($i=1,\ldots,n$), training an exemplar
linear classifier consists in minimizing the error
\begin{equation}
E(w,b)=\frac{1}{n}\sum_{i=1}^n L(-1,w\cdot x_i+b)+ L(1,w\cdot
x_0+b)+\frac{\lambda}{2} ||w||^2
\label{eq:general}
\end{equation}
with respect to $w$ in $\RR^p$ and $b$ in $\RR$.

The traditional examplar SVM uses the hinge loss for $L$. Let us use
instead the square loss. Equation~(\ref{eq:general}) becomes
\begin{equation}
E(w,b)=\frac{1}{2n}\sum_{i=1}^n (w\cdot x_i+b+1)^2+ \frac{1}{2}(w\cdot
x_0+b-1)^2+\frac{\lambda}{2} ||w||^2.
\label{eq:square}
\end{equation}

The interest of the square loss is that the minimization can be done
in closed form. Indeed, writing that $\partial E/\partial b=0$ at a
minimum yields
\begin{equation}
b=-\frac{1}{2}(x_0+\mu)\cdot w,
\end{equation}
where $\mu=\frac{1}{n}\sum_{i=1}^n x_i$ is the center of mass of the
negative samples $x_i$.

Substituting in Eq.~(\ref{eq:square}) now reduces our optimization problem
to the minimization with respect to $w$ of
\begin{equation}
\begin{array}{lcl}\hat{E}(w)&=&
\displaystyle\frac{1}{2n}\sum_{i=1}^n [w\cdot (x_i-\frac{1}{2}(x_0+\mu))+1]^2
+\frac{1}{2}[w\cdot (x_0-\frac{1}{2}(x_0+\mu))-1]^2+\frac{\lambda}{2}||w||^2\\
&=&
\displaystyle\min_{w\in\RR^p}
\frac{1}{2n}\sum_{i=1}^n [w\cdot (\sigma_i-\frac{1}{2}\delta) +1]^2
+\frac{1}{2}(\frac{1}{2}w\cdot\delta-1)^2+\frac{\lambda}{2}||w||^2,
\end{array}
\label{eq:weq}
\end{equation}
where $\delta=x_0-\mu$ and $\sigma_i=x_i-\mu$.

Finally, writing $\partial\hat{E}/\partial w=0$ at a minimum yields
\begin{equation}
\frac{1}{n}\sum_{i=1}^n [w\cdot(\sigma_i-\frac{1}{2}\delta)+1](\sigma_i-\frac{1}{2}\delta)
+\frac{1}{2}(\frac{1}{2}w\cdot\delta-1)\delta +\lambda w=0,
\end{equation}
or
\begin{equation}
[\frac{1}{n}\sum_{i=1}^n(\sigma_i-\frac{1}{2}\delta)(\sigma_i^T-\frac{1}{2}\delta^T)] w
+\frac{1}{n}\sum_{i=1}^n (\sigma_i-\frac{1}{2}\delta) +\frac{1}{2}\delta\frac{1}{2}\delta^T w -\frac{1}{2}\delta +\lambda w
=0.
\end{equation}

Using the fact that the variables $\sigma_i$ have zero
mean, we obtain
\begin{equation}
[\frac{1}{n}\sum_{i=1}^n\sigma_i\sigma_i^T +\frac{1}{2}\delta\delta^T +\lambda\text{Id}]
w=\delta,
\end{equation}
or, putting everything together and reverting to the original notation,
\begin{equation}
\left\{\begin{array}{l}
\displaystyle w=A^{-1} (x_0-\mu),\quad\text{where}\quad
A=\Sigma
+\frac{1}{2}(x_0-\mu)(x_0-\mu)^T +\lambda\text{Id},\\
\displaystyle b=-\frac{1}{2}(x_0+\mu)\cdot w,
\end{array}\right.
\label{eq:final}
\end{equation}
where
\begin{equation}
\Sigma=\frac{1}{n}\sum_{i=1}^n(x_i-\mu)(x_i-\mu)^T
=\frac{1}{n}\sum_{i=1}^n x_ix_i^T-\mu\mu^T=\frac{1}{n}XX^T-\mu\mu^T,
\label{eq:Sigmadef}
\end{equation}
is the covariance matrix of the negative samples $x_i$, and
$X$ denotes the $p\times n$ matrix with columns $x_i^T$ ($i=1,\ldots,n$).

The vector $w$ (or the vector $[w;b]$) of this {\em square-loss
examplar machine} can potentially be used as a new representation for
the corresponding positive feature $x_0$. Note that this gives
asymmetric roles to positive and negative examples.

A similar approach (with no regularizer) is advocated for region
matching in~\cite{ARS14}. We will come back to this issue in the last
section of this presentation.

\section{Kernelized version}
Let us recall a few basic facts avout kernel methods for supervised
classification. We consider a reprodcing kernel Hilbert space (RKHS)
$H$ formed by real functions over some set
$X$, and denote by $k$ the corresponding reproducing kernel. We
address the following learning problem over $H\times\RR$:
\begin{equation}
\min_{h\in H,b\in\RR}
\sum_{i=1}^n L[y_i,h(x_i)+b] + \frac{\lambda}{2}||h||^2,
\label{eq:kernel}
\end{equation}
where the pairs $(x_i,y_i)$ ($i=1,\ldots,n$) in $X\times Y$ are
training samples, $Y=\RR$ (regression) or $Y=\{-1,1\}$
(classification), and $L: Y\times \RR\rightarrow\RR$ is some arbitrary
loss function. By definition of a reproducing kernel,
Eq.~(\ref{eq:kernel}) can be rewritten as
\begin{equation}
\min_{h\in H,b\in\RR}
\sum_{i=1}^n L[y_i,( \varphi(x_i)|h)+b] +
\frac{\lambda}{2}||h||^2,
\label{eq:kernelaff}
\end{equation}
where $\varphi$ is the {\em feature map} over $X$ associated with the
kernel $k$ (which may not admit a known explicit form). We dub
problems with the general form of (\ref{eq:kernelaff}) {\em affine}
supervised learning problems since, given some fixed element $h$ of
$H$ and some scalar $b$, $(h|h')+b$ is an affine function of $h'$,
whose zero set defines an affine hyperplane of $H$ considered itself
as an affine space.

Let $K$ denote the kernel matrix with entries $k_{ij}=[\varphi(x_i)|
\varphi(x_j)]$ and rows $k_i^T$ ($i=1,\ldots,n$). We assume from now
on that $L$ is convex and continuous. We have the following theorem.

\begin{theorem}\label{th:affinelearn}
Equation~(\ref{eq:kernelaff}) admits equivalent primal and dual
versions.

The primal problem is
\begin{equation}
\min_{\alpha\in\RR^n,b\in\RR} \sum_{i=1}^n L[y_i,k_i^T\alpha+b]
+\frac{\lambda}{2}\alpha^TK\alpha,
\label{eq:primal1}
\end{equation}
and any solution $(\alpha^\star,b^\star)$ to (\ref{eq:primal1})
provides
a solution $(h^\star,b^\star)$ to (\ref{eq:kernelaff}) with
$h^\star=\sum_{i=1}^n \alpha_i^\star\varphi(x_i)$.

Let $r$ denote the rank of the kernel matrix $K$, a
{\em reduced} version of the primal problem is
\begin{equation}
\min_{\beta\in\RR^r,b\in\RR} \sum_{i=1}^n L[y_i,b_i^T\beta+b]
+\frac{\lambda}{2}||\beta||^2,
\label{eq:primal2}
\end{equation}
where $B$ is the $n\times r$ square root of $K$ and $b_i^T$ denotes
its $i$th row,
and any solution $(\beta^\star,b^\star)$ to (\ref{eq:primal2})
provides
a solution $(h^\star,b^\star)$ to (\ref{eq:kernelaff}) with
$h^\star=\sum_{i=1}^n \alpha_i^\star\varphi(x_i)$, and
$\alpha^\star=P\beta^\star$, where $P$ is the pseudo-inverse of $B^T$.

The dual problem is
\begin{equation}
\min_{\alpha\in\RR^n}
\sum_{i=1}^n L^*(y_i,-\lambda\alpha_i)
+\frac{\lambda}{2}\alpha^TK\alpha
\quad\text{subject to}\quad \sum_{i=1}^n \alpha_i=0,
\label{eq:finaldual}
\end{equation}
and the optimal value $b^*$ of $b$ given any solution
$\alpha^\star$ of (\ref{eq:finaldual}) can be found by solving
\begin{equation}
\min_{b\in \RR} \sum_{i=1}^n L(y_i,k_i^T\alpha^{\star}+b).
\label{eq:minb}
\end{equation}
The pair $(\alpha^\star,b^\star)$ then provides a solution
$(h^\star,b^\star)$ to (\ref{eq:kernelaff}) with $h^\star=
\sum_{i=1}^n\alpha_i^\star\varphi(x_i)$.
\end{theorem}

Note that, given a solution $(\alpha^\star,b^\star)$ of either the
primal or the dual version of (\ref{eq:kernelaff}), the optimal
prediction function is
\begin{equation}
[h^\star|\varphi(x)]+b^\star= \sum_{i=1}^n \alpha_i^{\star} K(x,x_i) +b^\star.
\end{equation}

Theorem\ref{th:affinelearn} follows fromt the representer theorem
~\cite{SHS01,Wahba90} and elementary results from analysis and
optimization theory.

In our context, the reduced primal version of Theorem 1 is
particularly interesting since the results of Section~\ref{sec:sqesvm}
directly apply to it: One simply has to replace $x_0$ and $\mu$ in
Eq.~(\ref{eq:final}) by appropriate columns of $B^T$, or linear
combinations thereof (in this case of course, the kernel matrix has
dimension $(n+1\times n+1)$). It is of course not clear whether the
corresponding vector $w$ would prove useful as a feature for
classification or other tasks.

The next section shows how to train the kernel square-loss examplar
machine efficiently.

\section{Efficient implementation}
Let us consider the $n\times n$ kernel matrix $\hat{K}$ associated with
the $n$ negative examples, and a low-rank factorization $\hat{K}=\hat{B}\hat{B}^T$
of
this matrix, where $\hat{B}$ is an
$n\times r$ square root of $\hat{K}$. This factorization can be computed in $O(nr^2)$ time
and $O(nr)$ storage~\cite{BaJo05,FiSc01}. The singular value
decomposition $UWV^T$ of $\hat{B}$ and its $r\times n$ pseudoinverse
$\hat{B}^\dagger=(\hat{B}^T\hat{B})^{-1}\hat{B}^T=VW^{-1}U^T$ such that $\hat{B}^\dagger
\hat{B}=\text{Id}$ can also be computed with the same time and storage
costs.

Let us consider the augmented
kernel matrix $K$ obtained by adding one more (positive) sample. This matrix
can be written as
\begin{equation}
K=\qmatrix{k_{00} & k_0^T\\k_0& \hat{K}},
\end{equation}
where $k_{00}$ is a scalar and $k_0$ is an element of $\RR^n$.
The following
lemma show that a $(n+1)\times (r+1)$ factorization of $K$ can
be computed efficiently as well.

\begin{lemma}
The augmented kernel matrix $\hat{K}$ can be factorized as
\begin{equation}
\hat{K}=BB^T,\quad\text{where}\quad
B=\qmatrix{u & v^T\\0 & \hat{B}}\quad\text{and}\quad
v=\hat{B}^\dagger k_0,\,\, u=\sqrt{k_{00}-||v||^2}
\end{equation}
in time $O(nr)$ and storage $O(n)$.
\end{lemma}
\begin{proof}
Indeed,
\begin{equation}
BB^T=\qmatrix{u^2+||v||^2 & v^T\hat{B}^T\\ \hat{B}v & \hat{B}\hat{B}^T}.
\end{equation}
Since $K$ is positive semidefinite $k_0$ must lie in the column space
$\cal \hat{B}$ of $\hat{B}$,\footnote{Let
us assume that $k_0$ does not belong to $\cal \hat{B}$, then it can be
decomposed uniquely as $k_0=l_0+m_0$, where $l_0$ belongs to $\cal \hat{B}$,
and $m_0\neq 0$ belong to the orthogonal complement $\cal \hat{B}^\perp$ of
$\cal \hat{B}$. Obviously $(1,-\nu m_0^T)K(1;-\nu m_0)=k_{00}-2\nu||m_0||^2$, which
is negative for large enough values of $\nu>0$. In turn, this contradicts the
fact that a kernel matrix is positive semidefinite.}
and $b=\hat{B}^\dagger k_0$ is thus an exact solution of
$\hat{B}b=k_0$.
The fact that $k_{00}-||b||^2$ is nonnegative follows immediately from
the fact that the Schur complement $\hat{K}-||k_0||^2/k_{00}$ of $k_{00}$ in
the positive semidefinite matrix $K$ is itself positive semidefinite.\footnote{Indeed,
since the matrix $k_{00}K-k_0k_0^T=\hat{B}(k_{00}\text{Id}-bb^T)\hat{B}^T$ is
positive semidefinite, and $\hat{B}$ has rank $r$, the $r\times r$ matrix
$k_{00}\text{Id}-bb^T$
is also positive definite, thus
$b^T(k_{00}\text{Id}-bb^T)b=||b||^2(k_{00}-||b||^2)\ge 0$.} The
cost of the factorization is dominated by the $O(nr)$ computation and $O(n)$ storage
of $b$
\end{proof}

Training the kernel square-loss examplar machine given $\hat{B}$ amounts to
computing
$\Sigma$ in (\ref{eq:Sigmadef}) and solving the $(r+1)\times (r+1)$
linear system in (\ref{eq:final}). Note that, because of the
particular structure of $B$, only one term in the sum in (\ref{eq:Sigmadef})
must be recomputed for each new positive example, with $O(r^2)$ cost.
The linear system itself can be solved in $O(r^3)$ cost.

Once the optimal parameters $\beta^\star$ haven been found, they can
be used directly as some representation of the positive example. The
corresponding value $\alpha^\star$ (or more correctly, {\em a}
corresponding value) can be computed as $\alpha^\star =P\beta^\star$,
where $P=B(B^TB)^{-1}$ is the pseudoinverse of $B^T$ with
$B^TP=\text{Id}$. The corresponding element of $\cal H$ is
$h^\star=\sum_{i=0}^n \alpha_i^\star\varphi(x_i)$.

Now, given two positive images $x_0$ and $x_0'$, they can be compared
by computing the normalized inner product
\begin{equation}
s(x_0,x_0')=\frac{(h|h')}{\sqrt{(h|h)}\sqrt{(h'|h')}}
\end{equation}
of the corresponding elements $h$ and $h'$ of $H$ as a similarity measure
(we drop the ``$\star$'' exponents from now on to avoid clutter in the
noration).

Concretely, we have
\begin{equation}
\begin{array}{lcl}
(h|h')&=&\displaystyle[\alpha_0\varphi(x_0)+\sum_{i=1}^n\alpha_i\varphi(x_i)
|\alpha_0'\varphi(x_0')+\sum_{i=1}^n\alpha_i'\varphi(x_i)]\\
&=&\displaystyle
\alpha_0\alpha_0'\displaystyle k(x_0,x_0')+ \alpha_0\sum_{i=1}^n\alpha_i' k(x_0,x_i)
+\alpha_0'\sum_{i=1}^n
\alpha_ik(x_0',x_i)+\hat{\alpha}^T\hat{K}\hat{\alpha}'\\
&=&\displaystyle\alpha_0\alpha_0'\displaystyle k(x_0,x_0')+ \alpha_0k_0^T\hat{\alpha}'
+\alpha_0'k_0'^T\hat{\alpha}
+\hat{\alpha}^T\hat{K}\hat{\alpha}'\\
\end{array}
\label{eq:lcl}
\end{equation}
where $\hat{\alpha}$ (resp. $\hat{\alpha}'$) denotes the vector of
$\RR^n$ formed by excluding the first (or rather zeroth) element
of $\alpha$ (resp. $\alpha'$).

If we write $\beta = [\beta_0; \hat{\beta}]$ similarly to
$\alpha=[\alpha_0; \hat{\alpha}]$ and keeping in mind that
$P=[\frac{1}{u}\ 0; -\frac{1}{u}\hat{P}v \ \hat{P}]$, where $\hat{P}$
is the pseudoinverse of
$\hat{B}^T$, we have that
\begin{equation}
\alpha_0 = \dfrac{\beta_0}{u}, \,\, \hat{\alpha}=-\dfrac{\beta_0}{u}Pv+P\hat{\beta}.
\end{equation}
Remembering that $\hat{K}=\hat{B}\hat{B}^T$ and and $k_0=\hat{B}v$
allows us to rewrite (\ref{eq:lcl}) as
\begin{align}
(h|h')& = \dfrac{\beta_0\beta_0'}{uu'}k(x_0,x_0')
+\dfrac{\beta_0}{u}v^T\hat{B}^T\left(-\dfrac{\beta_0'}{u'}Pv'+P\hat{\beta'}\right)
+\dfrac{\beta_0'}{u'}v'^T\hat{B}^T\left(-\dfrac{\beta_0}{u}Pv+P\hat{\beta}\right)\\ &+\left(\hat{B}^T\left(-\dfrac{\beta_0}{u}Pv+P\hat{\beta}\right)\right)^T
\left(\hat{B}^T\left(-\dfrac{\beta_0'}{u'}Pv'+P\hat{\beta'}\right)\right)\\ &=
\dfrac{\beta_0\beta_0'}{uu'}k(x_0,x_0')
+\dfrac{\beta_0}{u}v^T\left(-\dfrac{\beta_0'}{u'}v'+\hat{\beta'}\right)
+\dfrac{\beta_0'}{u'}v'^T\left(-\dfrac{\beta_0}{u}v+\hat{\beta}\right)
+
\left(-\dfrac{\beta_0}{u}v+\hat{\beta}\right)^T\left(-\dfrac{\beta_0'}{u'}v'+\hat{\beta'}\right)\\ &=
\hat{\beta}^T\hat{\beta'}
+\dfrac{\beta_0\beta_0'}{uu'}\left(k(x_0,x_0')-v^Tv'\right).
\end{align}
We also note that $(h|h)=||\beta||^2$ and $(h'|h')=||\beta'||^2$.

Furthermore, from the solution of the linear system (\ref{eq:final}), we find that $\beta_0 = \frac{cu}{\lambda}$, $\hat{\beta} = cG^{-1}(v-\mu)$ and $\beta_0' = \frac{cu'}{\lambda}$, $\hat{\beta}' = c'G^{-1}(v'-\mu)$, for $c,\ c'$ scalars and $G = \frac{1}{n}\hat{B}^T\hat{B}-\mu\mu^T+\lambda\text{Id}$. The final similarity score can be computed by
\begin{equation}
s(x_0,x_0') = \dfrac{(v-\mu)^TG^{-2}(v'-\mu)+\dfrac{1}{\lambda^2}(k(x_0,x_0')-v^Tv')}{\sqrt{(v-\mu)^TG^{-2}(v-\mu)+\dfrac{u^2}{\lambda^2}}\sqrt{(v'-\mu)^TG^{-2}(v'-\mu)+\dfrac{u'^2}{\lambda^2}}}
\end{equation}

\section{Notes: LDA and square-loss linear classifiers}
Let us now go back to linear classifiers, and suppose that we have
multiple positive samples. It is easy to show that in this case, the
corresponding linear classifier for the square loss is also given by
Eq.~(\ref{eq:final}), where $x_0$ denotes this time the center of mass
of the positive samples, {\em if} these samples have the {\em same}
covariance matrix $\Sigma$ as the negative ones.

This equal-covariance assumption is of course quite restrictive, and
probably unrealistic in general. It is interesting to note, however,
that this is exactly the assumption made by linear discriminant
analysis. As shown in~\cite{Hastie2009} for example, LDA can be seen as a
(non-regularized) linear classifier
with decision function $w'\cdot z+ b'$, where $z$ is a sample in
$\RR^p$, and
\begin{equation}
\left\{\begin{array}{l}
\displaystyle w'=\Sigma^{-1}(x_0-\mu),\\
\displaystyle b'=-\frac{1}{2}(x_0+\mu)\cdot w',
\end{array}\right.
\label{eq:lda}
\end{equation}

Note that when $\lambda=0$ (no regularization),
\begin{equation}
\Sigma w = Aw -\frac{1}{2}[(x_0-\mu)\cdot w] (x_0-\mu) =
[1-\frac{1}{2}(x_0-\mu)\cdot w](x_0-\mu)
=[1-\frac{1}{2}(x_0-\mu)\cdot w]\Sigma w'
\end{equation}
thus $w$ and $w'$ have the same direction. In other words, a
square-loss linear classifier is strictly equivalent to LDA when the
regularizer parameter $\lambda $ is zero.

More generally, as shown in~\cite{Hastie2009},
\begin{enumerate}
\item LDA is a nearest-centroid binary classifier using the
Mahalanobis distance under the assumption that the two classes
have the same covariance.
\item Its decision function is a Bayes classification rule for normal
distributions with different means but identical covariance matrices
when, in addition, the priors over the two classes are assumed to be
equal (the log-ratio of the prior probabilities must be added to the
offset otherwise). \footnote{In general, Hastie
{\em et al.}~recommend taking the priors $\pi_i$ equal to the proportion of points
in class $i$ among all points, that is the class with most samples is
most probable.}
\end{enumerate}

Once again, these facts are well known, but may be relevant (or not)
given recent renewed interest in LDA in modern classification
tasks~\cite{GMPD12,HMR12}, where some of these properties have been
rediscovered. In particular, as noted earlier, (non-regularized)
linear square-loss exemplar machines are proposed in~\cite{ARS14} as a
means to treat region matching as a classification problem. Aubry {\em
et al.} note in~\cite{ARS14} that the decision hyperplanes obtained by
this method and LDA are parallel (our analysis shows they are actually
identical --- there must be a mistake somewhere). They further show
that LDA scores calibrated using an additive constant give much better
results in their matching task. This is a bit puzzling since they
should give exactly the same results (same rankings and same scores up
to an affine transformation---before squaring the error term) for a
given positive example. Aubry {\em et al.} may mean calibration across
different positive examples. This needs to be clarified and perhaps
justified as well.

\section{SLEM versus whitening}
Due to Woodbury identity, Equation \ref{eq:final} can be rewritten as 
\begin{equation}
w = \Sigma^{-1}(x_0-\mu).
\end{equation}
In the original E-SVM work, the score of similarity between a query image $x_0$ and a possible result $x_0'$ is given by
\begin{equation}
s(x_0,x_0') = x_0'^T\Sigma^{-1}(x_0-\mu). \label{white:score}
\end{equation}

Under the assumption centered data and positive and negative data having the same covariance matrix (as proposed in the previous section), Equation \ref{white:score} corresponds to the score of whitened data:
\begin{align}
\hat{x_0} = \Sigma^{-\frac{1}{2}}(x_0-mu);\\
s(x_0,x_0') = \hat{x_0}^T\hat{x_0'}.
\end{align}
This observation has already been made by \cite{Koba15}. However, the score of similarity used in \cite{ZePe15} is
\begin{equation}
s(x_0,x_0') = (x_0'-\mu)^T\Sigma^{-2}(x_0-\mu). \label{esvm:score}
\end{equation}

Based on this, we propose the score $s_{\alpha}$ given by $s_{\alpha}(x_0,x_0') = (x_0-\mu)^T\Sigma^{\alpha}(x_0-\mu).$ $s_{-2}$ is the SLEM similarity score, $s_{-1}$ is the whitening similarity score and $s_{0}$ is the centered base features similarity score.

Strangely, both scores seem to perform similarly, even if their scores differ. Doing a singular value decomposition of $\Sigma$, $\Sigma^{-1}=VD^{-1}V^T$ and $\Sigma^{-\alpha}=VD^{-\alpha}V^T$.

\begin{figure}[!h]
\centering
\begin{subfigure}[b]{0.45\textwidth}
\includegraphics[width=\textwidth]{whitening_cnn.jpg}
\end{subfigure}
\begin{subfigure}[b]{0.45\textwidth}
\includegraphics[width=\textwidth]{whitening_vlad.jpg}
\end{subfigure}
\caption{Comparison of mAP for different similarity scores $s_{\alpha}$, varying $\alpha$. At left, CNN features. At right, VLAD features. On blue, $\Sigma$ is the covariance of the negative samples and on red, $\Sigma$ is the covariance of the positive and negative samples. On black, the mAP for similarity calculated as inner product of base features.}
\label{Sigma}
\end{figure}

\section{Experiment Evaluation}
\subsection{Base image encoding}
We use three base features as the representation in $\RR^p$ of our images. The two firsts are VLAD and CNN features, as used in \cite{ZePe15}. 
CNN features are non-negative and can be used both as $\mathbb{L}^1$ or $\mathbb{L}^2$ normalized features. The third features are pyramids of histograms of SIFT descriptors, as used in \cite{spk}. These are non-negative $\mathbb{L}^1$ normalized features.

\subsection{Kernels}
For all features, we test non-kernelized SLEM, as well as kernelized SLEM for gaussian (rbf) and polynomial kernel:
\begin{align}
k_{rbf}(x,y) &= \exp\left(-\gamma||x-y||^2\right);\\
k_{poly}(x,y) &= x^Ty+\gamma(x^Ty)^2.
\end{align}
For non-negative $\mathbb{L}^1$ normalized, we also test the intersection kernel: $k_{inter}(x,y) = \min(x,y).$\\
We observe that the non-kernelized obtains similar results to kernels polynomial and Gaussian. For the pyramid histogram of SIFT, the intersection kernel works much better than all the other kernels.

\begin{table}
\begin{center}
\begin{tabular}{|l|c|c|c|c|}
\hline
Method & $s_0$ & $s_{-2}$ with no kernel & $s_{-2}$ for gaussian kernel &  $s_{-2}$ for intersection kernel \\
\hline\hline
VLAD-64 & 73.3 & 77.9 & 78.1 & -\\
Caffe CNN & 69.7 & 72.1 & 72.9 & 70.2\\
Pyramid of Histogram & 38.7 & 43 & 44.6 & 63.5\\
\hline
\end{tabular}
\end{center}
\caption{Results.}
\end{table}

\bibliographystyle{plain}
\bibliography{sup}
\end{document}

### 2016-11-07 esvm_compversion.tex
first commit

--- 2016-11-07 esvm.tex
@@
 \usepackage{subcaption}
-%\usepackage{caption}
 \DeclareMathOperator*{\argmin}{\arg\!\min}
@@
 
-\section{Square-loss examplar machines\label{sec:sqesvm}}
-Given some loss function $L:[-1,+1]\times \RR\rightarrow \RR^+$, some
+\section{Square-loss exemplar machines\label{sec:sqesvm}}
+Given some loss function $l:[-1,+1]\times \RR\rightarrow \RR^+$, some
 positive training example $x_0$ in $\RR^p$ and $n$ negative training
@@
 \begin{equation} 
-E(w,b)=\frac{1}{n}\sum_{i=1}^n L(-1,w\cdot x_i+b)+ L(1,w\cdot
-x_0+b)+\frac{\lambda}{2} ||w||^2
+E(\omega,b)=\frac{1}{n}\sum_{i=1}^n l(-1,\omega^T x_i+b)+ \theta \ l(1,\omega^T x_0+b)+\frac{\lambda}{2} ||\omega||^2
 \label{eq:general}
 \end{equation} 
-with respect to $w$ in $\RR^p$ and $b$ in $\RR$.
+with respect to $\omega$ in $\RR^p$ and $b$ in $\RR$.
 
-The traditional examplar SVM uses the hinge loss for $L$. Let us use
+The traditional exemplar SVM uses the hinge loss for $l$. Let us use
 instead the square loss. Equation~(\ref{eq:general}) becomes
 \begin{equation}
-E(w,b)=\frac{1}{2n}\sum_{i=1}^n (w\cdot x_i+b+1)^2+ \frac{1}{2}(w\cdot
-x_0+b-1)^2+\frac{\lambda}{2} ||w||^2.
+E(\omega,b)=\frac{1}{2n}\sum_{i=1}^n (\omega^T x_i+b+1)^2+ \frac{\theta}{2}(\omega^T
+x_0+b-1)^2+\frac{\lambda}{2} ||\omega||^2.
 \label{eq:square}
@@
 \begin{equation}
-b=-\frac{1}{2}(x_0+\mu)\cdot w,
+b=\dfrac{\theta-1}{\theta+1}-\dfrac{1}{\theta +1}(\theta x_0+\mu)^T \omega,
 \end{equation}
@@
 Substituting in Eq.~(\ref{eq:square}) now reduces our optimization problem
-to the minimization with respect to $w$ of
+to the minimization with respect to $\omega$ of
 \begin{equation}
-\begin{array}{lcl}\hat{E}(w)&=&
-\displaystyle\frac{1}{2n}\sum_{i=1}^n [w\cdot (x_i-\frac{1}{2}(x_0+\mu))+1]^2
-+\frac{1}{2}[w\cdot (x_0-\frac{1}{2}(x_0+\mu))-1]^2+\frac{\lambda}{2}||w||^2\\
+\begin{array}{lcl}\hat{E}(\omega)&=&
+\displaystyle\frac{1}{2n}\sum_{i=1}^n [\omega^T (x_i-\frac{1}{2}(x_0+\mu))+1]^2
++\frac{\theta}{2}[\omega^T (x_0-\frac{1}{2}(x_0+\mu))-1]^2+\frac{\lambda}{2}||\omega||^2\\
 &=&
-\displaystyle\min_{w\in\RR^p}
-\frac{1}{2n}\sum_{i=1}^n [w\cdot (\sigma_i-\frac{1}{2}\delta) +1]^2
-+\frac{1}{2}(\frac{1}{2}w\cdot\delta-1)^2+\frac{\lambda}{2}||w||^2,
+\displaystyle\min_{\omega\in\RR^p}
+\frac{1}{2n}\sum_{i=1}^n [\omega^T (\sigma_i-\frac{1}{2}\delta) +1]^2
++\frac{\theta}{2}(\frac{1}{2}\omega^T\delta-1)^2+\frac{\lambda}{2}||\omega||^2,
 \end{array}
@@
 
-Finally, writing $\partial\hat{E}/\partial w=0$ at a minimum yields
+Finally, writing $\partial\hat{E}/\partial \omega=0$ at a minimum yields
 \begin{equation}
-\frac{1}{n}\sum_{i=1}^n [w\cdot(\sigma_i-\frac{1}{2}\delta)+1](\sigma_i-\frac{1}{2}\delta)
-+\frac{1}{2}(\frac{1}{2}w\cdot\delta-1)\delta +\lambda w=0,
+\frac{1}{n}\sum_{i=1}^n [\omega^T(\sigma_i-\frac{1}{2}\delta)+1](\sigma_i-\frac{1}{2}\delta)
++\frac{1}{2}(\frac{1}{2}\omega\cdot\delta-1)\delta +\lambda \omega=0,
 \end{equation}
@@
 \begin{equation}
-[\frac{1}{n}\sum_{i=1}^n(\sigma_i-\frac{1}{2}\delta)(\sigma_i^T-\frac{1}{2}\delta^T)] w
-+\frac{1}{n}\sum_{i=1}^n (\sigma_i-\frac{1}{2}\delta) +\frac{1}{2}\delta\frac{1}{2}\delta^T w -\frac{1}{2}\delta +\lambda w
+[\frac{1}{n}\sum_{i=1}^n(\sigma_i-\frac{1}{2}\delta)(\sigma_i^T-\frac{1}{2}\delta^T)] \omega
++\frac{1}{n}\sum_{i=1}^n (\sigma_i-\frac{1}{2}\delta) +\frac{1}{2}\delta\frac{1}{2}\delta^T \omega -\frac{1}{2}\delta +\lambda \omega
 =0.
@@
 [\frac{1}{n}\sum_{i=1}^n\sigma_i\sigma_i^T +\frac{1}{2}\delta\delta^T +\lambda\text{Id}]
-w=\delta,
+\omega=\delta,
 \end{equation}
@@
 \left\{\begin{array}{l}
-\displaystyle w=A^{-1} (x_0-\mu),\quad\text{where}\quad
-A=\Sigma
+\displaystyle \omega=U^{-1} (x_0-\mu),\quad\text{where}\quad
+U=\Sigma
 +\frac{1}{2}(x_0-\mu)(x_0-\mu)^T +\lambda\text{Id},\\
-\displaystyle b=-\frac{1}{2}(x_0+\mu)\cdot w,
+\displaystyle b=-\frac{1}{2}(x_0+\mu)\cdot \omega,
 \end{array}\right.
@@
 \Sigma=\frac{1}{n}\sum_{i=1}^n(x_i-\mu)(x_i-\mu)^T
-=\frac{1}{n}\sum_{i=1}^n x_ix_i^T-\mu\mu^T=\frac{1}{n}XX^T-\mu\mu^T,
+=\frac{1}{n}\sum_{i=1}^n x_ix_i^T-\mu\mu^T=\frac{1}{n}X^TX-\mu\mu^T,
 \label{eq:Sigmadef}
@@
 is the covariance matrix of the negative samples $x_i$, and
-$X$ denotes the $p\times n$ matrix with columns $x_i^T$ ($i=1,\ldots,n$).
+$X$ denotes the $n\times p$ matrix with rows $x_i^T$ ($i=1,\ldots,n$).
 
-The vector $w$ (or the vector $[w;b]$) of this {\em square-loss
+The vector $\omega$ (or the vector $[\omega;b]$) of this {\em square-loss
 examplar machine} can potentially be used as a new representation for
@@
 \section{Kernelized version}
-Let us recall a few basic facts avout kernel methods for supervised
-classification. We consider a reprodcing kernel Hilbert space (RKHS)
+Let us recall a few basic facts about kernel methods for supervised
+classification. We consider a reproducing kernel Hilbert space (RKHS)
 $H$ formed by real functions over some set
@@
 \min_{h\in H,b\in\RR}
-\sum_{i=1}^n L[y_i,h(x_i)+b] + \frac{\lambda}{2}||h||^2,
+\sum_{i=1}^n l[y_i,h(x_i)+b] + \frac{\lambda}{2}||h||^2,
 \label{eq:kernel}
@@
 \min_{h\in H,b\in\RR}
-\sum_{i=1}^n L[y_i,( \varphi(x_i)|h)+b] +
+\sum_{i=1}^n l[y_i,( \varphi(x_i)|h)+b] +
 \frac{\lambda}{2}||h||^2,
@@
 \begin{theorem}\label{th:affinelearn}
-Equation~(\ref{eq:kernelaff}) admits equivalent primal and dual
-versions.
+Equation~(\ref{eq:kernelaff}) admits equivalent primal and dual versions.
 
@@
 \begin{equation}
-\min_{\alpha\in\RR^n,b\in\RR} \sum_{i=1}^n L[y_i,k_i^T\alpha+b]
+\min_{\alpha\in\RR^n,b\in\RR} \sum_{i=1}^n l[y_i,k_i^T\alpha+b]
 +\frac{\lambda}{2}\alpha^TK\alpha,
@@
 \begin{equation}
-\min_{\beta\in\RR^r,b\in\RR} \sum_{i=1}^n L[y_i,b_i^T\beta+b]
+\min_{\beta\in\RR^r,b\in\RR}  \sum_{i=1}^n l[y_i,b_i^T\beta+b]
 +\frac{\lambda}{2}||\beta||^2,
@@
 \end{equation}
-where $B$ is the $n\times r$ square root of $K$ and $b_i^T$ denotes
+where $B$ is the $n\times r$ incomplete Cholesky decomposition of $K$ and $b_i^T$ denotes
 its $i$th row,
@@
 \min_{\alpha\in\RR^n}
-\sum_{i=1}^n L^*(y_i,-\lambda\alpha_i)
+\sum_{i=1}^n l^*(y_i,-\lambda\alpha_i) 
 +\frac{\lambda}{2}\alpha^TK\alpha
@@
 \begin{equation}
-\min_{b\in \RR} \sum_{i=1}^n L(y_i,k_i^T\alpha^{\star}+b).
+\min_{b\in \RR} \sum_{i=1}^n l(y_i,k_i^T\alpha^{\star}+b).
 \label{eq:minb}
@@
 
-Theorem\ref{th:affinelearn} follows fromt the representer theorem
+Theorem\ref{th:affinelearn} follows from the representer theorem
 ~\cite{SHS01,Wahba90} and elementary results from analysis and
@@
 
-The next section shows how to train the kernel square-loss examplar
-machine efficiently.
-
-\section{Efficient implementation}
-Let us consider the $n\times n$ kernel matrix $\hat{K}$ associated with
-the $n$ negative examples, and a low-rank factorization $\hat{K}=\hat{B}\hat{B}^T$
-of
-this matrix, where $\hat{B}$ is an
-$n\times r$ square root of $\hat{K}$. This factorization can be computed in $O(nr^2)$ time
-and $O(nr)$ storage~\cite{BaJo05,FiSc01}. The singular value
-decomposition $UWV^T$ of $\hat{B}$ and its $r\times n$ pseudoinverse
-$\hat{B}^\dagger=(\hat{B}^T\hat{B})^{-1}\hat{B}^T=VW^{-1}U^T$ such that $\hat{B}^\dagger
-\hat{B}=\text{Id}$ can also be computed with the same time and storage
-costs.
-
-Let us consider the augmented
-kernel matrix $K$ obtained by adding one more (positive) sample. This matrix
-can be written as
-\begin{equation}
-K=\qmatrix{k_{00} & k_0^T\\k_0& \hat{K}},
-\end{equation}
-where $k_{00}$ is a scalar and $k_0$ is an element of $\RR^n$.
-The following
-lemma show that a $(n+1)\times (r+1)$ factorization of $K$ can
-be computed efficiently as well.
-
-\begin{lemma}
-The augmented kernel matrix $\hat{K}$ can be factorized as
-\begin{equation}
-\hat{K}=BB^T,\quad\text{where}\quad
-B=\qmatrix{u & v^T\\0 & \hat{B}}\quad\text{and}\quad
-v=\hat{B}^\dagger k_0,\,\, u=\sqrt{k_{00}-||v||^2}
-\end{equation}
-in time $O(nr)$ and storage $O(n)$.
-\end{lemma}
-\begin{proof}
-Indeed,
-\begin{equation}
-BB^T=\qmatrix{u^2+||v||^2 & v^T\hat{B}^T\\ \hat{B}v & \hat{B}\hat{B}^T}.
-\end{equation}
-Since $K$ is positive semidefinite $k_0$ must lie in the column space
-$\cal \hat{B}$ of $\hat{B}$,\footnote{Let
-us assume that $k_0$ does not belong to $\cal \hat{B}$, then it can be
-decomposed uniquely as $k_0=l_0+m_0$, where $l_0$ belongs to $\cal \hat{B}$,
-and $m_0\neq 0$ belong to the orthogonal complement $\cal \hat{B}^\perp$ of
-$\cal \hat{B}$. Obviously $(1,-\nu m_0^T)K(1;-\nu m_0)=k_{00}-2\nu||m_0||^2$, which
-is negative for large enough values of $\nu>0$. In turn, this contradicts the
-fact that a kernel matrix is positive semidefinite.}
-and $b=\hat{B}^\dagger k_0$ is thus an exact solution of
-$\hat{B}b=k_0$.
-The fact that $k_{00}-||b||^2$ is nonnegative follows immediately from
-the fact that the Schur complement $\hat{K}-||k_0||^2/k_{00}$ of $k_{00}$ in
-the positive semidefinite matrix $K$ is itself positive semidefinite.\footnote{Indeed,
-since the matrix $k_{00}K-k_0k_0^T=\hat{B}(k_{00}\text{Id}-bb^T)\hat{B}^T$ is
-positive semidefinite, and $\hat{B}$ has rank $r$, the $r\times r$ matrix
-$k_{00}\text{Id}-bb^T$
-is also positive definite, thus
-$b^T(k_{00}\text{Id}-bb^T)b=||b||^2(k_{00}-||b||^2)\ge 0$.} The
-cost of the factorization is dominated by the $O(nr)$ computation and $O(n)$ storage
-of $b$
-\end{proof}
-
-Training the kernel square-loss examplar machine given $\hat{B}$ amounts to
-computing
-$\Sigma$ in (\ref{eq:Sigmadef}) and solving the $(r+1)\times (r+1)$
-linear system in (\ref{eq:final}). Note that, because of the
-particular structure of $B$, only one term in the sum in (\ref{eq:Sigmadef})
-must be recomputed for each new positive example, with $O(r^2)$ cost.
-The linear system itself can be solved in $O(r^3)$ cost.
-
 Once the optimal parameters $\beta^\star$ haven been found, they can
@@
 $B^TP=\text{Id}$. The corresponding element of $\cal H$ is
-$h^\star=\sum_{i=0}^n \alpha_i^\star\varphi(x_i)$.
+$h^\star=\sum_{i=1}^n \alpha_i^\star\varphi(x_i)$.  It can be compared
+to any other element $h'=\sum_i^n\alpha_i'\varphi(x_i')$ of $\cal H$ (say, the representation of
+another image) by computing $(h^\star|h')=\alpha^{\star T}K \alpha' = \beta^{\star T}\beta'$.
 
-Now, given two positive images $x_0$ and $x_0'$, they can be compared
-by computing the normalized inner product
+\section{Incomplete Cholesky decomposition}
+
+In this section we revisit how to obtain a low-rank decomposition $B$ that better approximates $K$, as seen in ~\cite{BaJo05,FiSc01}.
+Let us assume a sequence of indexes $I=\{i_1,i_2,\dots ,i_r\}$ from $\{1,2,\dots, n\}$ is given, where $r$ is the number of columns of $B$. 
+$I$ is the set of pivots, chosen such that to minimize the error approximation between $BB^T$ and $K$.
+We also note, for $t$ in $\{1,2,...,r\}$, $I_t\{i_1,...,i_t\}$, $J_t=\{1,2,...,n\}/ I_t$ and $J=J_r$.
+
+Given $I$ and $J$, we can calculate the $t$-th column, $t=1,2,...,r$, of $B$ interactively:
+\begin{align}
+\begin{cases}
+\vspace{3 mm}
+B(i_t, t) &= \left(K(i_t,i_t)-\displaystyle{\sum_{m=1}^{t-1}} B(i_{m},m)  \right)^{\frac{1}{2}},\\
+\vspace{3 mm}
+B(I_{t-1},t) &= 0,\\
+B(J_t, t) &= \dfrac{1}{B(i_t,t)}\left(K(J_t,i_t)-\displaystyle{\sum_{m=1}^{t-1}B(J_t,m)B(i_t,m)}\right).
+\end{cases}\label{icd:algo}
+\end{align}
+The time complexity of the $t$-th interaction is $O(tn)$, thus making the full algorithm $O(nr^2)$ in time. Since $B$ is $n\times r$, storage complexity is $O(nr)$.
+\begin{proposition}
+There is an unique $n\times n$ matrix $L$ such that
+\begin{itemize}
+\item L is symmetric and positive semidefinite,
+\item L(:,I) = K(:,I),
+\item the column space of $L$ is equal to the column space of $L(:,I)$.
+\end{itemize}
+L is such that 
 \begin{equation}
-s(x_0,x_0')=\frac{(h|h')}{\sqrt{(h|h)}\sqrt{(h'|h')}}
+L([I \ J],[I\ J]) = \qmatrix{K(I,I) & K(J,I)^T\\ K(J,I) & K(J,I)K(I,I)^{-1}K(J,I)^T}.
 \end{equation}
-of the corresponding elements $h$ and $h'$ of $H$ as a similarity measure
-(we drop the ``$\star$'' exponents from now on to avoid clutter in the
-noration).
+\end{proposition}
+From \textbf{Proposition 1}, $L$ is an unique approximation of $K$ for a given pair $I$ and $J$. The algorithm from (\ref{icd:algo}) construct $B$ such that $L=BB^T$.
 
-Concretely, we have
+The next section shows how to train the kernel square-loss exemplar machine efficiently.
+
+\section{Efficient implementation of kernel decomposition}
+
+Let us consider a low-rank factorization $K=BB^T$, where $B$ as given in the previous section. 
+%The singular value decomposition $UWV^T$ of $B$ and its $r\times n$ pseudoinverse
+%$B^\dagger=(B^TB)^{-1}B^T=VW^{-1}U^T$ such that $B^\dagger B=\text{Id}$ can also be computed with the same time and storage costs.
+
+Let us consider the augmented
+kernel matrix  $K'$ obtained by adding one more sample. This matrix
+can be written as
 \begin{equation}
-\begin{array}{lcl}
-(h|h')&=&\displaystyle[\alpha_0\varphi(x_0)+\sum_{i=1}^n\alpha_i\varphi(x_i)
-|\alpha_0'\varphi(x_0')+\sum_{i=1}^n\alpha_i'\varphi(x_i)]\\
-&=&\displaystyle
-\alpha_0\alpha_0'\displaystyle k(x_0,x_0')+ \alpha_0\sum_{i=1}^n\alpha_i' k(x_0,x_i)
-+\alpha_0'\sum_{i=1}^n
-\alpha_ik(x_0',x_i)+\hat{\alpha}^T\hat{K}\hat{\alpha}'\\
-&=&\displaystyle\alpha_0\alpha_0'\displaystyle k(x_0,x_0')+ \alpha_0k_0^T\hat{\alpha}'
-+\alpha_0'k_0'^T\hat{\alpha}
-+\hat{\alpha}^T\hat{K}\hat{\alpha}'\\
-\end{array}
-\label{eq:lcl}
+K'=\qmatrix{k_{00} & k_0^T\\k_0& K},
 \end{equation}
-where $\hat{\alpha}$ (resp. $\hat{\alpha}'$) denotes the vector of
-$\RR^n$ formed by excluding the first (or rather zeroth) element
-of $\alpha$ (resp. $\alpha'$).
+where $k_{00}$ is a scalar and $k_0$ is an element of $\RR^n$.
+The following
+lemma shows that a $(n+1)\times (r+1)$ factorization of $K'$ can
+be computed efficiently as well.
+In order to ease notations, we assume $[k_{00}; k_0]$ and $[k_{00}; k_0]^T$ are respectively the $0$-th column and row of $K'$.
+Hence, the $i$-th row of $K$ is part of the $i$-th row of $K'$, $i=\{1,2,...,n\}$.
+Also, let $I'=\{0\}\cup I$ be the set of pivots for the decomposition of $K'$.
 
-If we write $\beta = [\beta_0; \hat{\beta}]$ similarly to
-$\alpha=[\alpha_0; \hat{\alpha}]$ and keeping in mind that
-$P=[\frac{1}{u}\ 0; -\frac{1}{u}\hat{P}v \ \hat{P}]$, where $\hat{P}$
-is the pseudoinverse of
-$\hat{B}^T$, we have that
+\begin{lemma}
+The augmented kernel matrix $K'$ can be factorized as
+\begin{align}
+K'&\approx B'B'^T,\quad\text{where}\quad
+B'=\qmatrix{u & v^T\\w & B}\quad\text{and}\\
+v=B_I^{-1} k_0(I),\,\, u&=\sqrt{k_{00}-||v||^2},\,\, w(I) = 0,\,\, w(J)=\dfrac{1}{u}(k_0(J)-B_Jv)
+\end{align}
+in time $O(nr)$ and storage $O(n)$. Here, $B_I$ and $B_J$ denote $B(I,:)$ and $B(J,:)$, respectively.
+\end{lemma}
+\begin{proof}
+Let $L' = B'B'^T$. Indeed, 
 \begin{equation}
-\alpha_0 = \dfrac{\beta_0}{u}, \,\, \hat{\alpha}=-\dfrac{\beta_0}{u}Pv+P\hat{\beta}.
+L'=\qmatrix{u^2+||v||^2 & v^TB^T+uw^T\\ Bv+uw & BB^T+ww^T}. 
 \end{equation}
-Remembering that $\hat{K}=\hat{B}\hat{B}^T$ and and $k_0=\hat{B}v$
-allows us to rewrite (\ref{eq:lcl}) as
+Following \textbf{Proposition 1}, if $L'(I,:)=K'(I,:)$ and $L'(:,I)=K'(:,I)$, $w(I)=0$.
+From the fact that $B$ has full rank, $B_I$ is a $r\times r$ matrix of rank $r$.
+$B_I$ is therefore invertible, which guarantees the system $B_Iv=k_0(I)$ has a solution given by $v=B_I^{-1} k_0(I)$.
+To show that $k_{00}\ge ||v||^2$, we first note the 
+\begin{equation}
+K'(I',I') = \qmatrix{k_{00} & k_0(I)^T\\ k_0(I)& K(I,I)}
+\end{equation}
+is positive semidefinite. Thus the Schur complement of $k_{00}$ is also non-negative:
+\begin{equation}
+k_{00}-k_0(I)^TK(I,I)^{-1}k_0(I)\ge 0,
+\end{equation}
+from which it follows that $k_{00}-||v||^2\ge 0$. \footnote{$k_0(I)^TK(I,I)^{-1}k_0(I)=k_0(I)^T(B_IB_I^T)^{-1}k_0(I)=v^Tv$}
+$w(J)$ follows from $B_Jv+uw(J)=k_0(J).$
+
+The cost of the factorization is dominated by the $O(nr)$ computation and $O(n)$ storage
+of $w$.
+\end{proof}
+
+\subsection{Why $w\ne 0$?}
+In a previous version of this presentation, it argued for a decomposition of $K'$ with $w=0$.
+One of the keys justifications of this assumption is that $B^\dagger v=k_0$ has a solution because $k_0$ lie in the column space $\cal B$ of $B$. 
+This is not true however for a low-rank decomposition $B$, i.e. $r$ is smaller the rank of $K$. 
+Indeed, $K'$ is a positive semidefinite matrix, but $[k_{00}\ k_0^T; k_0\ BB^T]$ is not. 
+If $r$ is the rank of $K$, then we do find $w=0$.
+But in practice, for $k_0$ does not belong in $\cal B$ and $k_{00}<||v||^2$ for $r$ not big enough.
+
+\section{Efficient implementation of ridge regression}
+Let us rewrite the kernelized problem after adding the positive exemplar to Equation (\ref{eq:primal2}):
+\begin{equation}
+\min_{\beta\in\RR^{r+1}, \ b\in\RR}\left( \theta (1-\beta^Tb_0'-b)^2+\dfrac{1}{n}\sum_{i=1}^n (1+\beta^Tb_i'+b)^2 +||\beta||^2\right).
+\end{equation}
+We recall the solution (\ref{eq:final}):
 \begin{align}
-(h|h')& = \dfrac{\beta_0\beta_0'}{uu'}k(x_0,x_0')
-+\dfrac{\beta_0}{u}v^T\hat{B}^T\left(-\dfrac{\beta_0'}{u'}Pv'+P\hat{\beta'}\right)
-+\dfrac{\beta_0'}{u'}v'^T\hat{B}^T\left(-\dfrac{\beta_0}{u}Pv+P\hat{\beta}\right)\\ &+\left(\hat{B}^T\left(-\dfrac{\beta_0}{u}Pv+P\hat{\beta}\right)\right)^T
-\left(\hat{B}^T\left(-\dfrac{\beta_0'}{u'}Pv'+P\hat{\beta'}\right)\right)\\ &=
-\dfrac{\beta_0\beta_0'}{uu'}k(x_0,x_0')
-+\dfrac{\beta_0}{u}v^T\left(-\dfrac{\beta_0'}{u'}v'+\hat{\beta'}\right)
-+\dfrac{\beta_0'}{u'}v'^T\left(-\dfrac{\beta_0}{u}v+\hat{\beta}\right)
-+
-\left(-\dfrac{\beta_0}{u}v+\hat{\beta}\right)^T\left(-\dfrac{\beta_0'}{u'}v'+\hat{\beta'}\right)\\ &=
-\hat{\beta}^T\hat{\beta'}
-+\dfrac{\beta_0\beta_0'}{uu'}\left(k(x_0,x_0')-v^Tv'\right).
+\left\{\begin{array}{l}
+\vspace{5 mm}
+\displaystyle \beta^\star=\dfrac{2\theta}{\theta+1}U^{-1} \delta,\quad\text{where}\quad
+U=\dfrac{1}{n}\hat{B}^T\hat{B}-\mu'\mu'^T
++\frac{\theta}{\theta+1}\delta\delta^T +\lambda\text{Id},\\
+\displaystyle b^\star=\dfrac{\theta-1}{\theta+1}-\frac{1}{\theta+1}(b'_0+\mu')^T \omega,
+\end{array}\right.
+\label{kernel:final}
 \end{align} 
-We also note that $(h|h)=||\beta||^2$ and $(h'|h')=||\beta'||^2$.
+Here, $b_i'^T$ is the $i$-th row of $B'$
+\footnote{As for $K'$, we start counting at $0$ for $B'$ indexes.}, $\hat{B}$ is the matrix of rows $b_i'^T$, $i$ in $\{1,2,...,n\}$,
+$\mu' = \frac{1}{n}\sum_{i=1}^nb'_i$ and $\delta = b'_0-\mu'$.
+In parallel with the notations for $B$, we note as well $b_i^T$ is the $i$-th row of $B$, $\mu = \frac{1}{n}\sum_{i=1}^nb_i$.
 
-Furthermore, from the solution of the linear system (\ref{eq:final}), we find that $\beta_0 = \frac{cu}{\lambda}$, $\hat{\beta} = cG^{-1}(v-\mu)$ and $\beta_0' = \frac{cu'}{\lambda}$, $\hat{\beta}' = c'G^{-1}(v'-\mu)$, for $c,\ c'$ scalars and $G = \frac{1}{n}\hat{B}^T\hat{B}-\mu\mu^T+\lambda\text{Id}$.
-The final similarity score can be computed by
+We note $A$ the matrix $(n+1)\times (n+1)$ such that
 \begin{equation}
-s(x_0,x_0') = \dfrac{(v-\mu)^TG^{-2}(v'-\mu)+\dfrac{1}{\lambda^2}(k(x_0,x_0')-v^Tv')}{\sqrt{(v-\mu)^TG^{-2}(v-\mu)+\dfrac{u^2}{\lambda^2}}\sqrt{(v'-\mu)^TG^{-2}(v'-\mu)+\dfrac{u'^2}{\lambda^2}}}
+A = \dfrac{1}{n}B'^TB'-\mu'\mu'^T+\lambda\text{Id}.\label{eq:A}
 \end{equation}
 
+Training the kernel square-loss exemplar machine given $B$ amounts to computing
+$U^{-1}$ from $A^{-1}$ since $U^{-1} = \left(A+0.5\delta\delta^T\right)^{-1}$
+and solving the $(r+1)\times (r+1)$ linear system in (\ref{kernel:final}). The linear system itself can be solved in time $O(r^3)$.
+Note that, because of the structure of $A$, only its first row and first column have to be recalculated for each given positive exemplar. 
+Indeed,
+\begin{equation}
+A = \qmatrix{a_{00} & a_0^T\\ a_0 & G},
+\end{equation}
+where $a_{00}=\dfrac{1}{n}(w^Tw+\text{1}^Tw)+\lambda$, $a_0 = \dfrac{1}{n}(B^Tw+\text{1}^Tw)$ and $G=\dfrac{1}{n}B^TB-\mu\mu^T+\lambda\text{Id}_r$.
+
+$G$ depends only on $B$ and can be calculated in time $O(nr^2)$ and storage $O(nr)$ before adding exemplars. Its inverse $G^{-1}$ is computed at cost $O(r^3)$.
+Woodbury identity allows the computation of $A^{-1}$ from $G^{-1}$ at cost $O(r^2)$:
+\begin{equation}
+A^{-1} = \qmatrix{\gamma & -\gamma a_0^TG^{-1}\\ -\gamma G^{-1}a_0 & G^{-1}+\gamma G^{-1}a_0a_0^TG^{-1}},
+\end{equation}
+where $\gamma = \left(a_{00}-a_0^TG^{-1}a_0\right)^{-1}$.
+
+Using Woodbury again, we obtain $U^{-1}$ from $A^{-1}$ at the same cost:
+\begin{equation}
+U^{-1} = A^{-1} - \dfrac{1}{\delta^TA^{-1}\delta+2}A^{-1}\delta \delta^TA^{-1}.
+\end{equation}
+
+Therefore $\beta^\star$ is computed at time cost $O(r^2)$ for each positive exemplar, which corresponds to the cost of applying the Woodbury identity two times,
+instead of solving the system at (\ref{kernel:final}) at cost $O(r^3)$.
 \section{Notes: LDA and square-loss linear classifiers}
@@
 
-\section{SLEM versus whitening}
-Due to Woodbury identity, Equation \ref{eq:final} can be rewritten as 
+\subsection{Comparison between SLEM and LDA scores}
+Let us reconsider the non-kernelized SLEM of subsection \ref{SLEM}.
+In this case, $A = \frac{1}{n}XX^T-\mu\mu^T+\lambda\textbf{Id}_d$ is the covariance matrix of the negative samples $X$\footnote{A small term proportional to the identity matrix is added to insure the covariance matrix is positive-definite.}.
+After application of Equation (\ref{beta:fromwood}) to obtain $\omega^\star(x_0,X)$, we can calculate the similarity score $s(x_0,x_0')$ between a query image $x_0$ and an other image $x_0'$ as the inner product of its SLEM vectors:
 \begin{equation}
-w = \Sigma^{-1}(x_0-\mu).
+s(x_0,x_0') = (x_0-\mu)^TA^{-2}(x_0'-\mu).
 \end{equation}
-In the original E-SVM work, the score of similarity between a query image $x_0$ and a possible result $x_0'$ is given by
+One interesting fact is that a similar similarity score can be obtained by applying LDA to the positive data. 
+Indeed, $\hat{s}(x_0,x_0') = (x_0-\mu)^TA^{-1}(x_0'-\mu)$ is the similarity score for LDA-transformed vectors $x_0$ and $x_0'$. 
+\cite{Koba15} noticed that $\hat{s}$ corresponds to the similarity score of SLEM vectors if we use the similarity score proposed in \cite{Efros11} to zero-centered data.
+
+In order to compare both scores, we introduce the similarity score 
 \begin{equation}
-s(x_0,x_0') = x_0'^T\Sigma^{-1}(x_0-\mu). \label{white:score}
+s_{\alpha}(x_0,x_0') = (x_0-\mu)^TA^{\alpha}(x_0'-\mu),
 \end{equation}
-
-Under the assumption centered data and positive and negative data having the same covariance matrix (as proposed in the previous section), Equation \ref{white:score} corresponds to the score of whitened data:
-\begin{align}
-\hat{x_0} = \Sigma^{-\frac{1}{2}}(x_0-mu);\\
-s(x_0,x_0') = \hat{x_0}^T\hat{x_0'}.
-\end{align}
-This observation has already been made by \cite{Koba15}. However, the score of similarity used in \cite{ZePe15} is
-\begin{equation}
-s(x_0,x_0') = (x_0'-\mu)^T\Sigma^{-2}(x_0-\mu). \label{esvm:score}
-\end{equation}
-
-Based on this, we propose the score $s_{\alpha}$ given by $s_{\alpha}(x_0,x_0') = (x_0-\mu)^T\Sigma^{\alpha}(x_0-\mu).$ $s_{-2}$ is the SLEM similarity score, $s_{-1}$ is the whitening similarity score and $s_{0}$ is the centered base features similarity score.
-
-Strangely, both scores seem to perform similarly, even if their scores differ. Doing a singular value decomposition of $\Sigma$, $\Sigma^{-1}=VD^{-1}V^T$ and $\Sigma^{-\alpha}=VD^{-\alpha}V^T$.
-
+and we study how the mAP of \emph{Holidays} evolves when we vary $\alpha$.
+Notice that $s=s_{-2}$, $\hat{s}=s_{-1}$ and $s_0$ corresponds to zero-centered base features inner product.
+We also compare both scores to the score obtained by whitening the data (\emph{i.e.}, $\mu$ and $A$ are, respectively, the mean and covariance matrix of the assembled positive and negative data) and by the original base features inner product (\emph{i.e.}, $x_0^Tx_0'$).
+Performing this test on VLAD and CNN feature, as seen on Figure \ref{salpha:vladcnn}, we observe the mAP plateauing for $\alpha$ in $[-2,-1]$, suggesting there is no real improvement in using non-kernelized SLEM over LDA, or, for that matter, the similarity score of \cite{ZePe15}.
+(which is the most important contribution of this paper) over the similarity of \cite{Efros11}.
+\emph{\color{red} I'd need to test with E-SVM scores, but if this is true, Patrick's paper is kind of pointless!}
+On the other hand, the spatial pyramid features improves its results for smaller values of $\alpha$ in $[-4,0]$, as seen on Figure \ref{salpha:pyr}.
 \begin{figure}[!h]
@@
 \caption{Comparison of mAP for different similarity scores $s_{\alpha}$, varying $\alpha$.
-At left, CNN features.
-At right, VLAD features.
-On blue, $\Sigma$ is the covariance of the negative samples and on red, $\Sigma$ is the covariance of the positive and negative samples.
-On black, the mAP for similarity calculated as inner product of base features.}
-\label{Sigma}
+On the top, CNN features.
+At the bottom, VLAD features.
+On blue, $A$ is the covariance of the negative samples and on red, $A$ is the covariance of the positive and negative samples.
+In black, the mAP for similarity calculated as inner product of base features.}
+\label{salpha:vladcnn}
 \end{figure}
 
-\section{Experiment Evaluation}
-\subsection{Base image encoding}
-We use three base features as the representation in $\RR^p$ of our images. The two firsts are VLAD and CNN features, as used in \cite{ZePe15}. 
-CNN features are non-negative and can be used both as $\mathbb{L}^1$ or $\mathbb{L}^2$ normalized features.
-The third features are pyramids of histograms of SIFT descriptors, as used in \cite{spk}.
-These are non-negative $\mathbb{L}^1$ normalized features.
-
-\subsection{Kernels}
-For all features, we test non-kernelized SLEM, as well as kernelized SLEM for gaussian (rbf) and polynomial kernel:
-\begin{align}
-k_{rbf}(x,y) &= \exp\left(-\gamma||x-y||^2\right);\\
-k_{poly}(x,y) &= x^Ty+\gamma(x^Ty)^2.
-\end{align}
-For non-negative $\mathbb{L}^1$ normalized, we also test the intersection kernel: $k_{inter}(x,y) = \min(x,y).$\\
-We observe that the non-kernelized obtains similar results to kernels polynomial and Gaussian. For the pyramid histogram of SIFT, the intersection kernel works much better than all the other kernels.
-
-\begin{table}
-\begin{center}
-\begin{tabular}{|l|c|c|c|c|}
-\hline
-Method & $s_0$ & $s_{-2}$ with no kernel & $s_{-2}$ for gaussian kernel &  $s_{-2}$ for intersection kernel \\
-\hline\hline
-VLAD-64 & 73.3 & 77.9 & 78.1 & -\\
-Caffe CNN & 69.7 & 72.1 & 72.9 & 70.2\\
-Pyramid of Histogram & 38.7 & 43 & 44.6 & 63.5\\
-\hline
-\end{tabular}
-\end{center}
-\caption{Results.}
-\end{table}
+\begin{figure}
+\centering
+\includegraphics[width=.45\textwidth]{whitening_pyr.jpg}
+\caption{Comparison of mAP for different similarity scores $s_{\alpha}$, varying $\alpha$, for spatial pyramid of SIFT descriptors.}
+\label{salpha:pyr}
+\end{figure}
 

### 2016-11-07 ichol_writeup.tex
first commit

--- 2016-11-07 esvm.tex
@@
 \title{Kernel Square-Loss Examplar Machines}
-\author{Jean Ponce}
+\author{Rafael Rezende}
 %\date{}
@@
 
-\begin{abstract}
-Zepeda and P\'erez show in~\cite{ZePe15} that the weight vectors
-associated with examplar SVMs can be used as global features
-representing images in tasks such as retrieval or classification. A
-drawback is the cost of the procedure, which requires a (convex)
-optimization process to be run for each positive image. This note
-shows that a more efficient approach can be obtained using a square
-loss instead of the hinge loss. Building on~\cite{BaJo05,FiSc01}, it
-also shows that this approach is easily kernelized, and, assuming
-low-rank kernels, that the corresponding computational cost is linear
-in the number of negative examples for each positive example. The
-presentation concludes with a brief discussion of the relationship
-between the proposed approach with linear discriminant analysis and
-its use in region matching in~\cite{ARS14}.
-\end{abstract}
+\section{Incomplete Cholesky decomposition}
 
-\section{Square-loss examplar machines\label{sec:sqesvm}}
-Given some loss function $L:[-1,+1]\times \RR\rightarrow \RR^+$, some
-positive training example $x_0$ in $\RR^p$ and $n$ negative training
-examples $x_i$ in $\RR^p$ ($i=1,\ldots,n$), training an exemplar
-linear classifier consists in minimizing the error
+In previous write-ups, we wanted to extend a $n\times r$ low-rank decomposition of a $n\times n$ kernel matrix $K$ into a $(n+1)\times (r+1)$ of the matrix 
 \begin{equation}
-E(w,b)=\frac{1}{n}\sum_{i=1}^n L(-1,w\cdot x_i+b)+ L(1,w\cdot
-x_0+b)+\frac{\lambda}{2} ||w||^2
-\label{eq:general}
-\end{equation}
-with respect to $w$ in $\RR^p$ and $b$ in $\RR$.
-
-The traditional examplar SVM uses the hinge loss for $L$. Let us use
-instead the square loss. Equation~(\ref{eq:general}) becomes
-\begin{equation}
-E(w,b)=\frac{1}{2n}\sum_{i=1}^n (w\cdot x_i+b+1)^2+ \frac{1}{2}(w\cdot
-x_0+b-1)^2+\frac{\lambda}{2} ||w||^2.
-\label{eq:square}
+K'=\begin{pmatrix}
+K & k_0 \\
+k_0^T & k_{00}
+\end{pmatrix}.
 \end{equation}
 
-The interest of the square loss is that the minimization can be done
-in closed form. Indeed, writing that $\partial E/\partial b=0$ at a
-minimum yields
+We assumed the $(n+1)$-th column of this matrix decomposition $B'$ would be all zero except for the last writing
 \begin{equation}
-b=-\frac{1}{2}(x_0+\mu)\cdot w,
+B'=\begin{pmatrix}
+B & 0 \\
+v^T & u
+\end{pmatrix},\label{eq:old}
 \end{equation}
-where $\mu=\frac{1}{n}\sum_{i=1}^n x_i$ is the center of mass of the
-negative samples $x_i$.
+where $B$ is the incomplete Cholesky decomposition (ICD) of $K$.
+However, the pivot algorithm for ICD states $B'$ should be written as
+\begin{equation}
+B'=\begin{pmatrix}
+B & w \\
+v^T & u
+\end{pmatrix},\label{eq:new}
+\end{equation}
+with $u$ in $\RR$, $v$ in $\RR^r$ and $w$ in $\RR^n$. 
+Decomposing $K'\approx B'B'^T$ gives us 
+\begin{align}
+\begin{cases}
+Bv+uw=k_0,\\
+u+||v||^2=k_{00}.\label{syst:comp}
+\end{cases}
+\end{align}
 
-Substituting in Eq.~(\ref{eq:square}) now reduces our optimization problem
-to the minimization with respect to $w$ of
-\begin{equation}
-\begin{array}{lcl}\hat{E}(w)&=&
-\displaystyle\frac{1}{2n}\sum_{i=1}^n [w\cdot (x_i-\frac{1}{2}(x_0+\mu))+1]^2
-+\frac{1}{2}[w\cdot (x_0-\frac{1}{2}(x_0+\mu))-1]^2+\frac{\lambda}{2}||w||^2\\
-&=&
-\displaystyle\min_{w\in\RR^p}
-\frac{1}{2n}\sum_{i=1}^n [w\cdot (\sigma_i-\frac{1}{2}\delta) +1]^2
-+\frac{1}{2}(\frac{1}{2}w\cdot\delta-1)^2+\frac{\lambda}{2}||w||^2,
-\end{array}
-\label{eq:weq}
-\end{equation}
-where $\delta=x_0-\mu$ and $\sigma_i=x_i-\mu$.
+Let $I_n=\{i_1,i_2,...,i_r\}\subset \{1,2,...,n\}$ be the set of pivots of $B$ and $J_n$ be the sorted complement if $I_n$. 
+We add one more sample and we wish to decompose $K'$ in a $(n+1)\times (r+1)$ ICS. 
+For this, we add another pivot, making $I_{n+1} = I_n\cup \{n+1\}$ the set of pivots of $K'$. Following the ICS algorithm gives us a solution to Equation (\ref{syst:comp}) in the form of 
+\begin{align}
+\begin{cases}
+\vspace{3 mm}
+w(I_n) = 0, v = B(I_n,:)^+k_0(I_n),\\
+\vspace{3 mm}
+u = \sqrt{k_{00}-||v||^2},\\
+w(J_n) = \dfrac{1}{u}\left(K_0(J_n)-B(J_n,:)v \right),
+\end{cases}
+\end{align}\label{syst:sol}
+where $B(I,:)$ is the block matrix of $B$ with rows $I$ and all columns.\\
+Equation (\ref{syst:sol}) follows the same solution we had for (\ref{eq:old}), except by the third line. $B'$ from Equations (\ref{eq:old}) and (\ref{eq:new}) do not give same approximation of $K'$.
 
-Finally, writing $\partial\hat{E}/\partial w=0$ at a minimum yields
-\begin{equation}
-\frac{1}{n}\sum_{i=1}^n [w\cdot(\sigma_i-\frac{1}{2}\delta)+1](\sigma_i-\frac{1}{2}\delta)
-+\frac{1}{2}(\frac{1}{2}w\cdot\delta-1)\delta +\lambda w=0,
-\end{equation}
-or
-\begin{equation}
-[\frac{1}{n}\sum_{i=1}^n(\sigma_i-\frac{1}{2}\delta)(\sigma_i^T-\frac{1}{2}\delta^T)] w
-+\frac{1}{n}\sum_{i=1}^n (\sigma_i-\frac{1}{2}\delta) +\frac{1}{2}\delta\frac{1}{2}\delta^T w -\frac{1}{2}\delta +\lambda w
-=0.
-\end{equation}
+\section{Real data}
+With the ICS algorithm applied to real data ($4096$ dimensional normalized CNN features from Holidays dataset) for a big enough sample of data $n$, we consistently find $r=4096$, which is smaller than $n$ but not low dimensional.
+Also, when adding the $(r+1)$-th dimension of $B'$, we find that $||v||=k_{00}$ and $w$ is non defined from Equation (\ref{syst:sol}).
 
-Using the fact that the variables $\sigma_i$ have zero
-mean, we obtain
-\begin{equation}
-[\frac{1}{n}\sum_{i=1}^n\sigma_i\sigma_i^T +\frac{1}{2}\delta\delta^T +\lambda\text{Id}]
-w=\delta,
-\end{equation}
-or, putting everything together and reverting to the original notation,
-\begin{equation}
-\left\{\begin{array}{l}
-\displaystyle w=A^{-1} (x_0-\mu),\quad\text{where}\quad
-A=\Sigma
-+\frac{1}{2}(x_0-\mu)(x_0-\mu)^T +\lambda\text{Id},\\
-\displaystyle b=-\frac{1}{2}(x_0+\mu)\cdot w,
-\end{array}\right.
-\label{eq:final}
-\end{equation}
-where
-\begin{equation}
-\Sigma=\frac{1}{n}\sum_{i=1}^n(x_i-\mu)(x_i-\mu)^T
-=\frac{1}{n}\sum_{i=1}^n x_ix_i^T-\mu\mu^T=\frac{1}{n}XX^T-\mu\mu^T,
-\label{eq:Sigmadef}
-\end{equation}
-is the covariance matrix of the negative samples $x_i$, and
-$X$ denotes the $p\times n$ matrix with columns $x_i^T$ ($i=1,\ldots,n$).
+\section{Eliminating negatives from negative encoding}
 
-The vector $w$ (or the vector $[w;b]$) of this {\em square-loss
-examplar machine} can potentially be used as a new representation for
-the corresponding positive feature $x_0$. Note that this gives
-asymmetric roles to positive and negative examples.
-
-A similar approach (with no regularizer) is advocated for region
-matching in~\cite{ARS14}. We will come back to this issue in the last
-section of this presentation.
-
-\section{Kernelized version}
-Let us recall a few basic facts avout kernel methods for supervised
-classification. We consider a reprodcing kernel Hilbert space (RKHS)
-$H$ formed by real functions over some set
-$X$, and denote by $k$ the corresponding reproducing kernel. We
-address the following learning problem over $H\times\RR$:
-\begin{equation}
-\min_{h\in H,b\in\RR}
-\sum_{i=1}^n L[y_i,h(x_i)+b] + \frac{\lambda}{2}||h||^2,
-\label{eq:kernel}
-\end{equation}
-where the pairs $(x_i,y_i)$ ($i=1,\ldots,n$) in $X\times Y$ are
-training samples, $Y=\RR$ (regression) or $Y=\{-1,1\}$
-(classification), and $L: Y\times \RR\rightarrow\RR$ is some arbitrary
-loss function. By definition of a reproducing kernel,
-Eq.~(\ref{eq:kernel}) can be rewritten as
-\begin{equation}
-\min_{h\in H,b\in\RR}
-\sum_{i=1}^n L[y_i,( \varphi(x_i)|h)+b] +
-\frac{\lambda}{2}||h||^2,
-\label{eq:kernelaff}
-\end{equation}
-where $\varphi$ is the {\em feature map} over $X$ associated with the
-kernel $k$ (which may not admit a known explicit form). We dub
-problems with the general form of (\ref{eq:kernelaff}) {\em affine}
-supervised learning problems since, given some fixed element $h$ of
-$H$ and some scalar $b$, $(h|h')+b$ is an affine function of $h'$,
-whose zero set defines an affine hyperplane of $H$ considered itself
-as an affine space.
-
-Let $K$ denote the kernel matrix with entries $k_{ij}=[\varphi(x_i)|
-\varphi(x_j)]$ and rows $k_i^T$ ($i=1,\ldots,n$). We assume from now
-on that $L$ is convex and continuous. We have the following theorem.
-
-\begin{theorem}\label{th:affinelearn}
-Equation~(\ref{eq:kernelaff}) admits equivalent primal and dual
-versions.
-
-The primal problem is
-\begin{equation}
-\min_{\alpha\in\RR^n,b\in\RR} \sum_{i=1}^n L[y_i,k_i^T\alpha+b]
-+\frac{\lambda}{2}\alpha^TK\alpha,
-\label{eq:primal1}
-\end{equation}
-and any solution $(\alpha^\star,b^\star)$ to (\ref{eq:primal1})
-provides
-a solution $(h^\star,b^\star)$ to (\ref{eq:kernelaff}) with
-$h^\star=\sum_{i=1}^n \alpha_i^\star\varphi(x_i)$.
-
-Let $r$ denote the rank of the kernel matrix $K$, a
-{\em reduced} version of the primal problem is
-\begin{equation}
-\min_{\beta\in\RR^r,b\in\RR} \sum_{i=1}^n L[y_i,b_i^T\beta+b]
-+\frac{\lambda}{2}||\beta||^2,
-\label{eq:primal2}
-\end{equation}
-where $B$ is the $n\times r$ square root of $K$ and $b_i^T$ denotes
-its $i$th row,
-and any solution $(\beta^\star,b^\star)$ to (\ref{eq:primal2})
-provides
-a solution $(h^\star,b^\star)$ to (\ref{eq:kernelaff}) with
-$h^\star=\sum_{i=1}^n \alpha_i^\star\varphi(x_i)$, and
-$\alpha^\star=P\beta^\star$, where $P$ is the pseudo-inverse of $B^T$.
-
-The dual problem is
-\begin{equation}
-\min_{\alpha\in\RR^n}
-\sum_{i=1}^n L^*(y_i,-\lambda\alpha_i)
-+\frac{\lambda}{2}\alpha^TK\alpha
-\quad\text{subject to}\quad \sum_{i=1}^n \alpha_i=0,
-\label{eq:finaldual}
-\end{equation}
-and the optimal value $b^*$ of $b$ given any solution
-$\alpha^\star$ of (\ref{eq:finaldual}) can be found by solving
-\begin{equation}
-\min_{b\in \RR} \sum_{i=1}^n L(y_i,k_i^T\alpha^{\star}+b).
-\label{eq:minb}
-\end{equation}
-The pair $(\alpha^\star,b^\star)$ then provides a solution
-$(h^\star,b^\star)$ to (\ref{eq:kernelaff}) with $h^\star=
-\sum_{i=1}^n\alpha_i^\star\varphi(x_i)$.
-\end{theorem}
-
-Note that, given a solution $(\alpha^\star,b^\star)$ of either the
-primal or the dual version of (\ref{eq:kernelaff}), the optimal
-prediction function is
-\begin{equation}
-[h^\star|\varphi(x)]+b^\star= \sum_{i=1}^n \alpha_i^{\star} K(x,x_i) +b^\star.
-\end{equation}
-
-Theorem\ref{th:affinelearn} follows fromt the representer theorem
-~\cite{SHS01,Wahba90} and elementary results from analysis and
-optimization theory.
-
-In our context, the reduced primal version of Theorem 1 is
-particularly interesting since the results of Section~\ref{sec:sqesvm}
-directly apply to it: One simply has to replace $x_0$ and $\mu$ in
-Eq.~(\ref{eq:final}) by appropriate columns of $B^T$, or linear
-combinations thereof (in this case of course, the kernel matrix has
-dimension $(n+1\times n+1)$). It is of course not clear whether the
-corresponding vector $w$ would prove useful as a feature for
-classification or other tasks.
-
-The next section shows how to train the kernel square-loss examplar
-machine efficiently.
-
-\section{Efficient implementation}
-Let us consider the $n\times n$ kernel matrix $\hat{K}$ associated with
-the $n$ negative examples, and a low-rank factorization $\hat{K}=\hat{B}\hat{B}^T$
-of
-this matrix, where $\hat{B}$ is an
-$n\times r$ square root of $\hat{K}$. This factorization can be computed in $O(nr^2)$ time
-and $O(nr)$ storage~\cite{BaJo05,FiSc01}. The singular value
-decomposition $UWV^T$ of $\hat{B}$ and its $r\times n$ pseudoinverse
-$\hat{B}^\dagger=(\hat{B}^T\hat{B})^{-1}\hat{B}^T=VW^{-1}U^T$ such that $\hat{B}^\dagger
-\hat{B}=\text{Id}$ can also be computed with the same time and storage
-costs.
-
-Let us consider the augmented
-kernel matrix $K$ obtained by adding one more (positive) sample. This matrix
-can be written as
-\begin{equation}
-K=\qmatrix{k_{00} & k_0^T\\k_0& \hat{K}},
-\end{equation}
-where $k_{00}$ is a scalar and $k_0$ is an element of $\RR^n$.
-The following
-lemma show that a $(n+1)\times (r+1)$ factorization of $K$ can
-be computed efficiently as well.
-
-\begin{lemma}
-The augmented kernel matrix $\hat{K}$ can be factorized as
-\begin{equation}
-\hat{K}=BB^T,\quad\text{where}\quad
-B=\qmatrix{u & v^T\\0 & \hat{B}}\quad\text{and}\quad
-v=\hat{B}^\dagger k_0,\,\, u=\sqrt{k_{00}-||v||^2}
-\end{equation}
-in time $O(nr)$ and storage $O(n)$.
-\end{lemma}
-\begin{proof}
-Indeed,
-\begin{equation}
-BB^T=\qmatrix{u^2+||v||^2 & v^T\hat{B}^T\\ \hat{B}v & \hat{B}\hat{B}^T}.
-\end{equation}
-Since $K$ is positive semidefinite $k_0$ must lie in the column space
-$\cal \hat{B}$ of $\hat{B}$,\footnote{Let
-us assume that $k_0$ does not belong to $\cal \hat{B}$, then it can be
-decomposed uniquely as $k_0=l_0+m_0$, where $l_0$ belongs to $\cal \hat{B}$,
-and $m_0\neq 0$ belong to the orthogonal complement $\cal \hat{B}^\perp$ of
-$\cal \hat{B}$. Obviously $(1,-\nu m_0^T)K(1;-\nu m_0)=k_{00}-2\nu||m_0||^2$, which
-is negative for large enough values of $\nu>0$. In turn, this contradicts the
-fact that a kernel matrix is positive semidefinite.}
-and $b=\hat{B}^\dagger k_0$ is thus an exact solution of
-$\hat{B}b=k_0$.
-The fact that $k_{00}-||b||^2$ is nonnegative follows immediately from
-the fact that the Schur complement $\hat{K}-||k_0||^2/k_{00}$ of $k_{00}$ in
-the positive semidefinite matrix $K$ is itself positive semidefinite.\footnote{Indeed,
-since the matrix $k_{00}K-k_0k_0^T=\hat{B}(k_{00}\text{Id}-bb^T)\hat{B}^T$ is
-positive semidefinite, and $\hat{B}$ has rank $r$, the $r\times r$ matrix
-$k_{00}\text{Id}-bb^T$
-is also positive definite, thus
-$b^T(k_{00}\text{Id}-bb^T)b=||b||^2(k_{00}-||b||^2)\ge 0$.} The
-cost of the factorization is dominated by the $O(nr)$ computation and $O(n)$ storage
-of $b$
-\end{proof}
-
-Training the kernel square-loss examplar machine given $\hat{B}$ amounts to
-computing
-$\Sigma$ in (\ref{eq:Sigmadef}) and solving the $(r+1)\times (r+1)$
-linear system in (\ref{eq:final}). Note that, because of the
-particular structure of $B$, only one term in the sum in (\ref{eq:Sigmadef})
-must be recomputed for each new positive example, with $O(r^2)$ cost.
-The linear system itself can be solved in $O(r^3)$ cost.
-
-Once the optimal parameters $\beta^\star$ haven been found, they can
-be used directly as some representation of the positive example. The
-corresponding value $\alpha^\star$ (or more correctly, {\em a}
-corresponding value) can be computed as $\alpha^\star =P\beta^\star$,
-where $P=B(B^TB)^{-1}$ is the pseudoinverse of $B^T$ with
-$B^TP=\text{Id}$. The corresponding element of $\cal H$ is
-$h^\star=\sum_{i=0}^n \alpha_i^\star\varphi(x_i)$.
-
-Now, given two positive images $x_0$ and $x_0'$, they can be compared
-by computing the normalized inner product
-\begin{equation}
-s(x_0,x_0')=\frac{(h|h')}{\sqrt{(h|h)}\sqrt{(h'|h')}}
-\end{equation}
-of the corresponding elements $h$ and $h'$ of $H$ as a similarity measure
-(we drop the ``$\star$'' exponents from now on to avoid clutter in the
-noration).
-
-Concretely, we have
-\begin{equation}
-\begin{array}{lcl}
-(h|h')&=&\displaystyle[\alpha_0\varphi(x_0)+\sum_{i=1}^n\alpha_i\varphi(x_i)
-|\alpha_0'\varphi(x_0')+\sum_{i=1}^n\alpha_i'\varphi(x_i)]\\
-&=&\displaystyle
-\alpha_0\alpha_0'\displaystyle k(x_0,x_0')+ \alpha_0\sum_{i=1}^n\alpha_i' k(x_0,x_i)
-+\alpha_0'\sum_{i=1}^n
-\alpha_ik(x_0',x_i)+\hat{\alpha}^T\hat{K}\hat{\alpha}'\\
-&=&\displaystyle\alpha_0\alpha_0'\displaystyle k(x_0,x_0')+ \alpha_0k_0^T\hat{\alpha}'
-+\alpha_0'k_0'^T\hat{\alpha}
-+\hat{\alpha}^T\hat{K}\hat{\alpha}'\\
-\end{array}
-\label{eq:lcl}
-\end{equation}
-where $\hat{\alpha}$ (resp. $\hat{\alpha}'$) denotes the vector of
-$\RR^n$ formed by excluding the first (or rather zeroth) element
-of $\alpha$ (resp. $\alpha'$).
-
-If we write $\beta = [\beta_0; \hat{\beta}]$ similarly to
-$\alpha=[\alpha_0; \hat{\alpha}]$ and keeping in mind that
-$P=[\frac{1}{u}\ 0; -\frac{1}{u}\hat{P}v \ \hat{P}]$, where $\hat{P}$
-is the pseudoinverse of
-$\hat{B}^T$, we have that
-\begin{equation}
-\alpha_0 = \dfrac{\beta_0}{u}, \,\, \hat{\alpha}=-\dfrac{\beta_0}{u}Pv+P\hat{\beta}.
-\end{equation}
-Remembering that $\hat{K}=\hat{B}\hat{B}^T$ and and $k_0=\hat{B}v$
-allows us to rewrite (\ref{eq:lcl}) as
-\begin{align}
-(h|h')& = \dfrac{\beta_0\beta_0'}{uu'}k(x_0,x_0')
-+\dfrac{\beta_0}{u}v^T\hat{B}^T\left(-\dfrac{\beta_0'}{u'}Pv'+P\hat{\beta'}\right)
-+\dfrac{\beta_0'}{u'}v'^T\hat{B}^T\left(-\dfrac{\beta_0}{u}Pv+P\hat{\beta}\right)\\ &+\left(\hat{B}^T\left(-\dfrac{\beta_0}{u}Pv+P\hat{\beta}\right)\right)^T
-\left(\hat{B}^T\left(-\dfrac{\beta_0'}{u'}Pv'+P\hat{\beta'}\right)\right)\\ &=
-\dfrac{\beta_0\beta_0'}{uu'}k(x_0,x_0')
-+\dfrac{\beta_0}{u}v^T\left(-\dfrac{\beta_0'}{u'}v'+\hat{\beta'}\right)
-+\dfrac{\beta_0'}{u'}v'^T\left(-\dfrac{\beta_0}{u}v+\hat{\beta}\right)
-+
-\left(-\dfrac{\beta_0}{u}v+\hat{\beta}\right)^T\left(-\dfrac{\beta_0'}{u'}v'+\hat{\beta'}\right)\\ &=
-\hat{\beta}^T\hat{\beta'}
-+\dfrac{\beta_0\beta_0'}{uu'}\left(k(x_0,x_0')-v^Tv'\right).
-\end{align}
-We also note that $(h|h)=||\beta||^2$ and $(h'|h')=||\beta'||^2$.
-
-Furthermore, from the solution of the linear system (\ref{eq:final}), we find that $\beta_0 = \frac{cu}{\lambda}$, $\hat{\beta} = cG^{-1}(v-\mu)$ and $\beta_0' = \frac{cu'}{\lambda}$, $\hat{\beta}' = c'G^{-1}(v'-\mu)$, for $c,\ c'$ scalars and $G = \frac{1}{n}\hat{B}^T\hat{B}-\mu\mu^T+\lambda\text{Id}$.
-The final similarity score can be computed by
-\begin{equation}
-s(x_0,x_0') = \dfrac{(v-\mu)^TG^{-2}(v'-\mu)+\dfrac{1}{\lambda^2}(k(x_0,x_0')-v^Tv')}{\sqrt{(v-\mu)^TG^{-2}(v-\mu)+\dfrac{u^2}{\lambda^2}}\sqrt{(v'-\mu)^TG^{-2}(v'-\mu)+\dfrac{u'^2}{\lambda^2}}}
-\end{equation}
-
-\section{Notes: LDA and square-loss linear classifiers}
-Let us now go back to linear classifiers, and suppose that we have
-multiple positive samples. It is easy to show that in this case, the
-corresponding linear classifier for the square loss is also given by
-Eq.~(\ref{eq:final}), where $x_0$ denotes this time the center of mass
-of the positive samples, {\em if} these samples have the {\em same}
-covariance matrix $\Sigma$ as the negative ones.
-
-This equal-covariance assumption is of course quite restrictive, and
-probably unrealistic in general. It is interesting to note, however,
-that this is exactly the assumption made by linear discriminant
-analysis. As shown in~\cite{Hastie2009} for example, LDA can be seen as a
-(non-regularized) linear classifier
-with decision function $w'\cdot z+ b'$, where $z$ is a sample in
-$\RR^p$, and
-\begin{equation}
-\left\{\begin{array}{l}
-\displaystyle w'=\Sigma^{-1}(x_0-\mu),\\
-\displaystyle b'=-\frac{1}{2}(x_0+\mu)\cdot w',
-\end{array}\right.
-\label{eq:lda}
-\end{equation}
-
-Note that when $\lambda=0$ (no regularization),
-\begin{equation}
-\Sigma w = Aw -\frac{1}{2}[(x_0-\mu)\cdot w] (x_0-\mu) =
-[1-\frac{1}{2}(x_0-\mu)\cdot w](x_0-\mu)
-=[1-\frac{1}{2}(x_0-\mu)\cdot w]\Sigma w'
-\end{equation}
-thus $w$ and $w'$ have the same direction. In other words, a
-square-loss linear classifier is strictly equivalent to LDA when the
-regularizer parameter $\lambda $ is zero.
-
-More generally, as shown in~\cite{Hastie2009},
-\begin{enumerate}
-\item LDA is a nearest-centroid binary classifier using the
-Mahalanobis distance under the assumption that the two classes
-have the same covariance.
-\item Its decision function is a Bayes classification rule for normal
-distributions with different means but identical covariance matrices
-when, in addition, the priors over the two classes are assumed to be
-equal (the log-ratio of the prior probabilities must be added to the
-offset otherwise). \footnote{In general, Hastie
-{\em et al.}~recommend taking the priors $\pi_i$ equal to the proportion of points
-in class $i$ among all points, that is the class with most samples is
-most probable.}
-\end{enumerate}
-
-Once again, these facts are well known, but may be relevant (or not)
-given recent renewed interest in LDA in modern classification
-tasks~\cite{GMPD12,HMR12}, where some of these properties have been
-rediscovered. In particular, as noted earlier, (non-regularized)
-linear square-loss exemplar machines are proposed in~\cite{ARS14} as a
-means to treat region matching as a classification problem. Aubry {\em
-et al.} note in~\cite{ARS14} that the decision hyperplanes obtained by
-this method and LDA are parallel (our analysis shows they are actually
-identical --- there must be a mistake somewhere). They further show
-that LDA scores calibrated using an additive constant give much better
-results in their matching task. This is a bit puzzling since they
-should give exactly the same results (same rankings and same scores up
-to an affine transformation---before squaring the error term) for a
-given positive example. Aubry {\em et al.} may mean calibration across
-different positive examples. This needs to be clarified and perhaps
-justified as well.
-
-\section{SLEM versus whitening}
-Due to Woodbury identity, Equation \ref{eq:final} can be rewritten as 
-\begin{equation}
-w = \Sigma^{-1}(x_0-\mu).
-\end{equation}
-In the original E-SVM work, the score of similarity between a query image $x_0$ and a possible result $x_0'$ is given by
-\begin{equation}
-s(x_0,x_0') = x_0'^T\Sigma^{-1}(x_0-\mu). \label{white:score}
-\end{equation}
-
-Under the assumption centered data and positive and negative data having the same covariance matrix (as proposed in the previous section), Equation \ref{white:score} corresponds to the score of whitened data:
-\begin{align}
-\hat{x_0} = \Sigma^{-\frac{1}{2}}(x_0-mu);\\
-s(x_0,x_0') = \hat{x_0}^T\hat{x_0'}.
-\end{align}
-This observation has already been made by \cite{Koba15}. However, the score of similarity used in \cite{ZePe15} is
-\begin{equation}
-s(x_0,x_0') = (x_0'-\mu)^T\Sigma^{-2}(x_0-\mu). \label{esvm:score}
-\end{equation}
-
-Based on this, we propose the score $s_{\alpha}$ given by $s_{\alpha}(x_0,x_0') = (x_0-\mu)^T\Sigma^{\alpha}(x_0-\mu).$ $s_{-2}$ is the SLEM similarity score, $s_{-1}$ is the whitening similarity score and $s_{0}$ is the centered base features similarity score.
-
-Strangely, both scores seem to perform similarly, even if their scores differ. Doing a singular value decomposition of $\Sigma$, $\Sigma^{-1}=VD^{-1}V^T$ and $\Sigma^{-\alpha}=VD^{-\alpha}V^T$.
-
-\begin{figure}[!h]
-\centering
-\begin{subfigure}[b]{0.45\textwidth}
-\includegraphics[width=\textwidth]{whitening_cnn.jpg}
-\end{subfigure}
-\begin{subfigure}[b]{0.45\textwidth}
-\includegraphics[width=\textwidth]{whitening_vlad.jpg}
-\end{subfigure}
-\caption{Comparison of mAP for different similarity scores $s_{\alpha}$, varying $\alpha$.
-At left, CNN features.
-At right, VLAD features.
-On blue, $\Sigma$ is the covariance of the negative samples and on red, $\Sigma$ is the covariance of the positive and negative samples.
-On black, the mAP for similarity calculated as inner product of base features.}
-\label{Sigma}
-\end{figure}
-
-\section{Experiment Evaluation}
-\subsection{Base image encoding}
-We use three base features as the representation in $\RR^p$ of our images. The two firsts are VLAD and CNN features, as used in \cite{ZePe15}. 
-CNN features are non-negative and can be used both as $\mathbb{L}^1$ or $\mathbb{L}^2$ normalized features.
-The third features are pyramids of histograms of SIFT descriptors, as used in \cite{spk}.
-These are non-negative $\mathbb{L}^1$ normalized features.
-
-\subsection{Kernels}
-For all features, we test non-kernelized SLEM, as well as kernelized SLEM for gaussian (rbf) and polynomial kernel:
-\begin{align}
-k_{rbf}(x,y) &= \exp\left(-\gamma||x-y||^2\right);\\
-k_{poly}(x,y) &= x^Ty+\gamma(x^Ty)^2.
-\end{align}
-For non-negative $\mathbb{L}^1$ normalized, we also test the intersection kernel: $k_{inter}(x,y) = \min(x,y).$\\
-We observe that the non-kernelized obtains similar results to kernels polynomial and Gaussian. For the pyramid histogram of SIFT, the intersection kernel works much better than all the other kernels.
-
-\begin{table}
-\begin{center}
-\begin{tabular}{|l|c|c|c|c|}
-\hline
-Method & $s_0$ & $s_{-2}$ with no kernel & $s_{-2}$ for gaussian kernel &  $s_{-2}$ for intersection kernel \\
-\hline\hline
-VLAD-64 & 73.3 & 77.9 & 78.1 & -\\
-Caffe CNN & 69.7 & 72.1 & 72.9 & 70.2\\
-Pyramid of Histogram & 38.7 & 43 & 44.6 & 63.5\\
-\hline
-\end{tabular}
-\end{center}
-\caption{Results.}
-\end{table}
-
-\bibliographystyle{plain}
-\bibliography{sup}
 \end{document}

### 2016-11-07 jean_original.tex
first commit

--- 2016-11-07 esvm.tex
@@
 
-\section{SLEM versus whitening}
-Due to Woodbury identity, Equation \ref{eq:final} can be rewritten as 
-\begin{equation}
-w = \Sigma^{-1}(x_0-\mu).
-\end{equation}
-In the original E-SVM work, the score of similarity between a query image $x_0$ and a possible result $x_0'$ is given by
-\begin{equation}
-s(x_0,x_0') = x_0'^T\Sigma^{-1}(x_0-\mu). \label{white:score}
-\end{equation}
-
-Under the assumption centered data and positive and negative data having the same covariance matrix (as proposed in the previous section), Equation \ref{white:score} corresponds to the score of whitened data:
-\begin{align}
-\hat{x_0} = \Sigma^{-\frac{1}{2}}(x_0-mu);\\
-s(x_0,x_0') = \hat{x_0}^T\hat{x_0'}.
-\end{align}
-This observation has already been made by \cite{Koba15}. However, the score of similarity used in \cite{ZePe15} is
-\begin{equation}
-s(x_0,x_0') = (x_0'-\mu)^T\Sigma^{-2}(x_0-\mu). \label{esvm:score}
-\end{equation}
-
-Based on this, we propose the score $s_{\alpha}$ given by $s_{\alpha}(x_0,x_0') = (x_0-\mu)^T\Sigma^{\alpha}(x_0-\mu).$ $s_{-2}$ is the SLEM similarity score, $s_{-1}$ is the whitening similarity score and $s_{0}$ is the centered base features similarity score.
-
-Strangely, both scores seem to perform similarly, even if their scores differ. Doing a singular value decomposition of $\Sigma$, $\Sigma^{-1}=VD^{-1}V^T$ and $\Sigma^{-\alpha}=VD^{-\alpha}V^T$.
-
-\begin{figure}[!h]
-\centering
-\begin{subfigure}[b]{0.45\textwidth}
-\includegraphics[width=\textwidth]{whitening_cnn.jpg}
-\end{subfigure}
-\begin{subfigure}[b]{0.45\textwidth}
-\includegraphics[width=\textwidth]{whitening_vlad.jpg}
-\end{subfigure}
-\caption{Comparison of mAP for different similarity scores $s_{\alpha}$, varying $\alpha$.
-At left, CNN features.
-At right, VLAD features.
-On blue, $\Sigma$ is the covariance of the negative samples and on red, $\Sigma$ is the covariance of the positive and negative samples.
-On black, the mAP for similarity calculated as inner product of base features.}
-\label{Sigma}
-\end{figure}
-
-\section{Experiment Evaluation}
-\subsection{Base image encoding}
-We use three base features as the representation in $\RR^p$ of our images. The two firsts are VLAD and CNN features, as used in \cite{ZePe15}. 
-CNN features are non-negative and can be used both as $\mathbb{L}^1$ or $\mathbb{L}^2$ normalized features.
-The third features are pyramids of histograms of SIFT descriptors, as used in \cite{spk}.
-These are non-negative $\mathbb{L}^1$ normalized features.
-
-\subsection{Kernels}
-For all features, we test non-kernelized SLEM, as well as kernelized SLEM for gaussian (rbf) and polynomial kernel:
-\begin{align}
-k_{rbf}(x,y) &= \exp\left(-\gamma||x-y||^2\right);\\
-k_{poly}(x,y) &= x^Ty+\gamma(x^Ty)^2.
-\end{align}
-For non-negative $\mathbb{L}^1$ normalized, we also test the intersection kernel: $k_{inter}(x,y) = \min(x,y).$\\
-We observe that the non-kernelized obtains similar results to kernels polynomial and Gaussian. For the pyramid histogram of SIFT, the intersection kernel works much better than all the other kernels.
-
-\begin{table}
-\begin{center}
-\begin{tabular}{|l|c|c|c|c|}
-\hline
-Method & $s_0$ & $s_{-2}$ with no kernel & $s_{-2}$ for gaussian kernel &  $s_{-2}$ for intersection kernel \\
-\hline\hline
-VLAD-64 & 73.3 & 77.9 & 78.1 & -\\
-Caffe CNN & 69.7 & 72.1 & 72.9 & 70.2\\
-Pyramid of Histogram & 38.7 & 43 & 44.6 & 63.5\\
-\hline
-\end{tabular}
-\end{center}
-\caption{Results.}
-\end{table}
-
 \bibliographystyle{plain}

### 2016-11-07 main_cvpr_review.tex
first commit

\documentclass[10pt,twocolumn,letterpaper]{article}

\usepackage{cvpr}
\usepackage{times}
\usepackage{epsfig}
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{amssymb}

\usepackage{graphicx}
\usepackage{caption}
\usepackage{subfigure}
\usepackage{subcaption}
\captionsetup{compatibility=false}
\usepackage{amssymb}
\usepackage{amsthm}
%\usepackage{ruler}
\usepackage{color}
\usepackage{booktabs}
%\usepackage[width=122mm,left=12mm,paperwidth=146mm,height=193mm,top=12mm,paperheight=217mm]{geometry}
\newtheorem{proposition}{Proposition}
\newtheorem{lemma}{Lemma}
%\newtheorem{proof}{Proof}
\newcommand{\RR}{\mathbb R}
\newcommand{\NNN}{\mathcal N}
\newcommand{\sumi}{\displaystyle{\sum_{i=1}^n}}
\DeclareMathOperator*{\argmin}{argmin}

% Highlighting
\usepackage{soul}
%\usepackage[usenames,dvipsnames]{xcolor}
\newcommand{\hlc}[2][yellow]{{\sethlcolor{#1}\hl{#2}}}
\newcommand{\RAF}[1]{\hlc[yellow]{(RR:) #1}}
\newcommand{\JZ}[1]{\hlc[pink]{(JZ:) #1}}
\newcommand{\PP}[1]{\hlc[green]{PP: #1}}
\newcommand{\un}[1]{\underline{#1}}

\usepackage{tikz}
\usepackage{pgfplots,pgfplotstable}
\usetikzlibrary{pgfplots.groupplots}
\pgfplotsset{grid=major,height=2in, width=\columnwidth}
\input{tikz_styles.tex}

% Include other packages here, before hyperref.

% If you comment hyperref and then uncomment it, you should delete
% egpaper.aux before re-running latex.  (Or just hit 'q' on the first latex
% run, let it finish, and you should be clear).
\usepackage[pagebackref=true,breaklinks=true,letterpaper=true,colorlinks,bookmarks=false]{hyperref}

% \cvprfinalcopy % *** Uncomment this line for the final submission

\def\cvprPaperID{882} % *** Enter the CVPR Paper ID here
\def\httilde{\mbox{\tt\raisebox{-.5ex}{\symbol{126}}}}

% Pages are numbered in submission mode, and unnumbered in camera-ready
\ifcvprfinal\pagestyle{empty}\fi
\begin{document}

%%%%%%%%% TITLE
\title{Kernel Square-Loss Exemplar Machines}

\author{Rafael S. de Rezende\\
INRIA Paris\\
Institution1 address\\
{\tt\small rafael.sampaio_de_rezende@inria.fr}
% For a paper whose authors are all at the same institution,
% omit the following lines up until the closing ``}''.
% Additional authors and addresses can be added with ``\and'',
% just like the second author.
% To save space, use either the email address or home page, not both
\and
Second Author\\
Institution2\\
First line of institution2 address\\
{\tt\small secondauthor@i2.org}
}

\maketitle
%\thispagestyle{empty}
%%%%%%%%% ABSTRACT
\begin{abstract}
This paper proposes a new feature encoding pipeline, based on the success of the exemplar SVM (ESVM) \cite{ZePe15} and linear discriminant analysis (LDA) as whitening \cite{HMR12}.
We first show that, by replacing the hinge loss by the square loss in the ESVM cost function, similar results in image retrieval can be obtained at a fraction of the computational cost. 
We call this model square-loss exemplar machine, or SLEM. 
Secondly, we introduce a kernelized SLEM version which benefits from the same computational advantages but may displays improved performance. 
Both SLEM variants exploit the fact that the negative examples are fixed for all the positives in the training set, so most of the SLEM computational complexity is relegated to an offline process independent of the positive examples. 
Our experiments establish the performance and computational advantages of our methods using a large array of base feature representations and standard image retrieval datasets.
\end{abstract}

\input{sec_intro.tex}

%\input{sec_prior.tex}

\input{sec_SLEM.tex}

\input{sec_kernel_methods.tex}

\input{sec_eff_imp.tex}

\input{sec_eval_raf.tex}

%\input{comp_sota.tex}
\input{conclusion.tex}

%% RR: uncomment for final version
%% \subsection*{Acknowledgments} This work was supported by the ERC grant VideoWorld.

{\small
\bibliographystyle{ieee} 
\bibliography{sup,jz}
}
\end{document}

### 2016-11-07 sec_SLEM.tex
first commit

\section{Square loss exemplar machine}\label{lsesvm}
In this section, we revisit the exemplar SVM model presented by \cite{Efros11} as a more general exemplar classifier model. Then, we introduce the square loss exemplar machine (SLEM) by solving a simple variant this model and study its properties.
\subsection{Exemplar classifiers} \label{esvm}
We are given base features in $\RR^d$ at training time, one positive example $x_0$ in $\RR^d$ and a set of negative examples $X = [x_1, x_2,...,x_n]$ in $\RR^{d\times n}$, each column of $X$ representing one example by a vector in $\RR^d$. 
We are also given a loss function $l:\{-1, 1\}\times \RR\rightarrow\RR^+$. Learning an exemplar classifier from these examples amounts to minimizing the function 
\begin{equation}
J(\omega, \nu) = \theta \ l(1, \omega^Tx_0+\nu) +\dfrac{1}{n}\sumi l(-1, \omega^T x_i+\nu)+\dfrac{\lambda}{2}\|\omega\|^2, 
\label{eq:first}
\end{equation}
w.r.t. $\omega$ in $\RR^d$ and $\nu$ in $\RR$. In Equation (\ref{eq:first}), $\lambda$ and $\theta$ are respectively a regularization parameter on $\omega$ and a positive scalar adjusting the weight of the positive exemplar.
%\footnote{We could also acknowledge a parameter $\theta$ other than $\frac{1}{n}$ as a regularization to the error of each negative example. But this parameters seems to be less important to cross-validate. Also, setting $\theta=\frac{1}{n}$ simplify Equation (\ref{omega:solution}) and allows the use of Woodbury identity.}

The exemplar classifiers of $x_0$ with respect to $X$ are the weights $\omega^\star(x_0,X)$ that minimizes the loss function $J$:
\begin{equation}
\big(\omega^\star(x_0, X), \nu^\star(x_0, X)\big) = \underset{(\omega,\nu)\in\RR^d\times\RR}{\argmin} \ J(\omega, \nu). \label{omega:first}
\end{equation}
To shorten the notation, we refer to these solutions as $\omega^\star$ and $\nu^\star$\footnote{Depending on the loss function $l$, $\nu^\star(x_0,X)$ may be not unique.} when their arguments are implicit.

%\cite{Efros11} introduced ensembles of exemplar classifiers for object detection, but single exemplar classifiers have been used as replacement of the original feature $x_0$ for image classification and retrieval \cite{ZePe15} and detection \cite{GMPD12}. \RAF{Maybe this should be in the introduction instead.}
Applications of ESVM use the hinge loss function, which guarantees that $J$ is a convex function. The solution of Equation (\ref{omega:first}) can be thus found by stochastic gradient descent~\cite{bottou10} individually for each positive sample. The next subsection show how to calculate all exemplar classifiers simultaneously by changing the loss function.

\subsection{Square loss function}\label{slem:intro}
Now, let us study the same learning problem for the square-loss function $l(y,\hat{y}) = \frac{1}{2}(y-\hat{y})^2$. As in the case of the hinge loss, Equation (\ref{eq:first}) is a convex problem. 
However it is now a ridge regression problem, whose unique solution can be found in closed form as
\begin{align}
\begin{cases}
\vspace{3 mm}
\omega^\star &= \dfrac{2\theta}{\theta+1}U^{-1}(x_0-\mu), \\
%\vspace{3 mm}
\nu^\star &= \dfrac{\theta-1}{\theta+1}-\dfrac{1}{\theta+1}(\theta x_0+\mu)^T\omega^\star,
\end{cases}
\label{omega:solution}
\end{align}
%\begin{align}
%\nu^\star &= \dfrac{\theta-1}{\theta+1}-\dfrac{1}{\theta+1}(\theta x_0+\mu)^T\omega^\star, \label{eq:nustar}\\
%\omega^\star &= \dfrac{2\theta}{\theta+1}U^{-1}(x_0-\mu) \label{eq:wstar}
%\end{align}
where:
\begin{align}
\begin{cases}
\mu = &\frac{1}{n}\sum_{i=1}^n x_i,\\ \\
U = &\frac{1}{n}XX^T-\mu\mu^T\\
&+\frac{\theta}{\theta+1}(x_0-\mu)(x_0-\mu)^T+\lambda\mathrm{Id}_d. \label{eq:U}
\end{cases}
\end{align}
Equation (\ref{omega:solution}) shows how to solve (\ref{eq:first}) simultaneously for all exemplars.
%Replacing the hinge loss by the square loss offers a more compact solution, but not necessarily more efficient.

\textbf{Woodbury identity.} 
We can modify Equation \ref{omega:solution} by simplifing Equation \ref{eq:U}.
Let us define $A = \frac{1}{n}XX^T-\mu\mu^T +\lambda\mathrm{Id}_d$ our regularized covariance matrix and assume its inverse $A^{-1}$ known. 
%$A$ is the covariance matrix of the negative samples $X$\footnote{added of a small term proportional to the identity matrix to insure $A$ is positive-definite.}. Let us assume its inverse $A^{-1}$ is known. 
The matrix $U$ now reads $U = A + \frac{\theta}{\theta+1}\delta\delta^T$, where $\delta=x_0-\mu$ is the centralized (w.r.t. the negative mean) positive sample. The Woodbury identity \cite{woodbury} gives us
\begin{equation}
U^{-1} = A^{-1} -\dfrac{\theta}{\theta\delta^TA^{-1}\delta+ \theta+1}A^{-1}\delta^T\delta A^{-1}. \label{invU}
\end{equation}
Substituting (\ref{invU}) in (\ref{omega:solution}) yields
\begin{equation}
\begin{split}
\omega^\star &= \dfrac{2\theta}{\theta +1}\left(A^{-1}\delta - \dfrac{\theta}{\theta\delta^TA^{-1}\delta+ \theta+1} A^{-1}\delta (\delta^TA^{-1}\delta)\right)\\
&= \dfrac{2\theta}{\theta\delta^TA^{-1}\delta+ \theta+1} A^{-1}\delta.\label{Wood:omega}
\end{split}
\end{equation}

Note that the positive sample weight $\theta$ does not influence the direction of the optimal vector $\omega^\star$, only its norm. This means that if search and ranking are based on the normalized feature $\frac{1}{\|\omega^\star\|}\omega^\star$, $\theta$ does not influence the matching score of the SLEM vectors of two different images. This sets SLEM appart from ESVM that requires this parameter to be calibrated \cite{Efros11,ZePe15}. The normalized exemplar classifier $\omega^\star$ is therefore a linear transformation of $x_0-\mu$
%Indeed, if $\omega$ and $\omega'$ are the $d$-dimensional SLEM vectors of $x$ and $x'$, respectively, their cosine similarity reads:
%we can denote $s$ the matching score scalar function defined in $\RR^d\times \RR^d$, which is given by
%\begin{equation}
%s(\omega, \omega') = \dfrac{\omega^T \omega'}{\|\omega\| \|\omega'\|} = \frac{(x-\mu)^TA^{-2}(x'-\mu)}{\|A^{-1}(x-\mu)\|\|A^{-1}(x'-\mu)\|},\label{match:score}
%\end{equation}
%which does not depend on the value of $\theta$. 
%This means that the weight of the positive sample can be fixed at any positive value without changing matching scores.

\subsection{LDA and SLEM}\label{sec:lda}

Let us now reanalyse Equation \ref{eq:first} for a particular case. Suppose that we have multiple positive samples. It can be shown that in this case, the corresponding linear classifier of Equation (\ref{eq:first}) for the square-loss is also given by
(\ref{omega:solution}), where $x_0$ denotes this time the center of mass
of the positive samples, {\em if} these samples have the {\em same} covariance matrix $\Sigma$ as the negative samples $X$.

This equal-covariance assumption is of course quite restrictive, and probably unrealistic in general. It is interesting to note, however, that this is exactly the assumption made by linear discriminant analysis. As shown in~\cite{Hastie2009} for example, LDA can be seen as a (non-regularized) linear classifier with decision function $\omega^T_{LDA} z+ \nu_{LDA}$, where $z$ is a sample in
$\RR^d$, and
\begin{equation}
\left\{\begin{array}{l}
\displaystyle \omega_{LDA}=\Sigma^{-1}(x_0-\mu),\\
\displaystyle \nu_{LDA}=-\frac{1}{2}(x_0+\mu)^T \omega_{LDA},
\end{array}\right.
\label{eq:lda}
\end{equation}

Note that SLEM is a generalized version of LDA. Indeed, when $\lambda=0$ in SLEM (i.e. no regularization), $A = \Sigma$, meaning $\omega_{LDA}$ of Eq. (\ref{eq:lda}) and $\omega^\star$ of Eq. (\ref{Wood:omega}) have the same direction.
%This observation has been equally made by \cite{Koba15}.

Many interesting properties of LDA has been used recently for classification tasks \cite{GMPD12,HMR12} and, more recently, for image retrieval with convolutional neural network activations \cite{babenko15}.
By simply generalizing LDA, we hope to obtain superior results.

%In the next section, we take one step further and generalize SLEM with kernel methods. \hlc{something to add to this sentence?}

%%% Local Variables:
%%% TeX-master: "main_eccv"
%%% End:

### 2016-11-07 sec_eff_imp.tex
first commit

\section{Efficient implementation}\label{eff_imp}

\input{figure_LR.tex}
When compared to the square-loss classifier of section \ref{slem:intro}, one drawback of a kernelized approach is that the dimension of our problem grows with the size $n$ of the negative samples.
The offline factorization $B$ of $K$ demands $O(nr)$ storage and at best $O(n^2r)$ time. This factorization can be obtained in different ways, depending on the aspect we wish to optimize. Efficiency in time and efficiency in storage are the two most common of these aspects. In this section we propose two different calculations of $B$, to be implemented depending on which of these two efficiencies we hope to achieve.

%As for the online step, solving Equation (\ref{beta:final}) amounts to, for all positive exemplars, calculating the row $[u, v^T]$ of $B'$  from Lemma 1 and solving a linear system in $A$, which is a $(r+1)\times (r+1)$ matrix. The first of these is computed in time $O(nr)$ and the second, $O(r^3)$.

\subsection{Time-efficient implementation}
The Cholesky decomposition is the most used factorization of positive-semidefinite matrices in kernel-based learning \cite{BaJo02,BaJo05,FiSc01}. 
The computation of the factor $B$ depends on a permutation $\pi(n) =\{i_1,i_2,\dots,i_n\}$ of $\{1,2,\dots, n\}$, called \textit{pivots}. The $t$-th column of $B$ is calculated iteratively as:

\begin{align}
\begin{cases}
\vspace{3 mm}
B(i_t, t) = \left(K(i_t,i_t)-\sum_{m=1}^{t-1} B(i_{m},m)  \right)^{\frac{1}{2}},\\
\vspace{3 mm}
B(i_j,t) = 0,\\
B(i_j, t) = \frac{1}{B(i_t,t)}(K(i_j,i_t)  -\sum_{m=1}^{t-1}B(i_j,m)B(i_t,m)), \end{cases}\label{icd:algo}
\end{align}
for all $t<j\le n$. There are two alternatives of Cholesky decomposition. The incomplete Cholesky decomposition (ICD) stops the algorithm (\ref{icd:algo}) after $r$ steps and has complexity $O(nr^2)$. To minimize the residue $\|K-BB^T\|$, the $t$-th pivot is calculated at the end of the $t$-th interaction. In the other hand, the complete Cholesky decomposition (CCD) consists of a full-rank decomposition, \textit{i.e.} the rank $r$ of $B$ is equal to $n$ and it has complexity $O(n^3)$. In this case, the pivots are irrelevant to the final residue (which is zero) and can be assumed to be $i_t=t$. The computation of pivots makes CCD more efficient for similar residues.
We make sure $K$ is positive-definite by adding a $\epsilon$ to its diagonal, where $\epsilon=\min(0,-\lambda_{min})$ and $\lambda_{min}$ is the smallest eigenvalue of $K$. Therefore, from the equation $BB^T=K+\epsilon\mathrm{Id}_n$, $B$ has rank $n$ and we can apply CCD.

\subsection{Storage-efficient implementation}\label{low-rank} %/Low-rank decomposition} 
Let us now consider the problem of improving retrieval results in minimal storage. As discussed in section \ref{simi_score}, for each positive exemplar we store its original encoder plus $2r$ dimensions, so we aim to decompose $K$ at a small rank $r$.
For small values of $r$, a kernel PCA (KPCA) factorization guarantees a better approximation of $K$ than the Cholesky decomposition at a higher time cost. Indeed, as illustrated in Figure \ref{fig:residue}, a kernel PCA gives smaller residues than ICD at small rank. We plot the ratio $tr(K-BB^T)/tr(K)$ when we vary the number of columns $r$ of $B$, and we see a faster convergence for the kernel PCA decomposition when $r$ goes to $n$. However, at these small values of $r$, KPCA is more time consuming than both ICD and CCD. This time cost comes from applying a singular values decomposition of K to obtain $B$:
\begin{equation}
    K = WDW^T; \ b_i = \sqrt{d_{ii}}w_i, \label{svd}
\end{equation}
where $w_i$ denotes the $i$-th column of $W$ and $d_{ii}$ the $i$-th diagonal element of $D$. Decomposing $K$ from Equation~(\ref{svd}) has complexity $O(n^3)$ in time.

As for ICD, it can be the fastest decomposition at small rank, but it underperformed KPCA; and for higher ranks, it is the slowest alternative and its performance approximates CCD's performance. We therefore concentrate our experiments in the two other methods. \hlc{RR: However, ICD is still the only $O(nr^2)$ in time, which could be useful for bigger values of $n$. I'm considering showing the experiments done in the ECCV supmat to show why ICD is obsolete.}

%\begin{figure}
%    \centering
%    \includegraphics[width=0.5\textwidth]{trace_residual.jpg}
%    \caption{Residual error of decompositions varying by its rank.}
%    \label{fig:residue}
%\end{figure}

### 2016-11-07 sec_eval.tex
first commit

\section{Experimental Evaluation}
\label{eval}

\subsection{Datasets and evaluation protocol} \label{eval:protocol}
\subsubsection{Holidays and Oxford5K.} We use two standard image retrieval datasets for our experiments. The first one is the INRIA \emph{Holidays} dataset \cite{holidays}. This dataset consists of 1491 images divided into 500 groups of matching images. %If the dataset of an experiment is not mentioned in what follows this means, by default, that it is performed on \emph{Holidays}.
The second one is  the \emph{Oxford5k} dataset \cite{oxford}, which consists of 5062 images divided into 55 groups of matching images.

In both datasets, each group contains one query image, the other images in the group being the only correct answers to the query. We use the mAP computation procedures provided along with the dataset to evaluate performance.

\subsubsection{Flickr $1e6$.} Besides these two datasets, we also use $1e6$ random distractor images downloaded from \emph{Flickr} to carry out large scale experiments, following standard protocol \cite{holidays,Delhumeau2013,ZePe15}.

%For each query image, we calculate its similarity to all other images in the database and rank them, in decreasing order.
%The average precision of a group is derived from the ranking of the images of the group for the similarity with the corresponding query image.
%The final mean average precision (mAP) for a dataset is the mean of the average precision over all its groups.
\subsubsection{Negative pool.} As a pool of negative images to build E-SVM and SLEM representations, we use the Flickr100k collection \cite{oxford}, composed of $10^5$ random Flickr images disjoint from those in Flickr $1e6$.. For a full rank decomposition, we use only a subset of it, containing between 6000 and 15000 images.
As stated in section \ref{low-rank} and further discussed in section \ref{time-scale}, a full rank decomposition does not scale well for bigger number of negative samples. For low rank decomposition, we use all 100000 images.

At evaluation time, for a dataset that consists of $p$ images and $q$ query images, we calculate its $p\times q$ \emph{similarity matrix} $S$, where each of its $q$ columns is the matching scores of the query image with all the $p$ images.

\subsection{Which kernel to choose?}
For the experiment of SLEM as a feature encoder, we tested three different kernels --linear, Gaussian, polynomial and Spatial Pyramid Matching (SPM) -- defined below:
\begin{align}
    k_1(x,y) = x^Ty; \
    k_2(x,y) = e^{-\gamma\|x-y \|^2}; \ 
    k_3(x,y) = x^Ty+\gamma(x^Ty)^2. \label{kernels}
\end{align}
%\JZ{We give results with spp, we should add it. Makes it clear that we use a wide range of kernels.}
%\JZ{The text below could be removed. But some brief text for SPM is a good idea, as it is less standard. I think SPP and SPM are not the same thing... We should use SPM... SPP is a finite dimensional feature map.}
% \textbf{Linear SLEM} The linear kernel of Equation (\ref{k:lin}) is the first default choice and equivalent to the non-kernelized version of SLEM.
% In the remaining of this paper, we reference to linear SLEM when we use the non-kernelized SLEM.
% \textbf{Gaussian SLEM} The radial basis function kernel of Equation (\ref{k:rbf}) is a
% well known reproducing kernel, used for classification with support vector machines.
% \textbf{Polynomial SLEM} The polynomial kernel of Equation (\ref{k:poly}) is a reproducing kernel often used in natural language processing.
%\textbf{SPM SLEM} The spatial pyramid matching kernel of $L+1$ levels in Equation \ref{k:spm} take as input a set of local descriptors and its location in pyramidal bins \cite{spk}. 
Note that Gaussian and polynomial kernels require a scalar parameter ($\gamma$ and $\lambda$, respectively), that we set using cross-validation \JZ{What is the cross-valid dataset?}.
The SPM kernel of $L+1$ levels in Equation \eqref{k:spm} takes as input a set of local descriptors and computes a similarity score that takes into account their geometrical layout \cite{spk}.\JZ{Need to define $H_x^{l,\gamma}$ ...}
%its location in pyramidal bins \cite{spk}.

\subsection{Base visual features}
We test our feature encoder for three different base features $x\in\mathbb{R}^d$ the hand-crafted VLAD image representation \cite{Delhumeau2013} and two learned features derived from the activation coefficients of deep Convolutional Neural Networks \cite{Krizhevsky2012,babenko15}.

We use the same VLAD variant of \cite{Delhumeau2013} used in \cite{ZePe15} that relies on densely-extracted rootSIFT \cite{3things} local descriptors, per-cluster, PCA-based rotations, and root normalization. Like \cite{ZePe15}, we use $64$ clusters, for a final feature size of $8192$.

The first CNN-based feature we use consists of the activation coefficients of the previous-to-last layer of the architecture of \cite{Krizhevsky2012}, based on a publicly available pre-trained model \cite{jia2014caffe}. This is the same feature considered in \cite{ZePe15} and originally considered as an image retrieval feature in \cite{Sharif}. We refer to it as FC-CNN. \JZ{Please verify...}

The second CNN-based feature, the SPoC feature of \cite{babenko15}, is tailored specifically for the image retrieval application. It consists of a spatially-weighted sum-pooling of the activations of the last convolutional layer of a 19-layer CNN \cite{Simonyan2014}.

%REMOVED
% We revisit the VLAD feature presented in \cite{VLAD} as an example of a hand-crafted representation. First we extract a set $\mathcal{F}$ of local descriptors of the image $I$. We use the 128 dimension RootSIFT \cite{3things} descriptors, extracted densely.
% Then, we hard-assign each descriptor $f$ in $\mathcal{F}$ to the closest among $K$ pre-trained codewords $c_k$, $k\in\{1\cdots K\}$,
% %one of $K$ set $\mathcal{C}_k$ of descriptors associated to codewords $\{c_k\}_{1\leq k\leq K}$, 
% and map $f$ to a $\RR^{128K}$ vector
% \begin{equation}
% \phi^{VL}_1(f) = \left[0 \cdots 0\quad \Phi_k^T\frac{(f-c_k)}{\|f-c_k\|} \quad 0 \cdots 0\right],
% \end{equation}
% where $\Phi_k$ is a $128\times 128$ PCA matrix learned on training features mapped to $k$-the codeword.
% %associated to descriptors in $\mathcal{C}_k$. 
% The final VLAD representation is the power-normalization and $l_2$ normalization of the sum-pooling of $\phi^{VL}_1$:
% \begin{equation}
% \phi^{VL}_2(I) \propto \mathrm{power}\big(\sum_{f\in \mathcal{F}}\phi^{VL}_1(f)\big),~\|\phi^{VL}_2(I)\|=1,
% \end{equation}
% with scalar power normalization $\mathrm{power}(v)=\mathrm{sign}(v)|v|^{0.5}$ applied component-wise.
% In experiments, we use $K=64$ codewords learned on images from Flickr. Our VLAD representation has $d=8192$ dimensions.

%REMOVED
% Convolutional features obtained from very deep convolutional neural networks (CNNs) have been shown to work as good local descriptors for matching \cite{SimonZisser15}. The SPoC representation \cite{babenko15} is a weighted sum-pooling of the activations of the last convolutional layer of a 19-layer CNN. For a given input image $I$, the activations in this layer are organized over a $W\times H$ spatial grid and over $d$ channels. Each position $(w,h)\in \{1 \cdots W\}\times \{1\cdots H\}$ can thus be equipped with a descriptor $f_{h,w}(I)\in\RR^D$.
% %Indeed, if our last convolutional layer has $D$ neurons, and each neuron a $W\times H$ map of activations of this neurons to the image $I$, 
% %each pair $(w,h)$ with $w$ in $\{1,2,..., W\}$ and $h$ in $\{1,2,...,H\}$ can be associated to a descriptor $f_{(h,w)}$ in $\RR^D$ of the responses of each neuron at coordinate $(w,h)$ of the maps. 
% We then sum-pool these descriptors, weighted accordingly to their distance to the center of the image:
% \begin{equation}
%     \phi^{SPoC}(I) = \sum_{w=1}^W\sum_{h=1}^H \alpha_{w,h}f_{w,h}(I),
% \end{equation}
% where
% \begin{equation}
%     \alpha_{w,h} = \exp \left(-\dfrac{(w-W/2)^2+(h-H/2)^2}{2\sigma^2}\right).
% \end{equation}
% In our experiments, we follow the implementation details of \cite{babenko15}: We resize all images to $586\times 586$ pixels before feeding them to the network. The last convolutional layer has $d=512$ channels with activation maps of size $(W,H)=(37,37)$. We also set $\sigma=\frac{H}{3}$.

% Finally, we also test a more convencional CNN feature, less deep and using two fully connected layers after the convolutional layers.
% Our final representation is a $4096$ non-negative feature. We based out implementation on CAFFE \cite{jia2014caffe}.

% \subsection{Implementation details}
% We cross validate $\gamma$ and $\lambda$ for Poly and Gaussian SLEM.

\subsection{Full rank results}
In Table \ref{fullrank:results}, we test our method using tye three recent base features discussed above (VLAD \cite{Delhumeau2013}, SPoC \cite{babenko15}, and FC-CNN \cite{Sharif}). We further compare against the ESVM method of \cite{ZePe15}, as well as the PCA+whitening strategy shown in \cite{babenko15} to drastically improve the retrieval results for SPoC. In our experiments, using PCA dimensionality reduction worsens the results, and hence we use the full PCA projection matrix for the PCA+whitening strategy.
%Hence we compare the improvements of PCA plus whitening, without compression, with our method.

Linear SLEM performs similarly to ESVM despite being much more time efficient (see discussion in section \ref{time-scale}).
Gaussian SLEM and Polynomial SLEM outperform all methods, for all datasets and all image representations.
The results are presented in Table \ref{fullrank:results}.

\begin{table*}[t]
\begin{center}
\caption{Mean average precision results for INRIA Holidays and Oxford buildings datasets, expressed as percentages. In this table, we present our results for VLAD-64 \ref{VLAD}, sum-pooling of convolutional features (SPoC) \ref{babenko15} and fully connected (fc) CNN \ref{jia2014caffe}}
\begin{tabular}{|c|c|c|c|c|c|c|c|}
\hline
Dataset & \multicolumn{3}{|c|}{\textbf{Holidays}} & \multicolumn{3}{|c|}{\textbf{Oxford 5k}}\\
\hline
Method, features & VLAD-64  & SPoC CNN & FC CNN & VLAD-64 & SPoC CNN & FC CNN\\
\hline\hline
Baseline            & 72.7         & 73.1         & 68.2         & 46.3           & 54.4         & 40.6\\
%Whitening           & -    & -    & -    & -    & -\\
PCA+whitening       & 69.4         & 77.5         & 69.2         & 50.9           & 63.7         & 45 \\
E-SVM               & \textbf{78.3} & 79.9         & 71.8         & 57.5           & 62.1         & 44.6\\
Linear SLEM         & 78           & 78.3         & 72.1         & \textbf{59.3}   & 64.1         & 45.5\\
Gaussian SLEM       & 78.1         & 81.4         & \textbf{72.9} & 59             & \textbf{64.9} & 46.1\\
Poly SLEM           & 78.1         & \textbf{82}   & \textbf{72.9} & \textbf{59.3}  & 64.8         & 46\\
\hline
\end{tabular}
\end{center}
\label{fullrank:results}
\end{table*}

\input{figuredistrac.tex}
\subsection{Time Scalability} \label{time-scale}
In this section we compare the time efficiency of our method and the E-SVM, as well as discuss which method to use accordingly with the number of positive and negative samples.

In Figure \ref{fullrank:results}, we see Linear SLEM efficiency does not change with $n$.
Indeed, if $d$ is the dimension of the base representation, $A$ is a $d\times d$ matrix for Linear SLEM, whereas for a full rank kernel, $A$ is a $n\times n$.
This explain the increasing running time for Gaussian and polynomial kernels: storage and solving a $n\times n$ system does not scale for large number of negative samples.
But retrieval results in Figure \ref{fullrank:results} suggest we can benefit from larger sets of negative samples. Linear SLEM and ESVM scales well, but do not perform so well.

\vspace{3 mm}

\input{figuren.tex}

\subsection{Low-rank decomposition evaluation}

%\begin{figure}[!h]
%\centering
%\begin{subfigure}[b]{0.48\textwidth}
%\includegraphics[width=\textwidth]{linear_decomposition_nolog.png}
%\end{subfigure}
%\begin{subfigure}[b]{0.48\textwidth}
%\includegraphics[width=\textwidth]{rbf_decomposition_nolog.png}
%\end{subfigure}
%\caption{Comparison between full rank and low rank. In blue, mAP results for full rank SLEM. %In red, mAP results for low rank decomposition of SLEM, varying the rank $r'$.
%At the left, linear SLEM results;at the right, Gaussian SLEM results. In this experiment, $n=10281$.}
%\label{no.ker.vs.linear2}
%\end{figure}

\subsection{Spatial pyramid matching kernel}

\begin{table}[!h]
    \centering
    \caption{Results for spatial pyramids kernel when compared with a Bag of Visual Words from the local descriptors}
    \begin{tabular}{|c|c|c|c|c|}
    \hline
    Dataset & \multicolumn{2}{|c|}{\textbf{Holidays}} & \multicolumn{2}{|c|}{\textbf{Oxford 5k}}\\
    \hline
        Method, feaures & RootSIFT & CNN & RootSIFT
        &CNN \\
    \hline
    \hline
    Baseline (BoW) & 56.3 & 49.8 & 31.1 & 32.4 \\
        SP SLEM    & 66.9 & 70.2 & 42.4 & 45 \\
    \hline
    \end{tabular}
    \label{tab:spk}
\end{table}
All previous kernel functions take as input image representations in a fixed sized vectorial space. In this section, we propose using the spatial pyramid kernel \cite{GrauDa05}, that takes as input a set of local descriptors and its coordinates in the image.

We revisit the spatial pyramid scheme presented in \cite{spk} using the generalized intersection function of histograms.

Let $X$, $Y$ be the sets of local descriptors of a pair of images.
These local descriptors can be associated to one of $M$ codewords $\{c_1,c_2,...,c_M\}$, so that $X=\bigcup_{1\leq m \leq M}X_m$ and $Y=\bigcup_{1\leq m \leq M}Y_m$, so that $X_m$, $Y_m$ are the set of descriptors associated to $c_m$.
For level $l$ in $\{0,1,..., L\}$, we divide the images in grids $2^l\times 2^l$ and we call $H^{(l,m)}_X$, $H^{(l,m)}_Y$ the $4^l$-dimensional histograms of the occurrences of the of the visual word $c_m$ in each bin of the grid.
The kernel $K_{sp}(X,Y)$ is given by sum over all codewords of the intersection histogram of $H^{(l,m)}_X$ and $H^{(l,m)}_Y$, $l=0,...,L$,  weighted proportionally to the number of bins in each level.

We test this kernel for RootSIFT extracted in multiscales, as used in \cite{spk}, and learn the codewords with K-means. Inspired by the results of \cite{SPPCNN}, we also test it for activations of the last convolutional layer of the network of \cite{SimonZisser15}. Each neuron of the last layer correspont to a codeword. The results are presented in Table \ref{tab:spk}.

%The spatial pyramid kernel is given by
%\begin{align}
%    K_{sp}(X,Y) &= \sum_{m=1}^M \kappa^L(X_m, Y_m), \ where \\
%    \kappa^L(X_m,Y_m) &= 2^{-L}I^{l}(X_m,Y_m) +\sum_{l=1}^L 2^{-L+l-1}I^{l}(X_m,Y_m) \label{kappa}\\
%    I^{l}(X_m,Y_m) = \sum_{i=1}^{4^l}\min()
%\end{align}

%% Local Variables:
%% TeX-master: "main_eccv"
%% End:

### 2016-11-07 sec_eval_raf.tex
first commit

\section{Experimental Evaluation}
\label{eval}

\subsection{Datasets and evaluation protocol} \label{eval:protocol}
We perform our experiments in three standard datasets for image retrieval. 
\begin{itemize}
    \item The INRIA \emph{Holidays} dataset \cite{holidays} consists of 1491 images divided in 500 groups of matching images. We manually rotate by 90 degrees some images that are not in their natural orientation to compensate for the fact that CNN features are not rotation invariant~\cite{babenko14,Arandjelovic15,RaToCh16}
    \item The \emph{Oxford5k} dataset \cite{oxford}, which consists of 5063 images separated in 55 groups of matching images, each group associated to a landmark of Oxford. We use the "full" crop, ignoring the region of interest of each image.
    \item The \emph{Oxford105k} dataset \cite{oxford}, large-scale dataset containing the same images and queries from Oxford5k plus \emph{Flickr100k}, a collection of $1e^5$ distractor images Flickr.
\end{itemize}
As pool of negative images to build SLEM, we use the Flickr100k for both Holidays and Oxford5k. When training on Oxford105k, we us instead the \emph{Paris} dataset \cite{PhiChIsSiZi08} as negative samples.
%In both datasets, each group contains one query image, the other images in the group being the only correct answers to the query.
%For each query image, we calculate its similarity to all other images in the database and rank them in decreasing order.
%The average precision of a group is derived from the ranking of the images of the group for the similarity with the corresponding query image.
%The final mean average precision (mAP) for a dataset is the mean of the average precision over all its groups.
%As pool of negative images to build SLEM, we use the Flickr100k collection \cite{oxford}, composed of $10^5$ images from Flickr. For a full rank decomposition, we use only a subset of it, containing between 6000 and 15000 images.
%As stated in section \ref{low-rank} and further discussed in section \ref{time-scale}, a full rank decomposition does not scale well for bigger number of negative samples. For low rank decomposition, we use all 100000 images.

At evaluation time, for a dataset that consists of $p$ images and $q$ query images, we calculate its $p\times q$ \emph{similarity matrix} $S$, where each of its $q$ columns is the matching scores of the query image with all the $p$ images. \hlc{RR: maybe take out this paragraph completely.}

\subsection{Which kernel to choose?}
For the experiment of SLEM as a feature encoder, we tested two different kernels --Gaussian and polynomial-- each with a scalar parameter $\gamma$:
\begin{align}
    k_1(x,y) = e^{-\gamma\|x-y \|^2}; \ 
    k_2(x,y) = x^Ty+\gamma(x^Ty)^2. \label{kernels}
\end{align}
\textbf{Gaussian SLEM.} The radial basis function kernel $k_1$ is a
well known reproducing kernel, used for classification with support vector machines.\\
\textbf{Poly SLEM.} The polynomial kernel $k_2$ is a reproducing kernel often used in natural language processing.
%\textbf{SPM SLEM} The spatial pyramid matching kernel of $L+1$ levels in Equation \ref{k:spm} take as input a set of local descriptors and its location in pyramidal bins \cite{spk}.

\subsection{Base visual features}
We test our feature encoder for four different base features $x$ in$\mathbb{R}^d$; the hand-crafted VLAD image representation and three learned features derived from the activation coefficients of deep Convolutional Neural Networks.

We use the same VLAD variant of \cite{Delhumeau2013} used in \cite{ZePe15} that relies on densely-extracted rootSIFT \cite{3things} local descriptors, per-cluster, PCA-based rotations, and root normalization. Like \cite{ZePe15}, we use $64$ clusters, for a final feature of size $8192$.

The first CNN-based features we use consist of the activation coefficients of the previous-to-last layer of the AlexNet architecture \cite{Krizhevsky2012}, based on a publicly available pre-trained model \cite{jia2014caffe}. These are also the features used in \cite{ZePe15}.

The Sum-Pooling of Convolutional (SPoC) features \cite{babenko15}, which are tailored specifically for the image retrieval application, consist of spatially-weighted sums of the activations of the last convolutional layer of a 19-layer CNN \cite{SimonZisser15}.

Finally, we use the NetVLAD features \cite{Arandjelovic15}, trained for place recognition. These features are obtained by adding a differentiable version of the VLAD algorithm~\cite{Delhumeau2013} as a layer at the end of a convolutional architecture.

%We revisit the VLAD feature presented in \cite{VLAD} as an example of a hand-crafted representation. First we extract a set $\mathcal{F}$ of local descriptors of the image $I$. We use the 128 dimension RootSIFT \cite{3things} descriptors, extracted densely.
%Then, we hard-assign each descriptor $f$ in $\mathcal{F}$ to the closest among $K$ pre-trained codewords $c_k$, $k\in\{1\cdots K\}$,
%one of $K$ set $\mathcal{C}_k$ of descriptors associated to codewords $\{c_k\}_{1\leq k\leq K}$, 
%and map $f$ to a $\RR^{128K}$ vector
%\begin{equation}
%\phi^{VL}_1(f) = \left[0 \cdots 0\quad \Phi_k^T\frac{(f-c_k)}{\|f-c_k\|} \quad 0 \cdots 0\right],
%\end{equation}
%where $\Phi_k$ is a $128\times 128$ PCA matrix learned on training features mapped to $k$-the codeword.
%associated to descriptors in $\mathcal{C}_k$. 
%The final VLAD representation is the power-normalization and $l_2$ normalization of the sum-pooling of $\phi^{VL}_1$:
%\begin{equation}
%\phi^{VL}_2(I) \propto \mathrm{power}\big(\sum_{f\in \mathcal{F}}\phi^{VL}_1(f)\big),~\|\phi^{VL}_2(I)\|=1,
%\end{equation}
%with scalar power normalization $\mathrm{power}(v)=\mathrm{sign}(v)|v|^{0.2}$ applied component-wise.
%In experiments, we use $K=64$ codewords learned on images from Flickr. Our VLAD representation has $d=8192$ dimensions.

%Convolutional features obtained from very deep convolutional neural networks (CNNs) have been shown to work as good local descriptors for matching \cite{SimonZisser15}. The SPoC representation \cite{babenko15} is a weighted sum-pooling of the activations of the last convolutional layer of a 19-layer CNN. For a given input image $I$, the activations in this layer are organized over a $W\times H$ spatial grid and over $d$ channels. Each position $(w,h)$ in $\{1 \cdots W\}\times \{1\cdots H\}$ can thus be equipped with a descriptor $f_{h,w}(I)$ in $\RR^D$.
%Indeed, if our last convolutional layer has $D$ neurons, and each neuron a $W\times H$ map of activations of this neurons to the image $I$, 
%each pair $(w,h)$ with $w$ in $\{1,2,..., W\}$ and $h$ in $\{1,2,...,H\}$ can be associated to a descriptor $f_{(h,w)}$ in $\RR^D$ of the responses of each neuron at coordinate $(w,h)$ of the maps. 
%We then sum-pool these descriptors, weighted accordingly to their distance to the center of the image:
%\begin{equation}
%    \phi^{SPoC}(I) = \sum_{w=1}^W\sum_{h=1}^H \alpha_{w,h}f_{w,h}(I),
%\end{equation}
%where
%\begin{equation}
%    \alpha_{w,h} = \exp \left(-\dfrac{(w-W/2)^2+(h-H/2)^2}{2\sigma^2}\right).
%\end{equation}
%In our experiments, we follow the implementation details of \cite{babenko15}: We resize all images to $586\times 586$ pixels before feeding them to the network. The last convolutional layer has $d=512$ channels with activation maps of size $(W,H)=(37,37)$, and we set $\sigma=\frac{H}{3}$. We use the same network architecture to extract the convolutional maps but different weights in the convolutional layer, which might explain the difference between the results in Table \ref{fullrank:results} and \cite{babenko15}.

%Finally, we also test a more convencional CNN feature, less deep and using two fully connected layers after the convolutional layers.
%Our final representation is a $4096$ non-negative feature. We based out implementation on CAFFE \cite{jia2014caffe}.

\subsection{Image retrieval results}

We test our method against the original ESVM, PCA and LDA algorithms. The results are presented in Table \ref{fullrank:results}.
\textit{Babenko et al.}~\cite{babenko15} and \textit{Arandjelovi\'c et al.}~\cite{Arandjelovic15} improve the retrieval results applying a PCA compression followed by whitening. 
Thus for these features, we considere their version as presented in \cite{Arandjelovic15,babenko15} in the \textit{PCA+whitening} row and the not-compressed, not-rotated and not-whitened version as \textit{Baseline}. 
For VLAD and AlexNet, we considere the version presented in~\cite{ZePe15} as \textit{Baseline} and a rotated, whitened and compressed to half its original dimensions as \textit{PCA+whitening}. 
As for the version of each feature to which we apply the other algorithms, it changes according to the dataset. 
For SPoC, we apply LDA, ESVM and SLEM to the baseline, as a replacement of PCA. 
For NetVLAD, we use the rotated and compressed version of the trained networks correspondent to each dataset as suggested by~\cite{Arandjelovic15}. 
For Holidays we obtain better results with whitened features and for Oxford5k, non-whitened.

For \textit{LDA}, we follow our description in section \ref{sec:lda} and calculate $\Sigma$ as the convariance of the negative samples.
%In our experiments, the compression worsens the results when compared with non-compressed. \textit{Delhumeau et al.}~\cite{Delhumeau2013}
%Hence we compare the improvements of PCA plus whitening, without compression, with our method.
Linear SLEM performs similarly to ESVM despite being much more time efficient (see discussion in section \ref{time-scale}). 
The fact that a hinge-loss classifier does not outperform a square-loss classifier can seem counter-intuitive, but both were shown to be equivalent for binary classification under mild constraints~\cite{YeXi07}.

We perform both Gaussian SLEM and Polynomial SLEM with two decompositions: one full-rank CCD decomposition signaled by (fr) and two low-rank KPCA decompositions signaled by the rank of the decomposition. 
We train our exemplar classifiers for between $6000$ and $10000$ negative samplers. We outperform all methods, for all datasets and all image representations. \hlc{RR: Do we? I'm putting it here because I'm an optimistic, but we have to wait the results.}

\begin{table*}[h!]
\begin{center}
\setlength{\tabcolsep}{.2em}
%\begin{tabular}{|c|c|c|c|c|c|c|c|}
%\small
\begin{tabular}{c@{\hskip 1em}cccc@{\hskip 1em}cccc@{\hskip 1em}cc}%@{\hskip 1em}c}
\toprule
Dataset & \multicolumn{4}{c}{\textbf{Holidays}} & \multicolumn{4}{c}{\textbf{Oxford 5k}} & \multicolumn{2}{c}{\textbf{Oxford 105k}} \\%& \textbf{Oxford 105k}\\
\midrule
Model, features & VLAD  & SPoC & AlexNet & NetVLAD & VLAD & SPoC & AlexNet & NetVLAD & SPoC & NetVLAD\\% & SPoC\\
\midrule
Baseline            & 72.7 & 76.4 & 68.2  &  85.4 & 46.3 & 54.4 & 40.6 & 67.5 & - & - \\% & 50.1\\
%Whitening           & -    & -    & -    & -    & -\\
PCA+whitening       & 69.4 & 81.6 & 69.2 & 88.3  & 50.9  & 63.7 & 45.0 & 69.1 & - & - \\% & 55.5\\
LDA                 & -   &   -   &  -   &   -   &   -   &   -  & 40.5 &  -  & - & - \\
E-SVM               & 77.5 & 84.0 & 71.3 & 90.5 & 57.2  & 62.1 & 43.9 & 70.4(?) & - & - \\% & 56.5\\ netvlad_hol: 91.6
Linear SLEM         & 78.0   & 82.3 & 72.1 & 90.9& \ul{59.3}  & 64.1 & 46.9/42.2(?) & 72.4 & - & - \\% & 56.7\\
Gaussian SLEM (16)  & - & - & - & 91.1 & - & - & - & 70.8 & - & - \\
Gaussian SLEM (32)  & - & 82.2 & - & 91.0 & - & - & - & 71.1 & - & - \\
Gaussian SLEM (fr)  & 78.1 & 85.2 & \ul{72.9} & 91.4 & 59.0   & \ul{64.9} & 47.0 & \bf{73.2} & - & - \\% & 59.5\\
Poly SLEM (16)      & - & 82.3 & - & 91.3 & - & 63.6 & - & 71.2 & - & - \\
Poly SLEM (32)      & - & 82.4 & - & \bf{91.7} & - & 63.6 & - & 71.7 & - & - \\
Poly SLEM   (fr)    & 78.1 & \ul{86.3}  & \ul{72.9} & \bf{91.7} & \ul{59.3}  & 64.8 & \ul{47.3} & \bf{73.2} & - & - \\% & \bf{62.3}\\
%Zepeda \textit{et al.} \cite{ZePe15} & \underline{78.3} & - & 71.8 & - & 57.5 & - & 44.6 & - & - \\
%Babenko \textit{et al.} \cite{babenko15} & - & 81.8 & - & - & - & \textbf{65.7} & - & - & 57.8\\
\hline
\end{tabular}
\caption{Mean average precision results for INRIA Holidays and Oxford buildings datasets, expressed as percentages. In this table, we present our results for VLAD \cite{Delhumeau2013}, sum-pooling of convolutional features (SPoC) \cite{babenko15}, activation coefficients from the previous-to-last CNN layer (AlexNet) \cite{Krizhevsky2012} and activation of NetVLAD layer~\cite{Arandjelovic15}. In parenthesis, the rank of he decomposition ('fr' for full rank decomposition)}
\end{center}
\label{fullrank:results}
\end{table*}

For all the experiments we calibrate the regularization cost $\lambda$, as well as the parameter $\gamma$. 
%For some experiments, specially using VLAD, the optimal value of $\gamma$ is rather small ($\gamma\sim 0.05$), which suggests the linear kernel is already close to the optimal kernel for these features.

\subsection{Time and storage scalability} \label{time-scale}
In this section we compare the time efficiency of our method and the E-SVM, as well as discuss which method to use accordingly with the number of positive and negative samples.

In Figure \ref{fullrank:results}, we that the see Linear SLEM efficiency does not change with $n$.
Indeed, if $d$ is the dimension of the base representation, $A$ is a $d\times d$ matrix for linear SLEM, whereas for a full rank kernel, $A$ is $n\times n$.
This explains the increasing running time for Gaussian and polynomial kernels: storage and solving a $n\times n$ system does not scale for large number of negative samples.
Retrieval results for kernelized SLEM in Figure \ref{fullrank:results} suggest we can benefit from larger sets of negative samples, but its improvements are not steadily. Indeed, the calculation time for a full rank kernel matrix offline is $O(n^3)$, as it can be seen in Figure \ref{fullrank:results}.
When in consider only the online procedure of our model, \emph{i.e.} the calculation of $\beta^\star$, our kernelized model similarly to ESVM. Therefore, we can process the kernel SLEM for the Gaussian and polynomial kernel similar running time to ESVM if we pre-process our negative samples offline.
For the low-rank decomposition \hlc{not yet presented in Figure \ref{fullrank:results}}, the offline procedure is $O(n^3)$, similarly to the full rank. However this step is much more time consuming for the low-rank, as shown in Figure \ref{fig:residue}. We need test different values of $n$, but for fixed rank $r$, we should expect better results for smaller $n$.

\vspace{3 mm}

\input{figuren2.tex}

\subsection{Comparation to the state of the art}
Results shown in Table~\ref{sota:results}. We do not include re-ranking nor query expansion.

\begin{table}[t]
\begin{center}
\setlength{\tabcolsep}{.2em}
\small
\begin{tabular}{l|c|ll}
\toprule
Features & dim & \textbf{Hol.} & \textbf{Ox5k} \\%& \textbf{Oxford 105k}\\
\midrule
Babenko \textit{et al.}\cite{babenko15}  & 256 & 80.2 & 58.9 \\ %& 57.8\\
Radenovi\'c \textit{et al.} \cite{RaToCh16}   & 256 & 81.5 & \un{77.4} \\
Arandjelovi\'c \textit{et al.} \cite{Arandjelovic15}& 256 & 86.0 & 62.5 \\ %& - \\
Kalantidis  \textit{et al.} \cite{KaMeOs16}   & 256 & 83.1 & 65.4 \\
SPoC + Linear SLEM & 256 & 82.3 & 64.1 \\
SPoC + Poly SLEM & 288 & 82.3 & 60.6 \\
SPoC + Poly SLEM & 320 & 82.4 & 60.7 \\
NetVLAD + Linear SLEM & 256 & \un{88.5} & - \\
NetVLAD + Poly SLEM & 288 & 87.7 & 65.5 \\
NetVLAD + Poly SLEM & 320 & 88.3 & 65.4 \\
\midrule
Radenovi\'c \textit{et al.} \cite{RaToCh16}   & 512 & 82.5 & 79.7 \\
Arandjelovi\'c \textit{et al.} \cite{Arandjelovic15}& 512 & 86.7 & 65.6 \\
Kalantidis \textit{et al.} \cite{KaMeOs16}   & 512 & 84.9 & 68.2 \\
Gordo \textit{et al.} \cite{GoAlReLa16} & 512 & 89.1^{\dag} & \bf{83.1}^{\dag} \\
NetVLAD + Linear SLEM & 512 & 89.3 & 68.9 \\
NetVLAD + Poly SLEM & 544 & 89.9 & 69.0 \\
NetVLAD + Poly SLEM & 576 & \un{89.9} & 69.2 \\
\midrule
Arandjelovi\'c \textit{et al.} \cite{Arandjelovic15}& 4096 & 88.3 & 69.1 \\
Li \textit{et al.} \cite{LiLaHa15}& - & 89.2^{\dag} & 73.7 \\
NetVLAD + Linear SLEM & 4096 & 90.9 & 72.1 \\
NetVLAD + Poly SLEM & 4128 & 91.3 & 71.2 \\
NetVLAD + Poly SLEM & 4160 & \bf{91.7} & 71.7 \\
%Zepeda \textit{et al.} \cite{ZePe15} & \underline{78.3} & - & 71.8 & - & 57.5 & - & 44.6 & - & - \\
%Babenko \textit{et al.} \cite{babenko15} & - & 81.8 & - & - & - & \textbf{65.7} & - & - & 57.8\\
\hline 
\end{tabular}
\caption{Compared results to state-of-the-art features at similar dimensions, without re-ranking or query augmentation. The results using Poly SLEM or Gaussian SLEMa dd 32 or 64 dimensions to the original feature (for $r=16$ or $r=32$, respectively). Underlined results are the best at each dimension bracket and bold results are the general best. $\dag$ indicates the previous state-of-the-art.}
\end{center}
\label{sota:results}
\end{table}

%\subsection{Semantic correspondence results}

\subsection{Classification results}

### 2016-11-07 sec_intro.tex
first commit

\section{Introduction}

Effective image 
representations are a crucial component of image retrieval systems 
where, given some query image,  images must be retrieved from a large 
database, then ranked according to their degree of similarity with the 
query.
%Robust image representations are a crucial component of a vast number of computer vision applications. 
%Within these, the \emph{image retrieval} application is an important example wherein a query image is provided to the system, which must in turn find all matching images from within a large, unannotated database. 
The matching process is done entirely based on the pixel content of the query and database images, and the search must be robust to large image variations due to camera pose, color differences and scene illumination, amongst others.

The success of several existing systems in this challenging application has been enabled by advances in image representation. %An image representation function commonly maps a given input image to a vector in a fixed dimensional space called the image \emph{feature vector}. 
Indeed, once the feature space is given,
the image retrieval process is reduced to finding the feature vectors closest to the query image under, \textit{e.g.}, the Euclidean distance. An adequate image representation must thus be robust to image variations with preferably low computation and storage cost.

%A now pervasive example of an image feature representation is that consisting of the activation coefficients extracted from the previous-to-last layer of Convolutional Neural Networks (CNNs) \cite{Krizhevsky2012}. Although CNNs are trained in a fully-supervised manner for the image classification task, they have been shown to transfer well not only to new, unseen classes \cite{Oquaba,Chatfield2014}, but also to alternate tasks such as object detection \cite{Girshick2014} and, importantly, image retrieval \cite{Sharif}.

Many successful feature representations for image retrieval rely on unsupervised models such as $K$-means \cite{Delhumeau2013} or Gaussian mixture models \cite{Perronnin2010}. Before the neural networks \textit{renaissance}, these representations would outperform methods that exploit supervised learning of image features directly for the image retrieval task \cite{Bilen2015,Rana}.
Today, with the popularization of convolutional architectures, many representations consist of aggregation and/or pooling of the last convolutional layer \cite{babenko15,GoAlReLa16,KaMeOs16}.
Others propose new differentiable layers based on successful retrieval representations to be added at the end an existing architecture \cite{Arandjelovic15}.
%One of the main reasons for this is the lack of adequately large and varied supervised datasets that are expensive to collect.

The exemplar support vector machine (ESVM), originally proposed by Malisiewicz {\it et al.} \cite{Malisiewicza}, leverages the availability of large, unannotated pools of images within the context of supervised learning. It uses a large generic pool of images as a set of negative examples, while using a single image (the \emph{exemplar}) as a positive example. Given these training sets, a classifier is learned that can  generalize well, despite the drastically limited size of the set of positive examples.

Zepeda \emph{et al.} \cite{ZePe15} propose to instead treat the weights of the resulting classifier as a new feature vector for the exemplar image. %This new ESVM feature is derived from \emph{base} feature representations ({\it e.g.,} CNN activation features) of the exemplar and the negative pool.
%An extension \cite{ZePe15} of the ESVM formulation discussed above instead treats the resulting  classifier as an \emph{enhanced} representation of the image feature representing the exemplar image (the \emph{base} feature representation).
An ESVM feature is extracted from each database image, as well as from the query image, by treating each image as an exemplar while keeping a fixed pool of generic negative images. Searching amounts to computing distances between the query and the database ESVM features. Note that ESVM features can be derived from arbitrary \emph{base} features ({\it e.g.} CNN activations) of the exemplar and the images in the generic negative pool. An interesting interpretation of this approach is that the ESVM features represent what is unique about the exemplar relative to the pool of generic negatives, which is the role of a classifier.

One important drawback of the ESVM feature encoding approach is that computing the  classifier requires solving an optimization problem for each positive example. This can be time consuming for the large negative pool sizes required for good ESVM feature performance. In this work we propose using the square loss instead of the hinge loss, in effect converting the ESVM problem into a ridge regression that can be solved in closed form. The parameters of the corresponding classifier, which we call the squared loss exemplar machine (SLEM), had been noted to generalize LDA \cite{Koba15} and are shown to form features with comparable performance while drastically reducing the computational cost.

Since computing the SLEM features requires inverting a large matrix related to the training set's covariance matrix, we propose an efficient way to compute this inverse that exploits the fact that only a single (positive) example changes in the training set when computing SLEM features for different images. We show experimentally that our representation can match and even improve upon the performance of ESVM features on two standard datasets using a wide range of base features.

We further introduce a kernel variant of SLEM that enjoys similar computational advantages but improves retrieval performances. Further computational efficiency is obtained using a kernel singular values decomposition of the negative pool.
%low rank incomplete Cholesky decomposition of the negative pool kernel matrix.

The rest of this paper is organized as follows:
%In Section \ref{prior work} we provide an overview of various existing feature representation methods. 
In section \ref{lsesvm} we first review the original ESVM feature representation method and subsequently introduce the proposed linear SLEM model. We then introduce the kernelized SLEM variant in section \ref{nonlinear SLEM}. We then present the low-rank approximation of our method that enables efficient implementations in the non-linear case in section \ref{eff_imp}. We evaluate our proposed method in image retrieval in section \ref{eval}, and present conclusions in section \ref{conclusion}.

% Image search
% - Importance.
% - Feature representations.

% Feature representation - supervised learning
% - Generic framework.
% - Deep CNN features.
% - ImageNet.

% Image retrieval
% - Supervised - Eusipco, NetVLAD.
% - VLAD, Fisher.
% - Cheaper.

% Pool:
% LDA
% SVMs for feature representations - parts learning, Malizsiewics

%\section {Introduction}
% The exemplar SVM (E-SVM) was first introduced by Malisiewicz et al. in \cite{Efros11} as a conceptually simple framework for object detection and image classification where the training set has a small ratio of positive/negative examples. 
% At training time, many SVMs are learned from a large pool of negative against a single positive (so called exemplar). 
% At test time, the scores of the test images for each classifier are fitted by a logistic regression. 
% The final score of an image is then a non-linear combination of scores from multiples exemplar SVMs.

% They have also been used in \cite{Efros12} in image retrieval tasks, using the classifier score to rank matching candidates.
% However this transfer of information from classification to retrieval is severely limited. 
% Indeed, the purpose of a support vector machine is to separate positive and negative samples, i.e., predict discrete labels. 
% The distance between a negative sample and the classification hyperplane has no value as a measure of has \emph{a priori} no value of continuous matching score.

% Zepeda et al. address this problem in \cite{ZePe15} by firstly, performing an E-SVM to each image in a dataset instead of only for the query images and secondly, comparing its classifiers distance to the query's classifier instead of comparing scores. 
% These modifications guarantee we are ranking distances between two points instead of classification scores.
% Therefore \cite{ZePe15} refers to E-SVMs as features encoders: a pipeline that takes an image representation as input and returns an improved image representation. This type of feature enhancing is more akin to methods such as whitening, PCA and LDA.

% This paper introduces the square-loss Exemplar machine (SLEM), which consists of optimizing the same cost function of a regular E-SVM where the hinge loss is replaced by the square loss. 
% The square loss version has the advantage of having a much more efficient optimization. Indeed, the minimization of its cost function is solved by a linear system and can be done for all exemplar simultaneously, whereas the regular E-SVM cost function is generally minimized one exemplar at a time, normally by stochastic gradient descent \cite{bottou10}.
% Also, for the machine learning tasks of binary classification, both regular SVM and least-squares SVM have similar performances \cite{YeXi07}.
% %Also, for the machine learning tasks of binary classification, both regular SVM and least-squares SVM have the same performance if positive and negative samples are separable and the pool of samples is linearly independent \cite{YeXi07}. 
% We also introduce a kernelized version of SLEM, efficiently implemented, that gives better results and superior scalability for large-scale image retrieval problems.

%%% Local Variables:
%%% TeX-master: "main_eccv"
%%% End:

### 2016-11-07 sec_kernel_methods.tex
first commit

\section{The kernel SLEM}
\label{nonlinear SLEM}

\subsection{Introduction to kernel methods}\label{kernel:review}
Let us recall a few basic facts about kernel methods for supervised
classification. We consider a reproducing kernel Hilbert space (RKHS)
$H$ formed by real functions over some set
$X$, and denote by $k$ the corresponding reproducing kernel.  We
address the following learning problem over $H\times\RR$:
\begin{equation}
\min_{h\in H,\nu\in\RR}
\sum_{i=1}^n l(y_i,h(x_i)+\nu) + \frac{\lambda}{2}\|h\|_H^2,
\label{eq:kernel}
\end{equation}  
where the pairs $(x_i,y_i)$ in $X\times \{-1,1\}$, $i=1\dots n$ are training samples, and $l: \{-1,1\}\times \RR\rightarrow\RR^+$ is some arbitrary loss function. 
By definition of a reproducing kernel,
Equation (\ref{eq:kernel}) can be rewritten as
\begin{equation}
\min_{h\in H,\nu\in\RR}
\sum_{i=1}^n l(y_i,\langle \varphi(x_i),h\rangle+\nu) +
\dfrac{\lambda}{2}\|h\|_H^2,
\label{ker:aff}
\end{equation} 
where $\varphi$ is the {\em feature map} over $X$ associated with the
kernel $k$ (which may not admit a known explicit form) and $\langle \cdot, \cdot \rangle$ is the inner product of $H$. We dub problems with the general form of (\ref{ker:aff}) {\em affine}
supervised learning problems since, given some fixed element $h$ of
$H$ and some scalar $\nu$, $\langle h,h'\rangle+\nu$ is an affine function of $h'$,
whose zero set defines an affine hyperplane of $H$ considered itself
as an affine space.

Let $K$ denote the kernel matrix with entries $k_{ij}=\langle\varphi(x_i),
\varphi(x_j)\rangle$ and rows $k_i^T=[k_{i1}, k_{i2},...,k_{in}]$, $i$ in $\{1,\ldots,n\}$.  We assume from now on that $l$ is convex and continuous. Under this assumption, Equation (\ref{ker:aff}) admits an equivalent formulation
\begin{align}
\min_{\alpha\in\RR^{n}, \ \nu\in\RR} \left(\dfrac{1}{n}\sumi l(y_i, k_i^T\alpha+\nu)  +\dfrac{\lambda}{2}\alpha^TK\alpha\right), \label{ker:first}
\end{align}
and any solution $(\alpha^\star,\nu^\star)$ to (\ref{ker:first})
provides
a solution $(h^\star,\nu^\star)$ to (\ref{ker:aff}) with
$h^\star=\sum_{i=1}^n \alpha_i^\star\varphi(x_i)+\nu^\star$. This result follows from the Riesz representation theorem~\cite{SHS01,Wahba90}.

%where $K$ is a $(n+1)\times (n+1)$  matrix and $k_i$ is its $(i+1)$-th column matrix for $i$ in $\{0, 1, ..., n\}$.\\

Assuming our reproducing kernel is semidefinite positive, 
%\footnote{\textit{i.e.} for any $z$ in $\RR^n$, $z^TKz\ge 0$. This is equivalent to all eigenvalues of $K$ being non-negative.} 
$K$ is a semidefinite positive matrix and can be decomposed as $K=BB^T$.
%, with $B$ a rank $r$ upper triangular matrix of size $n\times r$. 
Using this factorization, the kernelized problem can be expressed as 
\begin{equation}
\min_{\beta\in\RR^r,\nu\in\RR} \left( \dfrac{1}{n}\sumi l(y_i, b_i^T\beta+\nu)+\dfrac{\lambda}{2}\|\beta\|^2\right), \label{beta:first}
\end{equation}
where $b_i^T$ denotes the $i$-th row of $B$ and $r$ is the number of columns of $B$. 
If $(\beta^\star, \nu^\star)$ is the solution Equation (\ref{beta:first}), the corresponding vector $\alpha^\star$ (or, more correctly, \textit{a} corresponding vector of dimension $n\geq r$) 
can be computed by $\alpha^\star=P\beta^\star$, where $P$ is the pseudoinverse of $B^T$.

\subsection{Preprocessing of negative samples offline}\label{offline}
Equation (\ref{beta:first}) has the same structure of Equation (\ref{eq:first}) if we were to consider $\theta=0$ and all labels $y_i=-1$.
In that case, all actions described in subsections \ref{slem:intro} and \ref{kernel:review} can be done independently of positive samples. 
They are therefore done offline and their results are stored. In the processing of positive samples, i.e. the online phase, these results are loaded.
The preprocessing phase consists of the calculation of $B$, $\mu = \frac{1}{n}\sum_{i=1}^n b_i^T$ and $A = \frac{1}{n}B^TB-\mu\mu^T +\lambda\mathrm{Id}_r.$

\subsection{Online addition of a positive sample}
\label{subsec:adding}
We now reconsider the exemplar problem, with one positive example $x_0$ and $n$ negative examples $X$. Here $K$ is the $n\times n$ kernel matrix of the negative samples $X$. 
%Let us assume the factor $B$ is given. %The calculation of a kernel Cholesky factorization is solved efficiently by \cite{BaJo05,BaJo02}. 
%We further analyse how to compute this factorization in subsection \ref{offline}.

We denote by $K'$ the augmented kernel matrix obtained by adding the positive samples $x_0$. Such matrix can be written as
\begin{equation}
K' = \begin{bmatrix}
k_{00} & k_0^T\\
k_0 & K
\end{bmatrix},
\end{equation}
where $k_{00}=\langle \varphi(x_0),\varphi(x_0)\rangle$ is a scalar and $k_0= [\langle \varphi(x_0),\varphi(x_i)\rangle]_{1\le i\le n}$ is a vector in $\RR^n$. 
The following lemma show how the factorization of $K'$ can be derived from the factorization of its sub-matrix $K$ and the solution of a $n\times n$ linear system.

\begin{lemma} The augmented kernel matrix $K'$ can be factorized as $K'= B'B'^T$ with
%\begin{align}
%K'&= B'B'^T,\quad\text{where}\quad
%B'=\begin{bmatrix}
%u & v^T\\0 & B
%\end{bmatrix}\\
%v&=B^\dagger k_0,\,\, u=\sqrt{k_{00}-||v||^2}.\label{eq:lemma1}
%\end{align}
\begin{equation}
B'=\begin{bmatrix}
u & v^T\\0 & B
\end{bmatrix},~
v = B^\dagger k_0,~ u=\sqrt{k_{00}-||v||^2}.
\label{eq:lemma1}
\end{equation}
\end{lemma}\label{lemma1}
\begin{proof}
For $B'$ defined by (\ref{eq:lemma1}), we have that
\begin{equation}
B'B'^T = 
\begin{bmatrix} u^2+\|v\|^2 & v^TB^T\\ 
Bv& BB^T\end{bmatrix} 
=\begin{bmatrix} k_{00} & v^TB^T\\ 
Bv& K\end{bmatrix} .
\end{equation}
Since $K'$ is positive semidefinite, $k_0$ must lie in the column space $\mathcal{B}$ of $B$.
%\footnote{
Indeed, if we suppose $k_0$ does not belong to $\mathcal{B}$, then it can be decomposed uniquely as $k_0=s+t$, $s\in\mathcal{B}$ and  $t\in\mathcal{B}^\perp$, with $t\ne 0$. In one hand, $K'$ being semidefinite positive implies that $[1, -at^T]K'[1; -at]=k_{00}-2a\|t\|^2\ge 0$ for all real value $a$. In the other hand, for $a$ large enough, $k_{00}-a\|t\|^2\le 0$, which is a contradiction.
%} 
Hence $v=B^\dagger k_0$ is an exact solution of $Bv=k_0$. The fact that $k_{00}-\|v\|^2$ is non-negative comes from the fact that the Schur complement $K-k_0 k_0^T / k_{00}$ of $k_{00}$ in $K'$ is itself positive semidefinite.
%\footnote{
Indeed, since the matrix $k_{00}K-k_0k_0^T=B(k_{00}\mathrm{Id}_r-vv^T$ is also positive semidefinite. Thus $v^T(k_{00}\mathrm{Id}_r-vv^T)v = \|v\|^2(k_{00}-\|v\|^2)\ge 0$. \qedhere
%}
\end{proof}

This lemma allows us to add a positive sample to Equation (\ref{beta:first}).
%$\beta^\star(x_0,X), \nu^\star(x_0,X) = \underset{\beta\in\RR^{r+1}, \nu\in\RR}{argmin}J'(\beta, \nu)$, for
With a positive exemplar, Equation \ref{beta:first} now reads
\begin{align}
 \theta\ l(1, b_0'^T\beta+\nu) +\dfrac{1}{n}\sumi l(-1,b_i'^T\beta+\nu)
+\dfrac{\lambda}{2}\|\beta\|^2,\label{beta:final}
\end{align}
with $b_i'^T$ being the $(i+1)$-th row of $B'$, $i$ in $\{0,1,...,n\}$. In particular, $b_0'=[u; \ v]$ and, for $i>0$, $b_i'=[0; \ b_i]$. Solution $(\beta^\star,\nu^\star)$ in $\RR^{r+1}\times\RR$ can be computed just as before by Equation (\ref{omega:solution}), replacing $x_0$ by $b_0'$, $\mu$ by $\mu' = \frac{1}{n}\sum_{i=1}^n b_i'$ and $X$ by the $(r+1)\times n$ matrix $Q$ of columns $b_1', b_2',...,b_n'$. $\alpha^\star$ is now calculated as
$\alpha^\star=P'\beta^\star$, where $P' = [u^{-1} \ 0^T; -u^{-1}Pv \ P]$ is the pseudoinverse of $B'^T$. This can be expressed by the linear system

\begin{equation}
\begin{bmatrix} \alpha_0^\star \\ \hat{\alpha}^\star \end{bmatrix} = \begin{bmatrix} \frac{1}{u} & 0^T \\-\frac{1}{u}Pv & P  \end{bmatrix} \begin{bmatrix}\beta_0^\star \\ \hat{\beta}^\star \end{bmatrix}.\label{alpha:to:beta}
\end{equation}

\subsection{Similarity score}\label{simi_score}
Once the optimal parameters $(\beta, \nu)$ from (\ref{beta:final}) and the coordinates $u$, $v$ of $b_0'$ from (\ref{eq:lemma1}) have been found\footnote{We drop the ``$\star$'' in this subsection to avoid cluttering the notation.}, they can be used directly for measuring similarity between matching images.

Indeed, suppose two image representations $x_0$ and $x_0'$ are given and we wish to calculate the similarity score between its SLEM representations $h$ and $h'$, denoted by $s(h,h')$. $h'$ can be written as $h'=\alpha_0'\varphi(x_0')+\sum_{i=1}^n \alpha_i'\varphi (x_i)+\nu'$. Using Equation \ref{alpha:to:beta} and ignoring bias $\nu$ and $\nu'$ which have empirically no influence, $s(h,h')$ is given by:
\begin{equation}
\begin{split}
s(h,h') & = \langle h, h'\rangle \\
		& = \hat{\alpha}^{T} K\hat{\alpha}'+\alpha_0k(X, x_0)^T\hat{\alpha}'+\alpha_0'k(X, x_0')^T\hat{\alpha} \\
		& \ \ \ \ \ +\alpha_0\alpha_0'k(x_0,x_0')\\
		& = \hat{\beta}^T\hat{\beta}'+\frac{\beta_0\beta_0'}{uu'}(k(x_0,x_0')-v^Tv').
\end{split}
\end{equation}

We need to store $x_0$, $\hat{\beta}$ and $v$ to calculate its matching score to whichever other image. Each image depends therefore on a vector of $p+2r$ dimensions.

%%% Local Variables:
%%% TeX-master: "main_eccv"
%%% End:

### 2016-11-07 suppmat.tex
first commit

\documentclass[runningheads]{llncs}

%\usepackage{times}
%\usepackage{epsfig}
\usepackage{graphicx}
\usepackage{caption}
%\usepackage{subfigure}
\usepackage{subcaption}
\captionsetup{compatibility=false}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{ruler}
\usepackage{color}
\usepackage{booktabs}
\usepackage{mathtools}
%\usepackage{pgfplots}
\usepackage[width=122mm,left=12mm,paperwidth=146mm,height=193mm,top=12mm,paperheight=217mm]{geometry}
%\newtheorem{proposition}{Proposition}
%\newtheorem{lemma}{Lemma}
%\newtheorem{proof}{Proof}
\newcommand{\RR}{\mathbb R}
\newcommand{\NNN}{\mathcal X}
\newcommand{\sumi}{\displaystyle{\sum_{i=1}^n}}
\DeclareMathOperator*{\argmin}{argmin}

% Highlighting
\usepackage{soul}
\usepackage{color}
%\usepackage[usenames,dvipsnames]{xcolor}
\newcommand{\hlc}[2][yellow]{{\sethlcolor{#1}\hl{#2}}}
\newcommand{\RAF}[1]{\hlc[yellow]{(RR:) #1}}
\newcommand{\JZ}[1]{\hlc[pink]{(JZ:) #1}}
\newcommand{\PP}[1]{\hlc[green]{PP: #1}}

\usepackage{tikz}
\usepackage{pgfplots,pgfplotstable}
\usetikzlibrary{pgfplots.groupplots}
\pgfplotsset{grid=major,height=2in, width=\columnwidth}
\input{tikz_styles.tex}

%\input{test_tikz_data.tex}

\begin{document}
% \renewcommand\thelinenumber{\color[rgb]{0.2,0.5,0.8}\normalfont\sffamily\scriptsize\arabic{linenumber}\color[rgb]{0,0,0}}
% \renewcommand\makeLineNumber {\hss\thelinenumber\ \hspace{6mm} \rlap{\hskip\textwidth\ \hspace{6.5mm}\thelinenumber}}
% \linenumbers

\pagestyle{headings}
\mainmatter
\def\ECCV16SubNumber{1390}  % Insert your submission number here

\title{Kernel Square-Loss Exemplar Machines\\ For Image Retrieval: Supplementary Material} % Replace with your title

\titlerunning{ECCV-16 submission ID \ECCV16SubNumber}

\authorrunning{ECCV-16 submission ID \ECCV16SubNumber}

\author{Anonymous ECCV submission}
\institute{Paper ID \ECCV16SubNumber}

\maketitle
%\thispagestyle{empty}

We present here some additional experiments and a discussion of 
different variants of the approach proposed in our submission.

\section{Rotation of Holidays' images}
In our submission, we compare our results to those obtained in the 
original SPoC paper of Babenko and Lempitsky \cite{babenko15}. We were surprised by not 
be able to reproduce their results on the Holidays Benchmark: 
Indeed, \cite{babenko15} reports a mAP of 81.8, but we only managed to obtain a mAP 
of 77.5 (SPoC+PCA+whitening) in our implementation. Upon careful 
examination of their paper, this is due in large part to the fact that 
Babenko and Limpitski manually rotate by 90 degrees some images that are 
not in their natural orientation to compensate for the fact that CNN 
features are not rotation invariant \cite{babenko15,babenko14}. Applying the {\em same} 
rotations to the {\em same} images boosts the performance of our 
implementation of their method to 81.6 mAP (the 0.2 difference is due to a different set of negative images used to learn the PCA matrix and singular values), and the performance of 
our full method (Poly SLEM) to a mAP of 86.1, clearly outperforming \cite{babenko15}. 
Note that, as required by the ECCV submission rules, these results are 
obtained using the {\em exact same} implementation and parameters as in 
our original submission, simply rotating the input images as in \cite{babenko15,babenko14}.
This phenomenon is not observed for the Oxford dataset, where all 
images appear in their "natural" orientation.

These results, shown in Table \ref{tab}, are comparable to the state of the art for the Holidays dataset which, accordingly to \cite{holidaysSOTA}, is 88.1 for all descriptors and 84.7 for global features such as the SPoC+SLEM. To our knowledge, the state of the art for global features in Holidays is 86.0 is \cite{netVLAD}, which again uses the same rotations as [3] and the additional 
experiments reported in this section.

\begin{table*}[!h]
\begin{center}
\caption{Mean average precision results for INRIA Holidays, expressed as percentages. On the left, results for Holidays dataset as presented in our submission. On the right, results for Holidays dataset with images on their natural orientation.}
\setlength{\tabcolsep}{.2em}
%\begin{tabular}{|c|c|c|c|c|c|c|c|}
\small
\begin{tabular}{c@{\hskip 1em}c@{\hskip 1em}c}
\toprule
Dataset & \textbf{Holidays} & \textbf{Holidays (rot)}\\
\midrule
Model, features  & SPoC & SPoC\\
\midrule
Baseline            & 73.1 & 76.4\\
PCA+whitening       & 77.5 & 81.6 \\
E-SVM               & 79.9 & 84.6 \\
Linear SLEM         & 78.3 & 82.3 \\
Gaussian SLEM       & 81.4 & 84.9 \\
Poly SLEM           & \underline{82.0} & \textbf{86.1} \\
Babenko \textit{et al.} \cite{babenko15} & - & 81.8\\
\hline
\label{tab}
\end{tabular}
\end{center}
\end{table*}

\section{Low rank decomposition evaluation}
We present here a more complete evaluation of the low rank decomposition. In this section, we use SPoC features as our base encoding, the polynomial kernel as our kernel function and Holidays dataset (not rotated, for lack of time) images.
As discussed in our submission, a significantly larger pool does not improve results after $15000$ negative examples. Indeed, for a pool of $100000$ negative examples, we obtain mAP 81.5, compared to 82.0 for $14500$ negative examples.
In Figure \ref{full_low}, we take a pool of 100000 negative examples and show their results (mAP and computation time) in green. For this same pool of negatives, we evaluate their low rank decomposition implementation of $r'\in\big\{ \lfloor \frac{10^5}{2^t}\rfloor, \ 0\le t\le 11 \big\}$, and shown their results in black.

We obtain mAP results close to full rank for small values of rank $r'$, at a fraction of the computation time of the full rank. 
As $r'$ increases, performance increases, but the computational cost eventually overcomes that of the full-rank method (around $r'=10000$).

It is not clear today whether it is advantageous to use a lower-rank approximation of the kernel or fewer negative samples.
\input{fig_low_n.tex}
%\input{fig_low_vs_full.tex}
%We then compare the low rank results for rank equals $r'$ to a full rank decomposition of a pool of $r'$ negative examples, for very small values of $r'$. What we observe is that a full rank decomposition obtain better retrieval results, even

\section{More large scale retrieval results}
In our submission, we compare our large scale results to those obtained in \cite{ZePe15} with recursive ESVM (RESVM-2) in the Oxford 5k benchmark using VLAD-64 as base features. 
We did not had at the time finished the same experiment for the Holidays dataset, in part due to our results not being superior to those presented in \cite{ZePe15}. 
We present them in Figure \ref{vlad:holidays}, for multiple kernel SLEMs, ESVM and RESVM-2 presented by \cite{ZePe15}. 
\input{holidaysDistrac.tex}

\section{Recursive square loss exemplar machine?} \label{RSLEM}
One interesting process proposed in \cite{ZePe15} is the recursive exemplar SVM encoding (or RESVM). The proposed idea is to implement the encoding pipeline $k$ times, $k>0$, taking the output of the $(k-1)$-th ESVM pipeline as input of the $k$-th ESVM pipeline. Here, we consider the base features as the output of the of the 0-th ESVM pipeline. 
Indeed, if write $\omega^\star$ as a function of the positive exemplar $x_0$ and the pool of negative examples $\NNN=\{x_1,x_2,...,x_n\}$\footnote{In our submission, we treated the pool of negatives as a matrix $X$. Exceptionally in this section, we treat the pool as a set in order to facilitate our notation}, \emph{i.e.}, 
\begin{equation}
\omega^\star(x_0, \NNN) = \argmin_{\omega\in\RR^d} \left(\theta \ l(1, \omega^Tx_0+\nu^\star) +\dfrac{1}{n}\sum_{x_i\in\NNN}l(1, \omega^Tx_i+\nu^\star)+\dfrac{\lambda}{2}\|\omega\|^2\right),
\end{equation}
and we denote the baseline encodings as $\omega^{(0)}=x_0$ and $\NNN^{(0)}=\NNN$,
the $k$-th recursion of ESVM is given by $\omega^{(k)}(x_0, \NNN) = \omega^\star (\omega^{(k-1)}, \NNN^{(k-1)})$, where $\NNN^{(k)} = \{\omega^\star(z, \NNN^{(k-1)}\setminus \{z\}), \ z\in\NNN^{(k-1)} \}$.
Note that the application of the $k$-th pipeline depends on the application of the $(k-1)$-th pipeline on the negative examples. In \cite{ZePe15}, this means taking the pool of negatives equals to $\NNN \setminus \{x_i\}$ for each negative example $x_i$. However, for SLEM, it means recalculating the matrix $A$ for each different pool of negative examples. This makes a recursive SLEM much more time costly than a regular SLEM, negating the computational advantages of 
our approach. Thus, we have not implemented recursive SLEM. One could 
also imagine keeping the positive example among the negative ones, which 
would avoid recomputing A at each iteration, but does not make much sense. Indeed, it would mean solving a classification problem where the only positive example is also a negative example.

\begin{equation}
    x_0\in \RR^d
\end{equation}
%The other alternative is to not exclude the positive example of the pool of negative, \emph{i.e.}, 
%$\NNN^{(k)} = \{\omega^\star(z, \NNN^{(k-1)}), \ z\in\NNN^{(k-1)} \}$. This avoids the recalculation of matrix $A$, but it means we solve a classification problem where the only positive example is also a negative example for each element in $\NNN$. As a result, in our experiments with recursive SLEM, we obtain a mAP 0.5 percentage points inferior to the regular SLEM results. These results were obtained in Holidays dataset, for Gaussian, linear kernels and $k=2$ recursive encodings.

\bibliographystyle{ieee} 
\bibliography{sup,jz}
\end{document}

%% Local Variables:
%% TeX-master: "suppmat"
%% End:

### 2016-11-07 writeup.tex
first commit

\documentclass[10pt,twocolumn,letterpaper]{article}

\usepackage{cvpr}
\usepackage{times}
\usepackage{epsfig}
\usepackage{graphicx}
\usepackage{caption}
\usepackage{subcaption}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{color}
\newtheorem{proposition}{Proposition}
\newtheorem{lemma}{Lemma}
\newtheorem{proof}{Proof}
\newcommand{\RR}{\mathbb R}
\newcommand{\NNN}{\mathcal N}
\newcommand{\sumi}{\displaystyle{\sum_{i=1}^n}}

% Include other packages here, before hyperref.

% If you comment hyperref and then uncomment it, you should delete
% egpaper.aux before re-running latex.  (Or just hit 'q' on the first latex
% run, let it finish, and you should be clear).
\usepackage[pagebackref=true,breaklinks=true,letterpaper=true,colorlinks,bookmarks=false]{hyperref}

% \cvprfinalcopy % *** Uncomment this line for the final submission

\def\cvprPaperID{****} % *** Enter the CVPR Paper ID here
\def\httilde{\mbox{\tt\raisebox{-.5ex}{\symbol{126}}}}

% Pages are numbered in submission mode, and unnumbered in camera-ready
\ifcvprfinal\pagestyle{empty}\fi
\begin{document}

%%%%%%%%% TITLE
\title{Square Loss Exemplar Machine as Image Encoding: Applications on Image Retrieval}

\author{Rafael Rezende\\
Institution1\\
Institution1 address\\
{\tt\small firstauthor@i1.org}
% For a paper whose authors are all at the same institution,
% omit the following lines up until the closing ``}''.
% Additional authors and addresses can be added with ``\and'',
% just like the second author.
% To save space, use either the email address or home page, not both
\and
Second Author\\
Institution2\\
First line of institution2 address\\
{\tt\small secondauthor@i2.org}
}

\maketitle
%\thispagestyle{empty}

%%%%%%%%% ABSTRACT
\begin{abstract}
   In this paper, we propose an improvement of the image encoder based on the exemplar SVM, first proposed bt Zepeda et al.
   First, we show a more efficient implementation, replacing the hinge loss by the square loss. 
   We obtain superior results on image search in a fraction of the execution time.
   Then, we introduce a low-rank kernel trick for exemplar SVM \emph{\color{red} talk about results} 
\end{abstract}

\input{sec_intro}

\section{Prior work}

\input{sec_SLEM.tex}

\input{sec_kernel_methods.tex}

\input{sec_eff_imp.tex}

\section{Experimental Evaluation}
\subsection{Databases and evaluation protocol} \label{eval:protocol}
Our default database in this report will be INRIA \emph{Holidays} database \cite{holidays}. This database consists of 1491 images divided in 500 groups. For the remaining of this report, if the database of an experiment is not mentioned, the experiment is performed on \emph{Holidays}.
We also perform in the \emph{Oxford} database \cite{oxford}, which consists of 5000 images separated in 55 groups.

Each group of these databases contains one query image. For each query image we rank all the other images in the database based on the similarity between them and the query image, in decreasing order. The average precision of a group is calculated by the ranking of the images of the group for the similarity with the corresponding query image. The final mean average precision (abbreviated for the rest of this report as mAP) for a database is the mean of the average precision over all its groups.
As negative sample of images, we use a small set of images from Flickr100K \cite{oxford}.
\subsection{Base Encoding of Images}
We use three base features as the representation in $\RR^d$ of our images. Firstly, we use VLAD representation, as used in \cite{ZePe15}. 
CNN features are non-negative and can be used both as $\mathbb{L}^1$ or $\mathbb{L}^2$ normalized features. The third features are spatial pyramids of SIFT descriptors, as used in \cite{spk}. These are non-negative $\mathbb{L}^1$ normalized features. \emph{\color{red} Details of these features construction are not relevant right now.}
\subsection{Implementation details}
\emph{\color{red} Not relevant right now.  To be written.}
$8192$ dimensional VLAD features.

E-SVM: $0.6$ second to solve one single exemplar.

SLEM: $30$ seconds to solve a $8192\times 8192$ linear system for all exemplars of the dataset.

Kernelized SLEM: at most $0.3$ seconds (for RBF kernel) for each iteration of (\ref{icd:algo}) algorithm plus at most $30$ seconds to solve a $r'\times r'$ system.

\subsection{Comparison between SLEM and LDA scores}
Let us reconsider the non-kernelized SLEM of subsection \ref{SLEM}. In this case, $A = \frac{1}{n}XX^T-\mu\mu^T+\lambda\textbf{Id}_d$ is the covariance matrix of the negative samples $X$\footnote{A small term proportional to the identity matrix is added to insure the covariance matrix is positive-definite.}. After application of Equation (\ref{beta:fromwood}) to obtain $\omega^\star(x_0,X)$, we can calculate the similarity score $s(x_0,x_0')$ between a query image $x_0$ and an other image $x_0'$ as the inner product of its SLEM vectors:
\begin{equation}
s(x_0,x_0') = (x_0-\mu)^TA^{-2}(x_0'-\mu).
\end{equation}
One interesting fact is that a similar similarity score can be obtained by applying LDA to the positive data. 
Indeed, $\hat{s}(x_0,x_0') = (x_0-\mu)^TA^{-1}(x_0'-\mu)$ is the similarity score for LDA-transformed vectors $x_0$ and $x_0'$. 
\cite{Koba15} noticed that $\hat{s}$ corresponds to the similarity score of SLEM vectors if we use the similarity score proposed in \cite{Efros11} to zero-centered data.

In order to compare both scores, we introduce the similarity score 
\begin{equation}
s_{\alpha}(x_0,x_0') = (x_0-\mu)^TA^{\alpha}(x_0'-\mu),
\end{equation}
and we study how the mAP of \emph{Holidays} evolves when we vary $\alpha$. Notice that $s=s_{-2}$, $\hat{s}=s_{-1}$ and $s_0$ corresponds to zero-centered base features inner product. We also compare both scores to the score obtained by whitening the data (\emph{i.e.}, $\mu$ and $A$ are, respectively, the mean and covariance matrix of the assembled positive and negative data) and by the original base features inner product (\emph{i.e.}, $x_0^Tx_0'$). 
Performing this test on VLAD and CNN feature, as seen on Figure \ref{salpha:vladcnn}, we observe the mAP plateauing for $\alpha$ in $[-2,-1]$, suggesting there is no real improvement in using non-kernelized SLEM over LDA, or, for that matter, the similarity score of \cite{ZePe15}. (which is the most important contribution of this paper) over the similarity of \cite{Efros11}. \emph{\color{red} I'd need to test with E-SVM scores, but if this is true, Patrick's paper is kind of pointless!}
On the other hand, the spatial pyramid features improves its results for smaller values of $\alpha$ in $[-4,0]$, as seen on Figure \ref{salpha:pyr}.
\begin{figure}[!h]
\centering
\begin{subfigure}[b]{0.45\textwidth}
\includegraphics[width=\textwidth]{whitening_cnn.jpg}
\end{subfigure}
\begin{subfigure}[b]{0.45\textwidth}
\includegraphics[width=\textwidth]{whitening_vlad.jpg}
\end{subfigure}
\caption{Comparison of mAP for different similarity scores $s_{\alpha}$, varying $\alpha$. On the top, CNN features. At the bottom, VLAD features. On blue, $A$ is the covariance of the negative samples and on red, $A$ is the covariance of the positive and negative samples. In black, the mAP for similarity calculated as inner product of base features.}
\label{salpha:vladcnn}
\end{figure}

\begin{figure}
\centering
\includegraphics[width=.45\textwidth]{whitening_pyr.jpg}
\caption{Comparison of mAP for different similarity scores $s_{\alpha}$, varying $\alpha$, for spatial pyramid of SIFT descriptors.}
\label{salpha:pyr}
\end{figure}

\begin{table*}[t]
\begin{center}
\begin{tabular}{|l|c|c|c|c|c|}
\hline
Method & $s_0$ & E-SVM \cite{ZePe15} &Linear SLEM & Gaussian SLEM &  Intersection SLEM \\
\hline\hline
VLAD-64 & 73.3 & 77.5 & 77.9 & 78.1 & -\\
Caffe CNN & 69.7 & 71.3 & 72.1 & 72.9 & 70.2\\
Spatial pyramid & 37.4 & 40.1 & 38.2 & 39.7 & 63.5\\
\hline
\end{tabular}
\end{center}
\caption{Results. For all but the second column, we used SLEM. The - means the tests can not be performed or was not performed yet.}
\end{table*}
\subsection{Which kernel to choose?}
For all base features, we test non-kernelized SLEM, as well as kernelized SLEM for Gaussian (rbf):
\begin{align}
k_{rbf}(x,y) &= \exp\left(-\gamma||x-y||^2\right).%;\\
%k_{poly}(x,y) &= x^Ty+\gamma(x^Ty)^2.
\end{align}
For non-negative $\mathbb{L}^1$ normalized, we also test the intersection kernel: 
\begin{equation}
k_{inter}(x,y) = \min(x,y).
\end{equation}
For Gaussian kernel, we can cross validate the parameter $\gamma$. Doing so, we obtained $\gamma$ small ($\gamma\sim 0.5$). Small values of $\gamma$ for the Gaussian kernel mean it is close to the linear kernels, and so its mAP results are not significantly better than the non-kernelized version.

Indeed, if we assume the base features are normalized, i.e. $||x||=||y||=1$, the Taylor series of the exponential function gives us

\begin{align}
k_{rbf}(x,y) &= 1-\gamma||x-y||^2+O(\gamma^2)\\
&=(1-2\gamma)+2\gamma x^Ty+O(\gamma^2)\\
&=C_1+C_2k_{lin}(x,y)+O(\gamma^2).\label{rbf:as:lin}
\end{align}

Equation (\ref{rbf:as:lin}) suggests that for small enough values of $\gamma$, the Gaussian kernel matrix is linearly dependent of the linear kernel matrix. In the context of an image search problem, both kernel matrices give the same result.%\footnote{As explained in section \ref{eval:protocol}, the similarity between positive samples $i$ and $j$ is given by the $j$-th entry of the $i$-th column of the similarity }.

For the spatial pyramid of SIFT, the intersection kernel works much better than all the other kernels. Strangely, this kernel doest not seem to alter the results for CNN features.

\subsection{Low-rank decomposition evaluation}
\emph{\color{red}subsection for the two images on the next page. Important for future work.}
\begin{figure}[!h]
\centering
\includegraphics[width=0.45\textwidth]{None_vs_linear.png}
\caption{Comparison between no kernel results and linear kernel results. In black, mAP of non kernelized SLEM. In red, mAP for different low-rank decompositions of linear kernel matrix. In this experiment, $r=2^{13}$ and we set $\log_2 r'$ in $\{6, 7,...,13\}$. }
\label{no.ker.vs.linear1}
\end{figure}

\begin{figure}[!h]
\centering
\includegraphics[width=0.45\textwidth]{rbf_decomposition2.png}
\caption{Comparison between no kernel results and linear kernel results. In black, mAP of non kernelized SLEM. In red, mAP for different low-rank decompositions of linear kernel matrix. In this experiment, $r=2^{13}$ and we set $\log_2 r'$ in $\{6, 7,...,13\}$. }
\label{no.ker.vs.linear2}
\end{figure}

\begin{figure}[!h]
\centering
\includegraphics[width=0.45\textwidth]{proj_b_0_2.png}
\caption{}
\label{proj}
\end{figure}
\bibliographystyle{ieee} 
\bibliography{sup}
\end{document}

### 2016-11-07 sec_eval_raf.tex
Update sec_eval_raf.tex

@@
 \textbf{Poly SLEM.} The polynomial kernel $k_2$ is a reproducing kernel often used in natural language processing.
-%\textbf{SPM SLEM} The spatial pyramid matching kernel of $L+1$ levels in Equation \ref{k:spm} take as input a set of local descriptors and its location in pyramidal bins \cite{spk}.
 
@@
 These features are obtained by adding a differentiable version of the VLAD algorithm~\cite{Delhumeau2013} as a layer at the end of a convolutional architecture.
-
-%We revisit the VLAD feature presented in \cite{VLAD} as an example of a hand-crafted representation.
-First we extract a set $\mathcal{F}$ of local descriptors of the image $I$.
-We use the 128 dimension RootSIFT \cite{3things} descriptors, extracted densely.
-%Then, we hard-assign each descriptor $f$ in $\mathcal{F}$ to the closest among $K$ pre-trained codewords $c_k$, $k\in\{1\cdots K\}$,
-%one of $K$ set $\mathcal{C}_k$ of descriptors associated to codewords $\{c_k\}_{1\leq k\leq K}$, 
-%and map $f$ to a $\RR^{128K}$ vector
-%\begin{equation}
-%\phi^{VL}_1(f) = \left[0 \cdots 0\quad \Phi_k^T\frac{(f-c_k)}{\|f-c_k\|} \quad 0 \cdots 0\right],
-%\end{equation}
-%where $\Phi_k$ is a $128\times 128$ PCA matrix learned on training features mapped to $k$-the codeword.
-%associated to descriptors in $\mathcal{C}_k$. 
-%The final VLAD representation is the power-normalization and $l_2$ normalization of the sum-pooling of $\phi^{VL}_1$:
-%\begin{equation}
-%\phi^{VL}_2(I) \propto \mathrm{power}\big(\sum_{f\in \mathcal{F}}\phi^{VL}_1(f)\big),~\|\phi^{VL}_2(I)\|=1,
-%\end{equation}
-%with scalar power normalization $\mathrm{power}(v)=\mathrm{sign}(v)|v|^{0.2}$ applied component-wise.
-%In experiments, we use $K=64$ codewords learned on images from Flickr. Our VLAD representation has $d=8192$ dimensions.
-
-%Convolutional features obtained from very deep convolutional neural networks (CNNs) have been shown to work as good local descriptors for matching \cite{SimonZisser15}.
-The SPoC representation \cite{babenko15} is a weighted sum-pooling of the activations of the last convolutional layer of a 19-layer CNN.
-For a given input image $I$, the activations in this layer are organized over a $W\times H$ spatial grid and over $d$ channels.
-Each position $(w,h)$ in $\{1 \cdots W\}\times \{1\cdots H\}$ can thus be equipped with a descriptor $f_{h,w}(I)$ in $\RR^D$.
-%Indeed, if our last convolutional layer has $D$ neurons, and each neuron a $W\times H$ map of activations of this neurons to the image $I$, 
-%each pair $(w,h)$ with $w$ in $\{1,2,..., W\}$ and $h$ in $\{1,2,...,H\}$ can be associated to a descriptor $f_{(h,w)}$ in $\RR^D$ of the responses of each neuron at coordinate $(w,h)$ of the maps. 
-%We then sum-pool these descriptors, weighted accordingly to their distance to the center of the image:
-%\begin{equation}
-%    \phi^{SPoC}(I) = \sum_{w=1}^W\sum_{h=1}^H \alpha_{w,h}f_{w,h}(I),
-%\end{equation}
-%where
-%\begin{equation}
-%    \alpha_{w,h} = \exp \left(-\dfrac{(w-W/2)^2+(h-H/2)^2}{2\sigma^2}\right).
-%\end{equation}
-%In our experiments, we follow the implementation details of \cite{babenko15}: We resize all images to $586\times 586$ pixels before feeding them to the network.
-The last convolutional layer has $d=512$ channels with activation maps of size $(W,H)=(37,37)$, and we set $\sigma=\frac{H}{3}$.
-We use the same network architecture to extract the convolutional maps but different weights in the convolutional layer, which might explain the difference between the results in Table \ref{fullrank:results} and \cite{babenko15}.
-
-%Finally, we also test a more convencional CNN feature, less deep and using two fully connected layers after the convolutional layers.
-%Our final representation is a $4096$ non-negative feature. We based out implementation on CAFFE \cite{jia2014caffe}.
 
@@
 \toprule
-Dataset & \multicolumn{4}{c}{\textbf{Holidays}} & \multicolumn{4}{c}{\textbf{Oxford 5k}} & \multicolumn{2}{c}{\textbf{Oxford 105k}} \\%& \textbf{Oxford 105k}\\
+Dataset & \multicolumn{4}{c}{\textbf{Holidays}} & \multicolumn{4}{c}{\textbf{Oxford 5k}} & \multicolumn{2}{c}{\textbf{Oxford 105k}} \\
 \midrule
-Model, features & VLAD  & SPoC & AlexNet & NetVLAD & VLAD & SPoC & AlexNet & NetVLAD & SPoC & NetVLAD\\% & SPoC\\
+Model, features & VLAD  & SPoC & AlexNet & NetVLAD & VLAD & SPoC & AlexNet & NetVLAD & SPoC & NetVLAD\\
 \midrule
-Baseline            & 72.7 & 76.4 & 68.2  &  85.4 & 46.3 & 54.4 & 40.6 & 67.5 & - & - \\% & 50.1\\
+Baseline            & 72.7 & 76.4 & 68.2  &  85.4 & 46.3 & 54.4 & 40.6 & 67.5 & - & - \\
 %Whitening           & -    & -    & -    & -    & -\\
-PCA+whitening       & 69.4 & 81.6 & 69.2 & 88.3  & 50.9  & 63.7 & 45.0 & 69.1 & - & - \\% & 55.5\\
+PCA+whitening       & 69.4 & 81.6 & 69.2 & 88.3  & 50.9  & 63.7 & 45.0 & 69.1 & - & - \\
 LDA                 & -   &   -   &  -   &   -   &   -   &   -  & 40.5 &  -  & - & - \\
-E-SVM               & 77.5 & 84.0 & 71.3 & 90.5 & 57.2  & 62.1 & 43.9 & 70.4(?) & - & - \\% & 56.5\\ netvlad_hol: 91.6
-Linear SLEM         & 78.0   & 82.3 & 72.1 & 90.9& \ul{59.3}  & 64.1 & 46.9/42.2(?) & 72.4 & - & - \\% & 56.7\\
+E-SVM               & 77.5 & 84.0 & 71.3 & 90.5 & 57.2  & 62.1 & 43.9 & 70.4(?) & - & - \\
+Linear SLEM         & 78.0   & 82.3 & 72.1 & 90.9& \ul{59.3}  & 64.1 & 46.9/42.2(?) & 72.4 & - & - \\
 Gaussian SLEM (16)  & - & - & - & 91.1 & - & - & - & 70.8 & - & - \\
 Gaussian SLEM (32)  & - & 82.2 & - & 91.0 & - & - & - & 71.1 & - & - \\
-Gaussian SLEM (fr)  & 78.1 & 85.2 & \ul{72.9} & 91.4 & 59.0   & \ul{64.9} & 47.0 & \bf{73.2} & - & - \\% & 59.5\\
+Gaussian SLEM (fr)  & 78.1 & 85.2 & \ul{72.9} & 91.4 & 59.0   & \ul{64.9} & 47.0 & \bf{73.2} & - & - \\
 Poly SLEM (16)      & - & 82.3 & - & 91.3 & - & 63.6 & - & 71.2 & - & - \\
 Poly SLEM (32)      & - & 82.4 & - & \bf{91.7} & - & 63.6 & - & 71.7 & - & - \\
-Poly SLEM   (fr)    & 78.1 & \ul{86.3}  & \ul{72.9} & \bf{91.7} & \ul{59.3}  & 64.8 & \ul{47.3} & \bf{73.2} & - & - \\% & \bf{62.3}\\
-%Zepeda \textit{et al.} \cite{ZePe15} & \underline{78.3} & - & 71.8 & - & 57.5 & - & 44.6 & - & - \\
-%Babenko \textit{et al.} \cite{babenko15} & - & 81.8 & - & - & - & \textbf{65.7} & - & - & 57.8\\
+Poly SLEM   (fr)    & 78.1 & \ul{86.3}  & \ul{72.9} & \bf{91.7} & \ul{59.3}  & 64.8 & \ul{47.3} & \bf{73.2} & - & - \\
 \hline

### 2016-11-07 sec_eval_raf.tex
Commenting Li et al.

@@
 Arandjelovi\'c \textit{et al.} \cite{Arandjelovic15}& 4096 & 88.3 & 69.1 \\
-Li \textit{et al.} \cite{LiLaHa15}& - & 89.2^{\dag} & 73.7 \\
+%Li \textit{et al.} \cite{LiLaHa15}& - & 89.2^{\dag} & 73.7 \\
 NetVLAD + Linear SLEM & 4096 & 90.9 & 72.1 \\

### 2016-11-13 conclusion.tex
Update conclusion.tex

@@
 \label{conclusion}
-In this paper, we address the problem of large scale image retrieval with global image representation. 
+In this paper, we address the problem of image retrieval with global image representation. 
 We contribute to the domain by presenting a simple idea, the kernelized square loss exemplar machine, and its efficient implementation.
-As a result, we obtain significant improvements over the image representations we tested and outperform similar encoders on different datasets and on large scale image search. 
-As future work, we must test our model on state of the art representations and see how it impacts their results.
-The use of other kernel functions is also worth investigating.
-The polynomial kernel outperforms the Gaussian kernel, even though the Hilbert space obtained from the Gaussian kernel has infinite dimensions and the Hilbert space obtained from the polynomial kernel does not.
-The spatial pyramid kernel can be used to improve representations based on local descriptors and it offers the better boosts when compared with its baseline.
-Many other performing matching kernel are yet to be tested \cite{Tolias13}, and this work introduces a model that we hope can be reproduced for all kernel functions.
+As a result, we obtain significant improvements over the image representations we tested and outperform similar encoders on different datasets. 
+As future work, we must work on a convolutional implementation similar to \cite{Arandjelovic15} so its parameters can be learned in a supervised manner.
+The use of other kernel functions is worth investigating.
+The polynomial kernel performs similarly to the Gaussian kernel, even though the Hilbert space obtained from the Gaussian kernel has infinite dimensions and the Hilbert space obtained from the polynomial kernel does not.
+Different kernels such as the spatial pyramid kernel \cite{spk} are an option for baseline feature that are not global descriptors, wich would increase the versatility of our approach.
+%The spatial pyramid kernel can be used to improve representations based on local descriptors and it offers the better boosts when compared with its baseline.

### 2016-11-13 main_cvpr_review.tex
Update main_cvpr_review.tex

@@
 \begin{abstract}
-This paper proposes a new feature encoding pipeline, based on the success of the exemplar SVM (ESVM) \cite{ZePe15} and linear discriminant analysis (LDA) as whitening \cite{HMR12}.
-We first show that, by replacing the hinge loss by the square loss in the ESVM cost function, similar results in image retrieval can be obtained at a fraction of the computational cost. 
+This paper proposes an extension of the exemplar SVM (ESVM) as a novel approach to feature encoding for feature retrieval applications.
+%This paper proposes a new feature encoding pipeline, based on the success of the exemplar SVM (ESVM) \cite{ZePe15} and linear discriminant analysis (LDA) as whitening \cite{HMR12}.
+We first show that, by replacing the hinge-loss by the square-loss in the ESVM cost function, equivalent results in image retrieval can be obtained at a fraction of the computational cost. 
 We call this model square-loss exemplar machine, or SLEM. 
-Secondly, we introduce a kernelized SLEM version which benefits from the same computational advantages but may displays improved performance. 
+We also introduce a kernelized SLEM which can be implemented efficiently through a low-rank matrix decomposition and
+%benefits from the same computational advantages but 
+displays improved performance. 
 Both SLEM variants exploit the fact that the negative examples are fixed for all the positives in the training set, so most of the SLEM computational complexity is relegated to an offline process independent of the positive examples.
-Our experiments establish the performance and computational advantages of our methods using a large array of base feature representations and standard image retrieval datasets.
+Our experiments establish the performance and computational advantages of our approach using a large array of base feature representations and standard image retrieval datasets.
 \end{abstract}

### 2016-11-13 sec_eff_imp.tex
Update sec_eff_imp.tex

@@
 \input{figure_LR.tex}
-When compared to the square-loss classifier of section \ref{slem:intro}, one drawback of a kernelized approach is that the dimension of our problem grows with the size $n$ of the negative samples.
-The offline factorization $B$ of $K$ demands $O(nr)$ storage and at best $O(n^2r)$ time.
+When compared to the linear square-loss classifier of Section \ref{slem:intro}, one drawback of a kernelized approach is that the dimension of our problem grows with the size $n$ of the negative samples.
+The offline factorization $B$ of $K$ demands $O(nr)$ storage and at best $O(nr^2)$ time.
 This factorization can be obtained in different ways, depending on the aspect we wish to optimize.
 Efficiency in time and efficiency in storage are the two most common of these aspects.
-In this section we propose two different calculations of $B$, to be implemented depending on which of these two efficiencies we hope to achieve.
+In this section we propose three different decompositions of $K$, to be applied depending if we want to minimize time or storage.
+%on which of these two efficiencies we hope to achieve.
 
@@
 \end{align}
-for all $t<j\le n$.
-There are two alternatives of Cholesky decomposition.
-The incomplete Cholesky decomposition (ICD) stops the algorithm (\ref{icd:algo}) after $r$ steps and has complexity $O(nr^2)$.
-To minimize the residue $\|K-BB^T\|$, the $t$-th pivot is calculated at the end of the $t$-th interaction.
+for all $t<j\le n$. There are two possible choices of Cholesky decomposition. The incomplete Cholesky decomposition (ICD) stops its algorithm after $r$ steps and has complexity $O(nr^2)$. 
+To minimize the residue $tr(K-BB^T)$, the $t$-th pivot is calculated at the end of the $t$-th interaction. 
 In the other hand, the complete Cholesky decomposition (CCD) consists of a full-rank decomposition, \textit{i.e.} the rank $r$ of $B$ is equal to $n$ and it has complexity $O(n^3)$. 
-In this case, the pivots are irrelevant to the final residue (which is zero) and can be assumed to be $i_t=t$.
-The computation of pivots makes CCD more efficient for similar residues.
-We make sure $K$ is positive-definite by adding a $\epsilon$ to its diagonal, where $\epsilon=\min(0,-\lambda_{min})$ and $\lambda_{min}$ is the smallest eigenvalue of $K$.
+In this case, the pivots are irrelevant to the final residue (which is zero) and can be assumed to be $i_t=t$. %The computation of pivots makes CCD more efficient for similar residues.
+We compare CCD and ICD performances in Fig.
+\ref{fig:residue}.
+We see that for small $r$, ICD is indeed faster.
+As we increase the value fo $r$, ICD becomes slower than CCD due to the pivot calculation but its residue decreases.
+We further compare both methods in Section \ref{time-scale}.
+
+We make sure $K$ is positive-definite by adding $\epsilon$ to its diagonal, where $\epsilon=\min(0,-\lambda_{min})$ and $\lambda_{min}$ is the smallest eigenvalue of $K$.
 Therefore, from the equation $BB^T=K+\epsilon\mathrm{Id}_n$, $B$ has rank $n$ and we can apply CCD.
@@
 \subsection{Storage-efficient implementation}\label{low-rank} %/Low-rank decomposition} 
-Let us now consider the problem of improving retrieval results in minimal storage.
-As discussed in section \ref{simi_score}, for each positive exemplar we store its original encoder plus $2r$ dimensions, so we aim to decompose $K$ at a small rank $r$.
+Let us now consider instead the problem of improving storage efficiency.
+As discussed in section \ref{simi_score}, for each positive exemplar we store its original encoder plus a $2r$ vector, so we aim to decompose $K$ at a small rank $r$.
 For small values of $r$, a kernel PCA (KPCA) factorization guarantees a better approximation of $K$ than the Cholesky decomposition at a higher time cost.
@@
 \end{equation}
-where $w_i$ denotes the $i$-th column of $W$ and $d_{ii}$ the $i$-th diagonal element of $D$. Decomposing $K$ from Equation~(\ref{svd}) has complexity $O(n^3)$ in time.
+where $w_i$ denotes the $i$-th column of $W$ and $d_{ii}$ the $i$-th diagonal element of $D$.
+Decomposing $K$ from Equation~(\ref{svd}) has complexity $O(n^2r)$ in time.
+\hlc{Francis, I need a reference for this complexity.}
 
-As for ICD, it can be the fastest decomposition at small rank, but it underperformed KPCA; and for higher ranks, it is the slowest alternative and its performance approximates CCD's performance.
-We therefore concentrate our experiments in the two other methods.
-\hlc{RR: However, ICD is still the only $O(nr^2)$ in time, which could be useful for bigger values of $n$.
+ICD may be the fastest decomposition for small ranks, but it is outperformed by KPCA; for higher ranks, ICD is the slowest alternative and its performance approximates CCD's performance.
+We therefore use KPCA for our storage-efficient.
+%\hlc{RR: However, ICD is still the only $O(nr^2)$ in time, which could be useful for bigger values of $n$.
 I'm considering showing the experiments done in the ECCV supmat to show why ICD is obsolete.}

### 2016-11-13 sec_eval_raf.tex
Update sec_eval_raf.tex

--- 2016-11-07 sec_eval_raf.tex
@@
 \subsection{Datasets and evaluation protocol} \label{eval:protocol}
-We perform our experiments in three standard datasets for image retrieval. 
+We perform our experiments on three standard datasets for image retrieval. 
 \begin{itemize}
@@
 We use the "full" crop, ignoring the region of interest of each image.
-    \item The \emph{Oxford105k} dataset \cite{oxford}, large-scale dataset containing the same images and queries from Oxford5k plus \emph{Flickr100k}, a collection of $1e^5$ distractor images Flickr.
+    \item The \emph{Oxford105k} dataset \cite{oxford}, large-scale dataset containing the same images and queries from Oxford5k plus \emph{Flickr100k}, a collection of $10^5$ distractor images Flickr.
 \end{itemize}
@@
 
-At evaluation time, for a dataset that consists of $p$ images and $q$ query images, we calculate its $p\times q$ \emph{similarity matrix} $S$, where each of its $q$ columns is the matching scores of the query image with all the $p$ images.
+At evaluation time, for a dataset that consists of $p$ images and $q$ query images, we calculate the $p\times q$ \emph{similarity matrix} $S$, where each of its $q$ columns is the matching scores of the query image with all the $p$ images.
 \hlc{RR: maybe take out this paragraph completely.}
 
+\begin{table*}[t!]
+\begin{center}
+\setlength{\tabcolsep}{.2em}
+%\begin{tabular}{|c|c|c|c|c|c|c|c|}
+%\small
+\begin{tabular}{c@{\hskip 1em}cccc@{\hskip 1em}cccc@{\hskip 1em}cc}%@{\hskip 1em}c}
+\toprule
+Dataset & \multicolumn{4}{c}{\textbf{Holidays}} & \multicolumn{4}{c}{\textbf{Oxford 5k}} & \multicolumn{2}{c}{\textbf{Oxford 105k}} \\%& \textbf{Oxford 105k}\\
+\midrule
+Model, features & VLAD  & SPoC & AlexNet & NetVLAD & VLAD & SPoC & AlexNet & NetVLAD & SPoC & NetVLAD\\% & SPoC\\
+\midrule
+Baseline            & 72.7 & 76.5 & 68.2  &  85.4 & 46.3 & 54.4 & 40.6 & 67.5 & 50.1 & - \\% & \\
+%Whitening           & -    & -    & -    & -    & -\\
+PCAW                & 75.5 & 81.7 & 69.2 & 88.3 & 50.9  & 63.7 & 45.0 & 69.1 & 55.5 & - \\% & \\
+LDA                 & 54.7 & 82.2 & 64.1 & 74.3 & 29.6  & 62.2 & 42.5 & 72.7 & 52.4 & - \\
+ESVM~\cite{ZePe15}  & $77.5^3$ & $84.0^3$ & 71.3 & $91.4^2$ & 57.2  & 62.1 & 43.9 & 72.5 & 56.5 & - \\% & \\ 
+Linear SLEM         & $78.0^2$ & 82.3 & 72.1 & $91.3^3$ & \ul{59.3}  & 64.1 & 46.2 & 72.9 & 56.7 & - \\% & \\ /42.2
+Gaussian SLEM (16)  & 76.8 & 80.3 & 71.2 & $91.4^2$ & 52.8 & 63.0 & 43.5 & 71.9 & 55.8 & - \\
+Gaussian SLEM (32)  & 77.4 & 81.7 & $72.0^3$ & $91.4^2$ & 54.9 & 63.1 & 44.0 & 71.1 & 56.0 & - \\
+Gaussian SLEM (fr)  & \bf{78.1} & $86.2^2$ & \bf{72.9} & \bf{91.7} & 59.0   & \ul{64.9} & 47.0 & \ul{74.4} & 59.5 & - \\% & \\ 
+Poly SLEM (16)      & 76.9 & 82.3 & 71.4 & $91.3^3$ & 53.0 & 63.6 & 43.6 & 71.4 & 56.1 & - \\
+Poly SLEM (32)      & 77.3 & 82.4 & $72.1^2$ & \bf{91.7} & 54.9 & 63.6 & 44.1 & 71.6 & 56.3 & - \\
+Poly SLEM   (fr)    & \bf{78.1} & \bf{86.3}  & \bf{72.9} & \bf{91.7} & \ul{59.3}  & 64.8 & \ul{47.3} & \ul{74.1} & 62.5 & - \\
+\hline
+\end{tabular}
+\caption{Mean average precision results for INRIA Holidays and Oxford buildings datasets, expressed as percentages.
+In this table, we present our results for VLAD \cite{Delhumeau2013}, sum-pooling of convolutional features (SPoC) \cite{babenko15}, activation coefficients from the previous-to-last CNN layer (AlexNet) \cite{Krizhevsky2012} and activation of NetVLAD layer~\cite{Arandjelovic15}.
+In parenthesis, the rank of he decomposition ('fr' for full rank decomposition)}
+\end{center}
+\label{fullrank}
+\end{table*}
+
 \subsection{Which kernel to choose?}
-For the experiment of SLEM as a feature encoder, we tested two different kernels --Gaussian and polynomial-- each with a scalar parameter $\gamma$:
+For the experiment of SLEM as a feature encoder, we tested two different kernels -- Gaussian and polynomial -- each with a scalar parameter $\gamma$:
 \begin{align}
@@
 well known reproducing kernel, used for classification with support vector machines.\\
-\textbf{Poly SLEM.} The polynomial kernel $k_2$ is a reproducing kernel often used in natural language processing.
+\textbf{Poly SLEM.} The polynomial kernel $k_2$ is a reproducing kernel often used in natural language processing.\\
+%Note the non-kernel version of SLEM is equivalent to a SLEM for the linear kernel.
+In order to distinguish it from other versions, we will refer to the non-kernel version as \textbf{Linear SLEM} for the remaining of this paper.
 %\textbf{SPM SLEM} The spatial pyramid matching kernel of $L+1$ levels in Equation \ref{k:spm} take as input a set of local descriptors and its location in pyramidal bins \cite{spk}.
@@
 \subsection{Base visual features}
-We test our feature encoder for four different base features $x$ in$\mathbb{R}^d$; the hand-crafted VLAD image representation and three learned features derived from the activation coefficients of deep Convolutional Neural Networks.
+We test our feature encoder for four different base features; the hand-crafted VLAD image representation and three learned features derived from the activation coefficients of deep Convolutional Neural Networks.
 
-We use the same VLAD variant of \cite{Delhumeau2013} used in \cite{ZePe15} that relies on densely-extracted rootSIFT \cite{3things} local descriptors, per-cluster, PCA-based rotations, and root normalization.
+We use the same VLAD variant of \cite{Delhumeau2013} used in \cite{ZePe15} that relies on densely-extracted rootSIFT \cite{3things} local descriptors, per-cluster normalization, PCA-based rotations, and root normalization.
 Like \cite{ZePe15}, we use $64$ clusters, for a final feature of size $8192$.
@@
 
-The Sum-Pooling of Convolutional (SPoC) features \cite{babenko15}, which are tailored specifically for the image retrieval application, consist of spatially-weighted sums of the activations of the last convolutional layer of a 19-layer CNN \cite{SimonZisser15}.
+The SPoC features \cite{babenko15}, which are tailored specifically for the image retrieval application, consist of spatially-weighted sums of the activations of the last convolutional layer of a 19-layer CNN \cite{SimonZisser15}.
 
@@
 \subsection{Image retrieval results}
+We use the base features of the previous subsection as baseline.
+Since \textit{Babenko et al.}~\cite{babenko15} and \textit{Arandjelovi\'c et al.}~\cite{Arandjelovic15} have improved retrieval results by applying PCA followed by whitening to their features, we also apply this post-processing to our base features as second baseline (PCAW), compressing its its dimension to half of the original.
+We then compare the baselines with the original ESVM, LDA and several variants of our approach (SLEM), since all the those methods are based on similar \hlc{???
+(I can't read the last word you wrote.} The results are presented in Table \ref{fullrank}.
 
-We test our method against the original ESVM, PCA and LDA algorithms. The results are presented in Table \ref{fullrank:results}.
-\textit{Babenko et al.}~\cite{babenko15} and \textit{Arandjelovi\'c et al.}~\cite{Arandjelovic15} improve the retrieval results applying a PCA compression followed by whitening. 
-Thus for these features, we considere their version as presented in \cite{Arandjelovic15,babenko15} in the \textit{PCA+whitening} row and the not-compressed, not-rotated and not-whitened version as \textit{Baseline}.
-For VLAD and AlexNet, we considere the version presented in~\cite{ZePe15} as \textit{Baseline} and a rotated, whitened and compressed to half its original dimensions as \textit{PCA+whitening}. 
-As for the version of each feature to which we apply the other algorithms, it changes according to the dataset. 
-For SPoC, we apply LDA, ESVM and SLEM to the baseline, as a replacement of PCA. 
-For NetVLAD, we use the rotated and compressed version of the trained networks correspondent to each dataset as suggested by~\cite{Arandjelovic15}. 
-For Holidays we obtain better results with whitened features and for Oxford5k, non-whitened.
-
-For \textit{LDA}, we follow our description in section \ref{sec:lda} and calculate $\Sigma$ as the convariance of the negative samples.
-%In our experiments, the compression worsens the results when compared with non-compressed. \textit{Delhumeau et al.}~\cite{Delhumeau2013}
-%Hence we compare the improvements of PCA plus whitening, without compression, with our method.
 Linear SLEM performs similarly to ESVM despite being much more time efficient (see discussion in section \ref{time-scale}). 
@@
 We perform both Gaussian SLEM and Polynomial SLEM with two decompositions: one full-rank CCD decomposition signaled by (fr) and two low-rank KPCA decompositions signaled by the rank of the decomposition.
-We train our exemplar classifiers for between $6000$ and $10000$ negative samplers.
-We outperform all methods, for all datasets and all image representations.
-\hlc{RR: Do we?
-I'm putting it here because I'm an optimistic, but we have to wait the results.}
+We train our exemplar classifiers for between $6000$ and $15000$ negative samplers.
+For all the experiments we calibrate the regularization cost $\lambda$, as well as the parameter $\gamma$.
 
-\begin{table*}[h!]
-\begin{center}
-\setlength{\tabcolsep}{.2em}
-%\begin{tabular}{|c|c|c|c|c|c|c|c|}
-%\small
-\begin{tabular}{c@{\hskip 1em}cccc@{\hskip 1em}cccc@{\hskip 1em}cc}%@{\hskip 1em}c}
-\toprule
-Dataset & \multicolumn{4}{c}{\textbf{Holidays}} & \multicolumn{4}{c}{\textbf{Oxford 5k}} & \multicolumn{2}{c}{\textbf{Oxford 105k}} \\%& \textbf{Oxford 105k}\\
-\midrule
-Model, features & VLAD  & SPoC & AlexNet & NetVLAD & VLAD & SPoC & AlexNet & NetVLAD & SPoC & NetVLAD\\% & SPoC\\
-\midrule
-Baseline            & 72.7 & 76.4 & 68.2  &  85.4 & 46.3 & 54.4 & 40.6 & 67.5 & - & - \\% & 50.1\\
-%Whitening           & -    & -    & -    & -    & -\\
-PCA+whitening       & 69.4 & 81.6 & 69.2 & 88.3  & 50.9  & 63.7 & 45.0 & 69.1 & - & - \\% & 55.5\\
-LDA                 & -   &   -   &  -   &   -   &   -   &   -  & 40.5 &  -  & - & - \\
-E-SVM               & 77.5 & 84.0 & 71.3 & 90.5 & 57.2  & 62.1 & 43.9 & 70.4(?) & - & - \\% & 56.5\\ netvlad_hol: 91.6
-Linear SLEM         & 78.0   & 82.3 & 72.1 & 90.9& \ul{59.3}  & 64.1 & 46.9/42.2(?) & 72.4 & - & - \\% & 56.7\\
-Gaussian SLEM (16)  & - & - & - & 91.1 & - & - & - & 70.8 & - & - \\
-Gaussian SLEM (32)  & - & 82.2 & - & 91.0 & - & - & - & 71.1 & - & - \\
-Gaussian SLEM (fr)  & 78.1 & 85.2 & \ul{72.9} & 91.4 & 59.0   & \ul{64.9} & 47.0 & \bf{73.2} & - & - \\% & 59.5\\
-Poly SLEM (16)      & - & 82.3 & - & 91.3 & - & 63.6 & - & 71.2 & - & - \\
-Poly SLEM (32)      & - & 82.4 & - & \bf{91.7} & - & 63.6 & - & 71.7 & - & - \\
-Poly SLEM   (fr)    & 78.1 & \ul{86.3}  & \ul{72.9} & \bf{91.7} & \ul{59.3}  & 64.8 & \ul{47.3} & \bf{73.2} & - & - \\% & \bf{62.3}\\
-%Zepeda \textit{et al.} \cite{ZePe15} & \underline{78.3} & - & 71.8 & - & 57.5 & - & 44.6 & - & - \\
-%Babenko \textit{et al.} \cite{babenko15} & - & 81.8 & - & - & - & \textbf{65.7} & - & - & 57.8\\
-\hline
-\end{tabular}
-\caption{Mean average precision results for INRIA Holidays and Oxford buildings datasets, expressed as percentages.
-In this table, we present our results for VLAD \cite{Delhumeau2013}, sum-pooling of convolutional features (SPoC) \cite{babenko15}, activation coefficients from the previous-to-last CNN layer (AlexNet) \cite{Krizhevsky2012} and activation of NetVLAD layer~\cite{Arandjelovic15}.
-In parenthesis, the rank of he decomposition ('fr' for full rank decomposition)}
-\end{center}
-\label{fullrank:results}
-\end{table*}
+The full-rank variant outperform all methods for all base features, although the gains when compared to linear SLEM are not always significant (\emph{e.g.} for VLAD features).
+It is interesting to notice that LDA, ESVM and Linear SLEM do not seem to outperform one another or the baselines consistently for all features and datasets.
+%For VLAD and AlexNet, we considere the version presented in~\cite{ZePe15} as \textit{Baseline} and a rotated, whitened and compressed to half its original dimensions as \textit{PCA+whitening}. 
+%As for the version of each feature to which we apply the other algorithms, it changes according to the dataset. 
+%For SPoC, we apply LDA, ESVM and SLEM to the baseline, as a replacement of PCA. 
+%For NetVLAD, we use the rotated and compressed version of the trained networks correspondent to each dataset as suggested by~\cite{Arandjelovic15}. 
+%For Holidays we obtain better results with whitened features and for Oxford5k, non-whitened.
 
-For all the experiments we calibrate the regularization cost $\lambda$, as well as the parameter $\gamma$. 
+%For \textit{LDA}, we follow our description in section \ref{sec:lda} and calculate $\Sigma$ as the convariance of the negative samples.
+%In our experiments, the compression worsens the results when compared with non-compressed. \textit{Delhumeau et al.}~\cite{Delhumeau2013}
+%Hence we compare the improvements of PCA plus whitening, without compression, with our method.
+
 %For some experiments, specially using VLAD, the optimal value of $\gamma$ is rather small ($\gamma\sim 0.05$), which suggests the linear kernel is already close to the optimal kernel for these features.
-
-\subsection{Time and storage scalability} \label{time-scale}
-In this section we compare the time efficiency of our method and the E-SVM, as well as discuss which method to use accordingly with the number of positive and negative samples.
-
-In Figure \ref{fullrank:results}, we that the see Linear SLEM efficiency does not change with $n$.
-Indeed, if $d$ is the dimension of the base representation, $A$ is a $d\times d$ matrix for linear SLEM, whereas for a full rank kernel, $A$ is $n\times n$.
-This explains the increasing running time for Gaussian and polynomial kernels: storage and solving a $n\times n$ system does not scale for large number of negative samples.
-Retrieval results for kernelized SLEM in Figure \ref{fullrank:results} suggest we can benefit from larger sets of negative samples, but its improvements are not steadily.
-Indeed, the calculation time for a full rank kernel matrix offline is $O(n^3)$, as it can be seen in Figure \ref{fullrank:results}.
-When in consider only the online procedure of our model, \emph{i.e.} the calculation of $\beta^\star$, our kernelized model similarly to ESVM.
-Therefore, we can process the kernel SLEM for the Gaussian and polynomial kernel similar running time to ESVM if we pre-process our negative samples offline.
-For the low-rank decomposition \hlc{not yet presented in Figure \ref{fullrank:results}}, the offline procedure is $O(n^3)$, similarly to the full rank.
-However this step is much more time consuming for the low-rank, as shown in Figure \ref{fig:residue}.
-We need test different values of $n$, but for fixed rank $r$, we should expect better results for smaller $n$.
-
-\vspace{3 mm}
 
@@
 
+\subsection{Time and storage scalability} \label{time-scale}
+In this section we compare the time efficiency of our method and the E-SVM, as well as discuss which method and decomposition to use accordingly with the number of positive and negative samples.
+
+In Fig. \ref{fullrank:results}, we see that the see Linear SLEM efficiency does not change with $n$.
+Indeed, if $d$ is the dimension of the base representation, $A$ is a $d\times d$ matrix for linear SLEM, whereas for a full rank kernel, $A$ is $n\times n$.
+This explains the increasing running time for Gaussian and polynomial kernels: storage and solving a $n\times n$ system does not scale for large number of negative samples.
+
+Retrieval results for full-rank kernelized SLEM in Fig.
+\ref{fullrank:results} suggest we can benefit from larger sets of negative samples.
+We however limit our full-rank experiments to 15000 negatives samples due to the $O(n^3)$ complexity of the offline step.
+When we consider only the online procedure of our model, \emph{i.e.} the calculation of $\beta^\star$, our kernelized model has a similar time-efficiency to ESVM.
+Therefore, we can process the kernel SLEM for the Gaussian and polynomial kernel similar running time to ESVM if we pre-process our negative samples offline.
+
+For the low-rank decomposition, we present in Fig.
+\ref{kpca:icd} a comparison in average precision between KPCA and ICD decompositions using SPoC on the Holidays dataset.
+The superior results justify our preference for KPCA, despite its time-inefficient offline step.
+The only advantage of ICD over KPCA is its time complexity, linear in the number of negative samples, that allows a bigger number of negative samples.
+In Fig.
+\ref{icd} we show results of ICD for bigger pools of negative samples.
+The results suggest that performance of low-rank SLEM is not sensible to the number of negative samples.
+
+\input{fig_KPCAvsICD.tex}
+
+\input{fig_ICD.tex}
+\vspace{3 mm}
+
 \subsection{Comparation to the state of the art}
-Results shown in Table~\ref{sota:results}. We do not include re-ranking nor query expansion.
+We compare our results to the state of the art for Holidays and Oxford 5k in Table ~\ref{sota}.
+We do not include re-ranking nor query expansion.
+We compare the state of the art global descriptors to both SPoC and NetVLAD features improved by Linear SLEM and low-rank Poly SLEM for rank equals to 16 and 32.
+We perform PCA and whitening to reduce the dimension of both descriptors to 256 and 512, and compare results by bracket of dimension.
+We also add a bracket of the full 4096-dimension NetVLAD for completeness, so we include our best performance.
+Our approach outperforms the state of the art for Holidays, despite not using descriptors as base feature.
 
@@
 Kalantidis  \textit{et al.} \cite{KaMeOs16}   & 256 & 83.1 & 65.4 \\
-SPoC + Linear SLEM & 256 & 82.3 & 64.1 \\
-SPoC + Poly SLEM & 288 & 82.3 & 60.6 \\
-SPoC + Poly SLEM & 320 & 82.4 & 60.7 \\
-NetVLAD + Linear SLEM & 256 & \un{88.5} & - \\
+SPoC + Linear SLEM & 256 & 81.5 & 64.7 \\
+SPoC + Poly SLEM & 288 & 80.1 & 63.6 \\
+SPoC + Poly SLEM & 320 & 81.8 & 63.6 \\
+NetVLAD + Linear SLEM & 256 & \un{88.5} & 65.9 \\
 NetVLAD + Poly SLEM & 288 & 87.7 & 65.5 \\
-NetVLAD + Poly SLEM & 320 & 88.3 & 65.4 \\
+NetVLAD + Poly SLEM & 320 & 88.3 & 65.6 \\
 \midrule
@@
 Gordo \textit{et al.} \cite{GoAlReLa16} & 512 & 89.1^{\dag} & \bf{83.1}^{\dag} \\
-NetVLAD + Linear SLEM & 512 & 89.3 & 68.9 \\
-NetVLAD + Poly SLEM & 544 & 89.9 & 69.0 \\
-NetVLAD + Poly SLEM & 576 & \un{89.9} & 69.2 \\
+SPoC + Linear SLEM & 512 & 82.3 & 64.1 \\
+SPoC + Poly SLEM & 544 & 82.3 & 63.0 \\
+SPoC + Poly SLEM & 576 & 82.4 & 63.1 \\
+NetVLAD + Linear SLEM & 512 & 89.3 & 72.3 \\
+NetVLAD + Poly SLEM & 544 & \un{89.9} & 71.9 \\
+NetVLAD + Poly SLEM & 576 & \un{89.9} & 72.3 \\
 \midrule
 Arandjelovi\'c \textit{et al.} \cite{Arandjelovic15}& 4096 & 88.3 & 69.1 \\
-Li \textit{et al.} \cite{LiLaHa15}& - & 89.2^{\dag} & 73.7 \\
-NetVLAD + Linear SLEM & 4096 & 90.9 & 72.1 \\
+%Li \textit{et al.} \cite{LiLaHa15}& - & 89.2^{\dag} & 73.7 \\
+NetVLAD + Linear SLEM & 4096 & 91.3 & 72.9 \\
 NetVLAD + Poly SLEM & 4128 & 91.3 & 71.2 \\
@@
 \caption{Compared results to state-of-the-art features at similar dimensions, without re-ranking or query augmentation.
-The results using Poly SLEM or Gaussian SLEMa dd 32 or 64 dimensions to the original feature (for $r=16$ or $r=32$, respectively).
+The results using Poly SLEM or Gaussian SLEM add 32 or 64 dimensions to the original feature (for $r=16$ or $r=32$, respectively).
 Underlined results are the best at each dimension bracket and bold results are the general best.
 $\dag$ indicates the previous state-of-the-art.}
+\label{sota}
 \end{center}
-\label{sota:results}
 \end{table}
@@
 
-\subsection{Classification results}
+%\subsection{Classification results}

### 2016-11-13 sec_intro.tex
Update sec_intro.tex

@@
 
-Effective image 
-representations are a crucial component of image retrieval systems 
+An effective image 
+representation is a crucial component of image retrieval system
 where, given some query image,  images must be retrieved from a large 
@@
 
-The success of several existing systems in this challenging application has been enabled by advances in image representation.
+The advances in image representation enabled the success of new systems in this challenging task.
 %An image representation function commonly maps a given input image to a vector in a fixed dimensional space called the image \emph{feature vector}.
-Indeed, once the feature space is given,
-the image retrieval process is reduced to finding the feature vectors closest to the query image under, \textit{e.g.}, the Euclidean distance.
-An adequate image representation must thus be robust to image variations with preferably low computation and storage cost.
-
+Indeed, many successful feature representations for image retrieval rely on unsupervised models such as $K$-means \cite{Delhumeau2013} or Gaussian mixture models \cite{Perronnin2010} and, before the neural networks \textit{renaissance}, these representations would outperform methods that exploit supervised learning of image features directly~\cite{Bilen2015,Rana}.
+Today, with the popularization of convolutional architectures, global image descriptors are often obtained by aggregation and/or pooling the last convolutional layers \cite{babenko15,GoAlReLa16,KaMeOs16} or addition of new differentiable layers to an existing architecture~\cite{Arandjelovic15}.
+%One of the main reasons for this is the lack of adequately large and varied supervised datasets that are expensive to collect.
 %A now pervasive example of an image feature representation is that consisting of the activation coefficients extracted from the previous-to-last layer of Convolutional Neural Networks (CNNs) \cite{Krizhevsky2012}.
 Although CNNs are trained in a fully-supervised manner for the image classification task, they have been shown to transfer well not only to new, unseen classes \cite{Oquaba,Chatfield2014}, but also to alternate tasks such as object detection \cite{Girshick2014} and, importantly, image retrieval \cite{Sharif}.
-
-Many successful feature representations for image retrieval rely on unsupervised models such as $K$-means \cite{Delhumeau2013} or Gaussian mixture models \cite{Perronnin2010}.
-Before the neural networks \textit{renaissance}, these representations would outperform methods that exploit supervised learning of image features directly for the image retrieval task \cite{Bilen2015,Rana}.
-Today, with the popularization of convolutional architectures, many representations consist of aggregation and/or pooling of the last convolutional layer \cite{babenko15,GoAlReLa16,KaMeOs16}.
-Others propose new differentiable layers based on successful retrieval representations to be added at the end an existing architecture \cite{Arandjelovic15}.
-%One of the main reasons for this is the lack of adequately large and varied supervised datasets that are expensive to collect.
 
@@
 
-Zepeda \emph{et al.} \cite{ZePe15} propose to instead treat the weights of the resulting classifier as a new feature vector for the exemplar image.
+Zepeda \emph{et al.} \cite{ZePe15} propose to instead treat the weights of the resulting classifier as a new feature vector for image retrieval.
 %This new ESVM feature is derived from \emph{base} feature representations ({\it e.g.,} CNN activation features) of the exemplar and the negative pool.
@@
 Searching amounts to computing distances between the query and the database ESVM features.
-Note that ESVM features can be derived from arbitrary \emph{base} features ({\it e.g.} CNN activations) of the exemplar and the images in the generic negative pool.
-An interesting interpretation of this approach is that the ESVM features represent what is unique about the exemplar relative to the pool of generic negatives, which is the role of a classifier.
+Note that ESVM features can be derived from arbitrary \emph{base} features ({\it e.g.}, CNN activations) of the exemplar and the images in the generic negative pool.
+%An interesting interpretation of this approach is that the ESVM features represent what is unique about the exemplar relative to the pool of generic negatives, which is the role of a classifier.
 
@@
 In this work we propose using the square loss instead of the hinge loss, in effect converting the ESVM problem into a ridge regression that can be solved in closed form.
-The parameters of the corresponding classifier, which we call the squared loss exemplar machine (SLEM), had been noted to generalize LDA \cite{Koba15} and are shown to form features with comparable performance while drastically reducing the computational cost.
+The parameters of the corresponding classifier, which we call the squared loss exemplar machine (SLEM), had been shown to both generalize LDA \cite{Koba15} and, for the general case of multiple positive samples (LS-SVM)~\cite{lssvm}, form features with comparable binary classification performance while drastically reducing the computational cost.
 
 Since computing the SLEM features requires inverting a large matrix related to the training set's covariance matrix, we propose an efficient way to compute this inverse that exploits the fact that only a single (positive) example changes in the training set when computing SLEM features for different images.
-We show experimentally that our representation can match and even improve upon the performance of ESVM features on two standard datasets using a wide range of base features.
+We show experimentally that our representation matches and even improves upon the performance of ESVM features on two standard datasets using a wide range of base features.
 
-We further introduce a kernel variant of SLEM that enjoys similar computational advantages but improves retrieval performances.
-Further computational efficiency is obtained using a kernel singular values decomposition of the negative pool.
+We also introduce a kernelized variant of SLEM that enjoys similar computational advantages but improves retrieval performances.
+Further computational efficiency is obtained using a kernel principal components analysis (KPCA) decomposition of the negative samples.
 %low rank incomplete Cholesky decomposition of the negative pool kernel matrix.
@@
 %In Section \ref{prior work} we provide an overview of various existing feature representation methods. 
-In section \ref{lsesvm} we first review the original ESVM feature representation method and subsequently introduce the proposed linear SLEM model.
-We then introduce the kernelized SLEM variant in section \ref{nonlinear SLEM}.
-We then present the low-rank approximation of our method that enables efficient implementations in the non-linear case in section \ref{eff_imp}.
-We evaluate our proposed method in image retrieval in section \ref{eval}, and present conclusions in section \ref{conclusion}.
+In Section \ref{lsesvm} we first review the original ESVM feature representation method and subsequently introduce the proposed linear SLEM model.
+We then introduce the kernelized SLEM variant in Section \ref{nonlinear SLEM}.
+We then present the low-rank approximation of our method that enables efficient implementations in the non-linear case in Section \ref{eff_imp}.
+We evaluate our proposed method in image retrieval in Section \ref{eval}, and present conclusions in Section \ref{conclusion}.
 

### 2016-11-13 sec_kernel_methods.tex
Update sec_kernel_methods.tex

@@
 
-\subsection{Introduction to kernel methods}\label{kernel:review}
+\subsection{Kernel methods}\label{kernel:review}
 Let us recall a few basic facts about kernel methods for supervised
@@
 $H$ formed by real functions over some set
-$X$, and denote by $k$ the corresponding reproducing kernel.  We
+$X$, and denote by $k$ and $\varphi$ the corresponding reproducing kernel and feature map (which may not admit a known explicit form) over $X$, respectively. We
 address the following learning problem over $H\times\RR$:
+%\begin{equation}
+%\min_{h\in H,\nu\in\RR}
+%\sum_{i=1}^n l(y_i,h(x_i)+\nu) + \frac{\lambda}{2}\|h\|_H^2,
+%\label{eq:kernel}
+%\end{equation}  
+%By definition of a reproducing kernel,
+%Equation (\ref{eq:kernel}) can be rewritten as
 \begin{equation}
 \min_{h\in H,\nu\in\RR}
-\sum_{i=1}^n l(y_i,h(x_i)+\nu) + \frac{\lambda}{2}\|h\|_H^2,
-\label{eq:kernel}
-\end{equation}  
-where the pairs $(x_i,y_i)$ in $X\times \{-1,1\}$, $i=1\dots n$ are training samples, and $l: \{-1,1\}\times \RR\rightarrow\RR^+$ is some arbitrary loss function. 
-By definition of a reproducing kernel,
-Equation (\ref{eq:kernel}) can be rewritten as
-\begin{equation}
-\min_{h\in H,\nu\in\RR}
-\sum_{i=1}^n l(y_i,\langle \varphi(x_i),h\rangle+\nu) +
+\dfrac{1}{n}\sum_{i=1}^n l(y_i,\langle \varphi(x_i),h\rangle+\nu) +
 \dfrac{\lambda}{2}\|h\|_H^2,
@@
 \end{equation} 
-where $\varphi$ is the {\em feature map} over $X$ associated with the
-kernel $k$ (which may not admit a known explicit form) and $\langle \cdot, \cdot \rangle$ is the inner product of $H$. We dub problems with the general form of (\ref{ker:aff}) {\em affine}
+where the pairs $(x_i,y_i)$ in $X\times \{-1,1\}$, $i=1\dots n$ are training samples, %and $l: \{-1,1\}\times \RR\rightarrow\RR^+$ is some arbitrary loss function. 
+%$\varphi$ is the {\em feature map} over $X$ associated with the kernel $k$ (which may not admit a known explicit form) 
+and $\langle h, h' \rangle$ is the inner product of element $h$ and $h'$ in $H$. We dub problems with the general form of (\ref{ker:aff}) {\em affine}
 supervised learning problems since, given some fixed element $h$ of
@@
 We assume from now on that $l$ is convex and continuous.
-Under this assumption, Equation (\ref{ker:aff}) admits an equivalent formulation
+Under this assumption, Eq.
+(\ref{ker:aff}) admits an equivalent formulation
 \begin{align}
@@
 
-\subsection{Preprocessing of negative samples offline}\label{offline}
-Equation (\ref{beta:first}) has the same structure of Equation (\ref{eq:first}) if we were to consider $\theta=0$ and all labels $y_i=-1$.
-In that case, all actions described in subsections \ref{slem:intro} and \ref{kernel:review} can be done independently of positive samples. 
-They are therefore done offline and their results are stored. In the processing of positive samples, i.e. the online phase, these results are loaded.
-The preprocessing phase consists of the calculation of $B$, $\mu = \frac{1}{n}\sum_{i=1}^n b_i^T$ and $A = \frac{1}{n}B^TB-\mu\mu^T +\lambda\mathrm{Id}_r.$
+Note that Eq.
+(\ref{beta:first}) allows us to write the kernel learning problem (\ref{ker:aff}) as an instance of Eq.
+(\ref{eq:first}) by setting $y_i=-1$ for all but one training sample.
+For our approach, we wish to solve (\ref{beta:first}) for many positive training samples (one at a time) against the same set of negative training samples.
+In the following subsections, we show how to take advantage of the fixed negative samples to efficiently solve (\ref{beta:first}).
+
+\subsection{Offline preprocess of negative samples}\label{offline}
+In order to calculate offline all operations that are dependent only on negative samples, let us suppose that the positive training sample is unknown and $K$ is the kernel matrix of the negative samples $X$.
+The preprocessing phase consists of the calculation of the decomposition $B$ and the constants of Eq.
+(\ref{Wood:omega}): $\mu = \frac{1}{n}\sum_{i=1}^n b_i^T$ and $A = \frac{1}{n}B^TB-\mu\mu^T +\lambda\mathrm{Id}_r.$
+These operations are done offline and their results are stored. In the processing of positive samples, \emph{i.e.} the online phase, these results are loaded.
+%Equation (\ref{beta:first}) has the same structure of Eq. (\ref{eq:first}) if we were to consider $\theta=0$ and all labels $y_i=-1$.
 
@@
 \label{subsec:adding}
-We now reconsider the exemplar problem, with one positive example $x_0$ and $n$ negative examples $X$. Here $K$ is the $n\times n$ kernel matrix of the negative samples $X$. 
+We now wish to write Eq. (\ref{beta:first}) as an exemplar classifier, with one positive example $x_0$ and $n$ negative examples $X$.
 %Let us assume the factor $B$ is given. %The calculation of a kernel Cholesky factorization is solved efficiently by \cite{BaJo05,BaJo02}. 
 %We further analyse how to compute this factorization in subsection \ref{offline}.
-
-We denote by $K'$ the augmented kernel matrix obtained by adding the positive samples $x_0$. Such matrix can be written as
+We denote by $K'$ the augmented kernel matrix obtained by adding the positive samples $x_0$. Such a matrix can be written as
 \begin{equation}
@@
 
-This lemma allows us to add a positive sample to Equation (\ref{beta:first}).
+This lemma allows us to add a positive sample to Eq. (\ref{beta:first}).
 %$\beta^\star(x_0,X), \nu^\star(x_0,X) = \underset{\beta\in\RR^{r+1}, \nu\in\RR}{argmin}J'(\beta, \nu)$, for
-With a positive exemplar, Equation \ref{beta:first} now reads
+With a positive exemplar, Eq. (\ref{beta:first}) now reads
 \begin{align}
- \theta\ l(1, b_0'^T\beta+\nu) +\dfrac{1}{n}\sumi l(-1,b_i'^T\beta+\nu)
+ \dfrac{1}{n}l(1, b_0'^T\beta+\nu) +\dfrac{1}{n}\sumi l(-1,b_i'^T\beta+\nu)
 +\dfrac{\lambda}{2}\|\beta\|^2,\label{beta:final}
@@
 In particular, $b_0'=[u; \ v]$ and, for $i>0$, $b_i'=[0; \ b_i]$.
-Solution $(\beta^\star,\nu^\star)$ in $\RR^{r+1}\times\RR$ can be computed just as before by Equation (\ref{omega:solution}), replacing $x_0$ by $b_0'$, $\mu$ by $\mu' = \frac{1}{n}\sum_{i=1}^n b_i'$ and $X$ by the $(r+1)\times n$ matrix $Q$ of columns $b_1', b_2',...,b_n'$.
+The solution $(\beta^\star,\nu^\star)$ in $\RR^{r+1}\times\RR$ can be computed just as before by Equation (\ref{omega:solution}), replacing $x_0$ by $b_0'$, $\mu$ by $\mu' = \frac{1}{n}\sum_{i=1}^n b_i'$ and $X$ by the $(r+1)\times n$ matrix $Q$ of columns $b_1', b_2',...,b_n'$.
 $\alpha^\star$ is now calculated as
-$\alpha^\star=P'\beta^\star$, where $P' = [u^{-1} \ 0^T; -u^{-1}Pv \ P]$ is the pseudoinverse of $B'^T$. This can be expressed by the linear system
-
-\begin{equation}
-\begin{bmatrix} \alpha_0^\star \\ \hat{\alpha}^\star \end{bmatrix} = \begin{bmatrix} \frac{1}{u} & 0^T \\-\frac{1}{u}Pv & P  \end{bmatrix} \begin{bmatrix}\beta_0^\star \\ \hat{\beta}^\star \end{bmatrix}.\label{alpha:to:beta}
-\end{equation}
+$\alpha^\star=P'\beta^\star$, where $P' = [u^{-1} \ 0^T; -u^{-1}Pv \ P]$ is the pseudoinverse of $B'^T$. 
+%This can be expressed by the linear system
 
@@
 
-Indeed, suppose two image representations $x_0$ and $x_0'$ are given and we wish to calculate the similarity score between its SLEM representations $h$ and $h'$, denoted by $s(h,h')$.
-$h'$ can be written as $h'=\alpha_0'\varphi(x_0')+\sum_{i=1}^n \alpha_i'\varphi (x_i)+\nu'$.
-Using Equation \ref{alpha:to:beta} and ignoring bias $\nu$ and $\nu'$ which have empirically no influence, $s(h,h')$ is given by:
+Indeed, suppose two image descriptors $x_0$ and $x_0'$ are given and we wish to calculate the similarity score between theirs SLEM representations $h$ and $h'$, denoted by $s(h,h')$. 
+We write $h'=\alpha_0'\varphi(x_0')+\sum_{i=1}^n \alpha_i'\varphi (x_i)+\nu'$. 
+Moreover, $\alpha$ can be expressed by the linear system
+\begin{equation}
+\begin{bmatrix} \alpha_0 \\ \hat{\alpha} \end{bmatrix} = \begin{bmatrix} \frac{1}{u} & 0^T \\-\frac{1}{u}Pv & P  \end{bmatrix} \begin{bmatrix}\beta_0 \\ \hat{\beta} \end{bmatrix}.\label{alpha:to:beta}
+\end{equation}
+Using Eq. (\ref{alpha:to:beta}) and ignoring bias $\nu$ and $\nu'$ which have empirically no influence, $s(h,h')$ is given by:
 \begin{equation}
@@
 		& \ \ \ \ \ +\alpha_0\alpha_0'k(x_0,x_0')\\
-		& = \hat{\beta}^T\hat{\beta}'+\frac{\beta_0\beta_0'}{uu'}(k(x_0,x_0')-v^Tv').
+		& = \hat{\beta}^T\hat{\beta}'+\lambda^{-2}(k(x_0,x_0')-v^T).
+		%& = \hat{\beta}^T\hat{\beta}'+\frac{\beta_0\beta_0'}{uu'}(k(x_0,x_0')-v^T).
 \end{split}
@@
 
-We need to store $x_0$, $\hat{\beta}$ and $v$ to calculate its matching score to whichever other image. Each image depends therefore on a vector of $p+2r$ dimensions.
+For a given image whose descriptor is $x_0$, we need to store $x_0$, $\hat{\beta}$ and $v$ to calculate its matching score to whichever other image for SLEM.
+Each image depends therefore on a vector of $p+2r$ dimensions.
 

### 2016-11-13 sec_SLEM.tex
Update sec_SLEM.tex

@@
 \section{Square loss exemplar machine}\label{lsesvm}
-In this section, we revisit the exemplar SVM model presented by \cite{Efros11} as a more general exemplar classifier model.
+In this section, we revisit the exemplar SVM model presented by \cite{Efros11} as an instance of a more general family of classifiers.
 Then, we introduce the square loss exemplar machine (SLEM) by solving a simple variant this model and study its properties.
@@
 $\omega$ in $\RR^d$ and $\nu$ in $\RR$.
-In Equation (\ref{eq:first}), $\lambda$ and $\theta$ are respectively a regularization parameter on $\omega$ and a positive scalar adjusting the weight of the positive exemplar.
+In Eq.
+(\ref{eq:first}), $\lambda$ and $\theta$ are respectively a regularization parameter on $\omega$ and a positive scalar adjusting the weight of the positive exemplar.
 %\footnote{We could also acknowledge a parameter $\theta$ other than $\frac{1}{n}$ as a regularization to the error of each negative example.
@@
 
-The exemplar classifiers of $x_0$ with respect to $X$ are the weights $\omega^\star(x_0,X)$ that minimizes the loss function $J$:
+Given a cost $l$, we define the correspondent \emph{exemplar classifiers} of $x_0$ with respect to $X$ are the weights $\omega^\star(x_0,X)$ that minimizes the loss function $J$:
 \begin{equation}
-\big(\omega^\star(x_0, X), \nu^\star(x_0, X)\big) = \underset{(\omega,\nu)\in\RR^d\times\RR}{\argmin} \ J(\omega, \nu). \label{omega:first}
+\big(\omega^\star, \nu^\star\big) = \underset{(\omega,\nu)\in\RR^d\times\RR}{\argmin} \ J(\omega, \nu). \footnote{Depending on the loss function $l$, $\nu^\star(x_0,X)$ may be not unique.}
+%\big(\omega^\star(x_0, X), \nu^\star(x_0, X)\big) = \underset{(\omega,\nu)\in\RR^d\times\RR}{\argmin} \ J(\omega, \nu). 
+\label{omega:first}
 \end{equation}
-To shorten the notation, we refer to these solutions as $\omega^\star$ and $\nu^\star$\footnote{Depending on the loss function $l$, $\nu^\star(x_0,X)$ may be not unique.} when their arguments are implicit.
+%To shorten the notation, we refer to these solutions as $\omega^\star$ and $\nu^\star$ when their arguments are implicit.
 
@@
 \RAF{Maybe this should be in the introduction instead.}
-Applications of ESVM use the hinge loss function, which guarantees that $J$ is a convex function.
-The solution of Equation (\ref{omega:first}) can be thus found by stochastic gradient descent~\cite{bottou10} individually for each positive sample.
+The exemplar SVM~\cite{Efros11,Efros12} is an instance of this model where $l$ is the hinge-loss. 
+%Applications of ESVM use the hinge loss function, which guarantees that $J$ is a convex function. 
+The solution of Eq.
+(\ref{omega:first}) can be thus found by stochastic gradient descent~\cite{bottou10} individually for each positive sample.
 The next subsection show how to calculate all exemplar classifiers simultaneously by changing the loss function.
@@
 Equation (\ref{omega:solution}) shows how to solve (\ref{eq:first}) simultaneously for all exemplars.
+\cite{lssvm}
 %Replacing the hinge loss by the square loss offers a more compact solution, but not necessarily more efficient.
@@
 \textbf{Woodbury identity.} 
-We can modify Equation \ref{omega:solution} by simplifing Equation \ref{eq:U}.
+We can simplify Eq. (\ref{omega:solution}) by modifying $U$ in Eq. (\ref{eq:U}).
 Let us define $A = \frac{1}{n}XX^T-\mu\mu^T +\lambda\mathrm{Id}_d$ our regularized covariance matrix and assume its inverse $A^{-1}$ known. 
@@
 The matrix $U$ now reads $U = A + \frac{\theta}{\theta+1}\delta\delta^T$, where $\delta=x_0-\mu$ is the centralized (w.r.t.
-the negative mean) positive sample.
+the negatives' mean) positive sample.
 The Woodbury identity \cite{woodbury} gives us
@@
 This means that if search and ranking are based on the normalized feature $\frac{1}{\|\omega^\star\|}\omega^\star$, $\theta$ does not influence the matching score of the SLEM vectors of two different images.
-This sets SLEM appart from ESVM that requires this parameter to be calibrated \cite{Efros11,ZePe15}.
-The normalized exemplar classifier $\omega^\star$ is therefore a linear transformation of $x_0-\mu$
+This sets SLEM appart from ESVM which requires this parameter to be calibrated \cite{Efros11,ZePe15}.
+We can thus set $\theta=\frac{1}{n}$ for the remaining of this work.
+%The normalized exemplar classifier $\omega^\star$ is therefore a linear transformation of $x_0-\mu$.
 %Indeed, if $\omega$ and $\omega'$ are the $d$-dimensional SLEM vectors of $x$ and $x'$, respectively, their cosine similarity reads:
@@
 
-Let us now reanalyse Equation \ref{eq:first} for a particular case.
-Suppose that we have multiple positive samples.
-It can be shown that in this case, the corresponding linear classifier of Equation (\ref{eq:first}) for the square-loss is also given by
+It is interesting to note the relationship between SLEM and the classical linear discriminant analisys (LDA).
+Let us reanalyse Eq.
+(\ref{eq:first})) an suppose that we have multiple positive samples.
+It can be shown that in this case, the corresponding linear classifier of Eq.
+(\ref{eq:first}) for the square-loss is also given by
 (\ref{omega:solution}), where $x_0$ denotes this time the center of mass
-of the positive samples, {\em if} these samples have the {\em same} covariance matrix $\Sigma$ as the negative samples $X$.
+of the positive samples {\em if} these samples have the {\em same} covariance matrix $\Sigma$ as the negative samples $X$.
 
@@
 
-Note that SLEM is a generalized version of LDA.
-Indeed, when $\lambda=0$ in SLEM (i.e.
-no regularization), $A = \Sigma$, meaning $\omega_{LDA}$ of Eq.
+This shows that SLEM is a generalized version of LDA: Indeed, when $\lambda=0$ in SLEM (i.e.
+no regularization), $A = \Sigma$, the vectors $\omega_{LDA}$ of Eq.
 (\ref{eq:lda}) and $\omega^\star$ of Eq.
@@
 %This observation has been equally made by \cite{Koba15}.
-
 Many interesting properties of LDA has been used recently for classification tasks \cite{GMPD12,HMR12} and, more recently, for image retrieval with convolutional neural network activations \cite{babenko15}.
-By simply generalizing LDA, we hope to obtain superior results.
+With our simple generalization of LDA, we hope to obtain superior results.
 

### 2016-11-13 main_cvpr_review.tex
up to section 4.

@@
 %%%%%%%%% TITLE
-\title{Kernel Square-Loss Exemplar Machines}
+\title{Kernel Square Loss Exemplar Machines}
 
@@
 \begin{abstract}
-This paper proposes an extension of the exemplar SVM (ESVM) as a novel approach to feature encoding for feature retrieval applications.
+This paper proposes an extension of the exemplar SVM (ESVM) as a novel approach to feature encoding for retrieval applications.
 %This paper proposes a new feature encoding pipeline, based on the success of the exemplar SVM (ESVM) \cite{ZePe15} and linear discriminant analysis (LDA) as whitening \cite{HMR12}.
-We first show that, by replacing the hinge-loss by the square-loss in the ESVM cost function, equivalent results in image retrieval can be obtained at a fraction of the computational cost. 
-We call this model square-loss exemplar machine, or SLEM. 
+We first show that, by replacing the hinge loss by the square loss in the ESVM cost function, equivalent results in image retrieval can be obtained at a fraction of the computational cost. 
+We call this model square loss exemplar machine, or SLEM. 
 We also introduce a kernelized SLEM which can be implemented efficiently through a low-rank matrix decomposition and
@@
 Both SLEM variants exploit the fact that the negative examples are fixed for all the positives in the training set, so most of the SLEM computational complexity is relegated to an offline process independent of the positive examples.
-Our experiments establish the performance and computational advantages of our approach using a large array of base feature representations and standard image retrieval datasets.
+Our experiments establish the performance \PP{could be a stronger comment, w.r.t.
+SoA?} and computational advantages of our approach using a large array of base feature representations and standard image retrieval datasets.
 \end{abstract}

### 2016-11-13 sec_SLEM.tex
up to section 4.

@@
 \section{Square loss exemplar machine}\label{lsesvm}
-In this section, we revisit the exemplar SVM model presented by \cite{Efros11} as an instance of a more general family of classifiers.
-Then, we introduce the square loss exemplar machine (SLEM) by solving a simple variant this model and study its properties.
+In this section, we revisit the exemplar SVM model proposed in \cite{Efros11} as an instance of a more general family of classifiers.
+Then, we introduce the square loss exemplar machine (SLEM) by solving a simple variant of this model and study its properties.
 \subsection{Exemplar classifiers} \label{esvm}
@@
 \begin{equation}
-\big(\omega^\star, \nu^\star\big) = \underset{(\omega,\nu)\in\RR^d\times\RR}{\argmin} \ J(\omega, \nu). \footnote{Depending on the loss function $l$, $\nu^\star(x_0,X)$ may be not unique.}
+\big(\omega^\star, \nu^\star\big) = \underset{(\omega,\nu)\in\RR^d\times\RR}{\argmin} \ J(\omega, \nu).
+\footnote{Depending on the loss function $l$, $\nu^\star(x_0,X)$ may be not unique.
+\PP{But, what about $w^{\star}$?}}
 %\big(\omega^\star(x_0, X), \nu^\star(x_0, X)\big) = \underset{(\omega,\nu)\in\RR^d\times\RR}{\argmin} \ J(\omega, \nu). 
@@
 \RAF{Maybe this should be in the introduction instead.}
-The exemplar SVM~\cite{Efros11,Efros12} is an instance of this model where $l$ is the hinge-loss. 
+The exemplar SVM~\cite{Efros11,Efros12} is an instance of this model where $l$ is the hinge loss. 
 %Applications of ESVM use the hinge loss function, which guarantees that $J$ is a convex function. 
-The solution of Eq.
-(\ref{omega:first}) can be thus found by stochastic gradient descent~\cite{bottou10} individually for each positive sample.
+The solution of (\ref{omega:first}) can be thus found by stochastic gradient descent~\cite{bottou10} individually for each positive sample.
 The next subsection show how to calculate all exemplar classifiers simultaneously by changing the loss function.
@@
 \subsection{Square loss function}\label{slem:intro}
-Now, let us study the same learning problem for the square-loss function $l(y,\hat{y}) = \frac{1}{2}(y-\hat{y})^2$. As in the case of the hinge loss, Equation (\ref{eq:first}) is a convex problem. 
+Now, let us study the same learning problem for the square loss function $l(y,\hat{y}) = \frac{1}{2}(y-\hat{y})^2$. As in the case of the hinge loss, Equation (\ref{eq:first}) is a convex problem. 
 However it is now a ridge regression problem, whose unique solution can be found in closed form as
@@
 \end{align}
-Equation (\ref{omega:solution}) shows how to solve (\ref{eq:first}) simultaneously for all exemplars.
+Equation (\ref{omega:solution}) shows how to solve (\ref{eq:first}) simultaneously for all individual exemplars.
 \cite{lssvm}
@@
 Note that the positive sample weight $\theta$ does not influence the direction of the optimal vector $\omega^\star$, only its norm.
-This means that if search and ranking are based on the normalized feature $\frac{1}{\|\omega^\star\|}\omega^\star$, $\theta$ does not influence the matching score of the SLEM vectors of two different images.
+This means that if search and ranking are based on the normalized feature $\frac{1}{\|\omega^\star\|}\omega^\star$, \eg using cosine similarity, $\theta$ does not influence the matching score of the SLEM vectors of two different images.
 This sets SLEM appart from ESVM which requires this parameter to be calibrated \cite{Efros11,ZePe15}.
@@
 It is interesting to note the relationship between SLEM and the classical linear discriminant analisys (LDA).
-Let us reanalyse Eq.
-(\ref{eq:first})) an suppose that we have multiple positive samples.
+Let us return to Eq.
+(\ref{eq:first}) an suppose that we have multiple positive samples.
 It can be shown that in this case, the corresponding linear classifier of Eq.
-(\ref{eq:first}) for the square-loss is also given by
+(\ref{eq:first}) for the square loss is also given by
 (\ref{omega:solution}), where $x_0$ denotes this time the center of mass
@@
 \displaystyle \omega_{LDA}=\Sigma^{-1}(x_0-\mu),\\
-\displaystyle \nu_{LDA}=-\frac{1}{2}(x_0+\mu)^T \omega_{LDA},
+\displaystyle \nu_{LDA}=-\frac{1}{2}(x_0+\mu)^T \omega_{LDA}.
 \end{array}\right.
@@
 
-This shows that SLEM is a generalized version of LDA: Indeed, when $\lambda=0$ in SLEM (i.e.
-no regularization), $A = \Sigma$, the vectors $\omega_{LDA}$ of Eq.
+This shows that SLEM is a generalized version of LDA: Indeed, when $\lambda=0$ in SLEM (\ie, no regularization), $A = \Sigma$, the vectors $\omega_{LDA}$ of Eq.
 (\ref{eq:lda}) and $\omega^\star$ of Eq.
@@
 %This observation has been equally made by \cite{Koba15}.
-Many interesting properties of LDA has been used recently for classification tasks \cite{GMPD12,HMR12} and, more recently, for image retrieval with convolutional neural network activations \cite{babenko15}.
+Many interesting properties of LDA have been used recently for classification tasks \cite{GMPD12,HMR12} and, more recently, for image retrieval with convolutional neural network activations \cite{babenko15}.
 With our simple generalization of LDA, we hope to obtain superior results.

### 2016-11-13 sec_eff_imp.tex
up to section 4.

@@
 Efficiency in time and efficiency in storage are the two most common of these aspects.
-In this section we propose three different decompositions of $K$, to be applied depending if we want to minimize time or storage.
+In this section we propose three different decompositions of $K$, to be applied depending on whether we want to minimize time or storage.
 %on which of these two efficiencies we hope to achieve.
@@
 for all $t<j\le n$. There are two possible choices of Cholesky decomposition. The incomplete Cholesky decomposition (ICD) stops its algorithm after $r$ steps and has complexity $O(nr^2)$. 
-To minimize the residue $tr(K-BB^T)$, the $t$-th pivot is calculated at the end of the $t$-th interaction. 
+To minimize the residue $\mathrm{tr}(K-BB^T)$, the $t$-th pivot is calculated at the end of the $t$-th interaction. 
 In the other hand, the complete Cholesky decomposition (CCD) consists of a full-rank decomposition, \textit{i.e.} the rank $r$ of $B$ is equal to $n$ and it has complexity $O(n^3)$. 
@@
 Let us now consider instead the problem of improving storage efficiency.
-As discussed in section \ref{simi_score}, for each positive exemplar we store its original encoder plus a $2r$ vector, so we aim to decompose $K$ at a small rank $r$.
+As discussed in section \ref{simi_score}, for each positive exemplar we store its original encoding plus a $2r$ vector, so we aim to decompose $K$ at a small rank $r$.
 For small values of $r$, a kernel PCA (KPCA) factorization guarantees a better approximation of $K$ than the Cholesky decomposition at a higher time cost.
 Indeed, as illustrated in Figure \ref{fig:residue}, a kernel PCA gives smaller residues than ICD at small rank.
-We plot the ratio $tr(K-BB^T)/tr(K)$ when we vary the number of columns $r$ of $B$, and we see a faster convergence for the kernel PCA decomposition when $r$ goes to $n$.
+We plot the ratio $\mathrm{tr}(K-BB^T)/\mathrm{tr}(K)$ when we vary the number of columns $r$ of $B$, and we see a faster convergence for the kernel PCA decomposition when $r$ goes to $n$.
 However, at these small values of $r$, KPCA is more time consuming than both ICD and CCD.
-This time cost comes from applying a singular values decomposition of K to obtain $B$:
+This time cost comes from applying a singular values decomposition of $K$ to obtain $B$:
 \begin{equation}
@@
 ICD may be the fastest decomposition for small ranks, but it is outperformed by KPCA; for higher ranks, ICD is the slowest alternative and its performance approximates CCD's performance.
-We therefore use KPCA for our storage-efficient.
+We therefore use KPCA for our storage-efficient kernel SLEM.
 %\hlc{RR: However, ICD is still the only $O(nr^2)$ in time, which could be useful for bigger values of $n$.

### 2016-11-13 sec_intro.tex
up to section 4.

@@
 An effective image 
-representation is a crucial component of image retrieval system
+representation is a crucial component of image retrieval systems
 where, given some query image,  images must be retrieved from a large 
@@
 Indeed, many successful feature representations for image retrieval rely on unsupervised models such as $K$-means \cite{Delhumeau2013} or Gaussian mixture models \cite{Perronnin2010} and, before the neural networks \textit{renaissance}, these representations would outperform methods that exploit supervised learning of image features directly~\cite{Bilen2015,Rana}.
-Today, with the popularization of convolutional architectures, global image descriptors are often obtained by aggregation and/or pooling the last convolutional layers \cite{babenko15,GoAlReLa16,KaMeOs16} or addition of new differentiable layers to an existing architecture~\cite{Arandjelovic15}.
+Today, with the popularization of convolutional architectures, global image descriptors are often obtained by aggregation and/or pooling the last convolutional layers \cite{babenko15,GoAlReLa16,KaMeOs16} or by addition of new differentiable layers to an existing architecture~\cite{Arandjelovic15}.
 %One of the main reasons for this is the lack of adequately large and varied supervised datasets that are expensive to collect.
@@
 It uses a large generic pool of images as a set of negative examples, while using a single image (the \emph{exemplar}) as a positive example.
-Given these training sets, a classifier is learned that can  generalize well, despite the drastically limited size of the set of positive examples.
+Given these training sets, an SVM classifier is learned that can  generalize well, despite the drastically limited size of the set of positive examples.
 
-Zepeda \emph{et al.} \cite{ZePe15} propose to instead treat the weights of the resulting classifier as a new feature vector for image retrieval.
+Zepeda \emph{et al.} \cite{ZePe15} propose to treat instead the weights of the resulting classifier as a new feature vector for image retrieval.
 %This new ESVM feature is derived from \emph{base} feature representations ({\it e.g.,} CNN activation features) of the exemplar and the negative pool.
@@
 
-One important drawback of the ESVM feature encoding approach is that computing the  classifier requires solving an optimization problem for each positive example.
+One important drawback of the ESVM feature encoding approach is that computing the  classifier requires solving an optimization problem for each positive example (\ie each query and each database image).
 This can be time consuming for the large negative pool sizes required for good ESVM feature performance.
 In this work we propose using the square loss instead of the hinge loss, in effect converting the ESVM problem into a ridge regression that can be solved in closed form.
-The parameters of the corresponding classifier, which we call the squared loss exemplar machine (SLEM), had been shown to both generalize LDA \cite{Koba15} and, for the general case of multiple positive samples (LS-SVM)~\cite{lssvm}, form features with comparable binary classification performance while drastically reducing the computational cost.
+The parameters of the corresponding classifier, which we call the square loss exemplar machine (SLEM), had been shown to both generalize LDA \cite{Koba15} and, for the general case of multiple positive samples (LS-SVM)~\cite{lssvm}, form features with comparable binary classification performance while drastically reducing the computational cost.
 

### 2016-11-13 sec_kernel_methods.tex
up to section 4.

@@
 where $b_i^T$ denotes the $i$-th row of $B$ and $r$ is the number of columns of $B$. 
-If $(\beta^\star, \nu^\star)$ is the solution Equation (\ref{beta:first}), the corresponding vector $\alpha^\star$ (or, more correctly, \textit{a} corresponding vector of dimension $n\geq r$) 
+If $(\beta^\star, \nu^\star)$ is the solution of (\ref{beta:first}), the corresponding vector $\alpha^\star$ (or, more correctly, \textit{a} corresponding vector of dimension $n\geq r$) 
 can be computed by $\alpha^\star=P\beta^\star$, where $P$ is the pseudoinverse of $B^T$.
@@
 \end{bmatrix},~
-v = B^\dagger k_0,~ u=\sqrt{k_{00}-||v||^2}.
+v = B^\dagger k_0,~ u=\sqrt{k_{00}-||v||^2},
 \label{eq:lemma1}
 \end{equation}
+where $B^\dagger$ is the pseudoinverse of $B$.
 \end{lemma}\label{lemma1}
@@
 %\footnote{
-Indeed, since the matrix $k_{00}K-k_0k_0^T=B(k_{00}\mathrm{Id}_r-vv^T$ is also positive semidefinite. Thus $v^T(k_{00}\mathrm{Id}_r-vv^T)v = \|v\|^2(k_{00}-\|v\|^2)\ge 0$. \qedhere
+Indeed, since the matrix $k_{00}K-k_0k_0^T=B(k_{00}\mathrm{Id}_r-vv^T)$ is also positive semidefinite. Thus $v^T(k_{00}\mathrm{Id}_r-vv^T)v = \|v\|^2(k_{00}-\|v\|^2)\ge 0$. \qedhere
 %}
@@
 The solution $(\beta^\star,\nu^\star)$ in $\RR^{r+1}\times\RR$ can be computed just as before by Equation (\ref{omega:solution}), replacing $x_0$ by $b_0'$, $\mu$ by $\mu' = \frac{1}{n}\sum_{i=1}^n b_i'$ and $X$ by the $(r+1)\times n$ matrix $Q$ of columns $b_1', b_2',...,b_n'$.
-$\alpha^\star$ is now calculated as
+Solution $\alpha^\star$ is now calculated as
 $\alpha^\star=P'\beta^\star$, where $P' = [u^{-1} \ 0^T; -u^{-1}Pv \ P]$ is the pseudoinverse of $B'^T$. 
@@
 For a given image whose descriptor is $x_0$, we need to store $x_0$, $\hat{\beta}$ and $v$ to calculate its matching score to whichever other image for SLEM.
-Each image depends therefore on a vector of $p+2r$ dimensions.
+Each image depends therefore on a vector of $p+2r$ dimensions \PP{Remind what are $p$ and $r$, I lost the former!}.
 

### 2016-11-13 sec_eval_raf.tex
sections 5 and 6

@@
 \subsection{Datasets and evaluation protocol} \label{eval:protocol}
-We perform our experiments on three standard datasets for image retrieval. 
+We perform our experiments on three standard datasets for image retrieval: 
 \begin{itemize}
-    \item The INRIA \emph{Holidays} dataset \cite{holidays} consists of 1491 images divided in 500 groups of matching images.
-We manually rotate by 90 degrees some images that are not in their natural orientation to compensate for the fact that CNN features are not rotation invariant~\cite{babenko14,Arandjelovic15,RaToCh16}
+    \item The INRIA \emph{Holidays} dataset \cite{holidays} which consists of 1491 images divided in 500 groups of matching images.
+We manually rotate by 90 degrees some images that are not in their natural orientation to compensate for the fact that CNN features are not rotation invariant~\cite{babenko14,Arandjelovic15,RaToCh16}.
     \item The \emph{Oxford5k} dataset \cite{oxford}, which consists of 5063 images separated in 55 groups of matching images, each group associated to a landmark of Oxford.
-We use the "full" crop, ignoring the region of interest of each image.
-    \item The \emph{Oxford105k} dataset \cite{oxford}, large-scale dataset containing the same images and queries from Oxford5k plus \emph{Flickr100k}, a collection of $10^5$ distractor images Flickr.
+We use the ``full'' crop, ignoring the region of interest of each image.
+    \item The \emph{Oxford105k} dataset \cite{oxford}, a large-scale dataset containing the same images and queries from Oxford5k plus \emph{Flickr100k}, a collection of $10^5$ distractor Flickr images.
 \end{itemize}
@@
 When training on Oxford105k, we us instead the \emph{Paris} dataset \cite{PhiChIsSiZi08} as negative samples.
+\PP{Not clear: negatives are not all images from Flicker100k or Paris.
+$n$ random samples?}
 %In both datasets, each group contains one query image, the other images in the group being the only correct answers to the query.
@@
 In this table, we present our results for VLAD \cite{Delhumeau2013}, sum-pooling of convolutional features (SPoC) \cite{babenko15}, activation coefficients from the previous-to-last CNN layer (AlexNet) \cite{Krizhevsky2012} and activation of NetVLAD layer~\cite{Arandjelovic15}.
-In parenthesis, the rank of he decomposition ('fr' for full rank decomposition)}
+In parenthesis, the rank of he decomposition (`fr' for full rank decomposition) \PP{meaning of bold and underline}}
+\label{fullrank}
 \end{center}
-\label{fullrank}
 \end{table*}
@@
 \subsection{Base visual features}
-We test our feature encoder for four different base features; the hand-crafted VLAD image representation and three learned features derived from the activation coefficients of deep Convolutional Neural Networks.
+We test our feature encoder for four different base features: the hand-crafted VLAD image representation and three learned features derived from the activation coefficients of deep Convolutional Neural Networks.
 
@@
 We use the base features of the previous subsection as baseline.
-Since \textit{Babenko et al.}~\cite{babenko15} and \textit{Arandjelovi\'c et al.}~\cite{Arandjelovic15} have improved retrieval results by applying PCA followed by whitening to their features, we also apply this post-processing to our base features as second baseline (PCAW), compressing its its dimension to half of the original.
+Since Babenko \etal~\cite{babenko15} and Arandjelovi\'c \etal~\cite{Arandjelovic15} have improved retrieval results by applying PCA followed by whitening to their features, we also apply this post-processing to our base features as second baseline (PCAW), compressing base feature dimension to half of the original.
 We then compare the baselines with the original ESVM, LDA and several variants of our approach (SLEM), since all the those methods are based on similar \hlc{???
 (I can't read the last word you wrote.} The results are presented in Table \ref{fullrank}.
+\PP{Explain why fewer columns for Oxford105k}
 
-Linear SLEM performs similarly to ESVM despite being much more time efficient (see discussion in section \ref{time-scale}). 
-The fact that a hinge-loss classifier does not outperform a square-loss classifier can seem counter-intuitive, but both were shown to be equivalent for binary classification under mild constraints~\cite{YeXi07}.
+Linear SLEM performs similarly to ESVM despite being much more time efficient (see discussion in Section \ref{time-scale}). 
+The fact that a hinge loss classifier does not outperform a square loss classifier can seem counter-intuitive, but both were shown to be equivalent for binary classification under mild constraints~\cite{YeXi07}.
 
-We perform both Gaussian SLEM and Polynomial SLEM with two decompositions: one full-rank CCD decomposition signaled by (fr) and two low-rank KPCA decompositions signaled by the rank of the decomposition.
+We perform both Gaussian SLEM and Polynomial SLEM with two decompositions: one full-rank CCD decomposition signaled by ``(fr)'' and two low-rank KPCA decompositions signaled by the rank of the decomposition.
 We train our exemplar classifiers for between $6000$ and $15000$ negative samplers.
@@
 
-The full-rank variant outperform all methods for all base features, although the gains when compared to linear SLEM are not always significant (\emph{e.g.} for VLAD features).
+The full-rank variant outperforms all methods for all base features, although the gains when compared to linear SLEM are not always significant (\emph{e.g.} for VLAD features).
 It is interesting to notice that LDA, ESVM and Linear SLEM do not seem to outperform one another or the baselines consistently for all features and datasets.
@@
 \subsection{Time and storage scalability} \label{time-scale}
-In this section we compare the time efficiency of our method and the E-SVM, as well as discuss which method and decomposition to use accordingly with the number of positive and negative samples.
+In this section we compare the time efficiency of our method and the ESVM, as well as discuss which method and decomposition to use accordingly with the number of positive and negative samples.
 
-In Fig. \ref{fullrank:results}, we see that the see Linear SLEM efficiency does not change with $n$.
+In Fig. \ref{fullrank:results}, we see that the Linear SLEM efficiency does not change with $n$.
 Indeed, if $d$ is the dimension of the base representation, $A$ is a $d\times d$ matrix for linear SLEM, whereas for a full rank kernel, $A$ is $n\times n$.
@@
 When we consider only the online procedure of our model, \emph{i.e.} the calculation of $\beta^\star$, our kernelized model has a similar time-efficiency to ESVM.
-Therefore, we can process the kernel SLEM for the Gaussian and polynomial kernel similar running time to ESVM if we pre-process our negative samples offline.
+Therefore, we can process the kernel SLEM for the Gaussian and polynomial kernels with similar running time to ESVM if we pre-process our negative samples offline.
 
@@
 Our approach outperforms the state of the art for Holidays, despite not using descriptors as base feature.
+\PP{A word on Ox5k, where we are far behind Gordo?
+Something special about this dataset?}
 

### 2016-11-15 main_cvpr_review.tex
sections 5 and 6
version J-1

--- 2016-11-13 main_cvpr_review.tex
@@
 \newcommand{\un}[1]{\underline{#1}}
+\def\dul#1{\underline{\underline{#1}}}
 
@@
 \begin{abstract}
-This paper proposes an extension of the exemplar SVM (ESVM) as a novel approach to feature encoding for feature retrieval applications.
+Zepeda and P\'erez \cite{ZePe15} have recently demonstrated the promise of the exemplar SVM (ESVM) as a feature encoder for image retrieval. This paper extends this approach in several directions:
+%This paper proposes an extension of the exemplar SVM (ESVM) as a novel approach to feature encoding for feature retrieval applications.
 %This paper proposes a new feature encoding pipeline, based on the success of the exemplar SVM (ESVM) \cite{ZePe15} and linear discriminant analysis (LDA) as whitening \cite{HMR12}.
-We first show that, by replacing the hinge-loss by the square-loss in the ESVM cost function, equivalent results in image retrieval can be obtained at a fraction of the computational cost. 
+We first show that replacing the hinge loss by the square-loss in the ESVM cost function significantly reduces encoding time with negligible effect on
+accuracy.
+%equivalent results in image retrieval can be obtained at a fraction of the computational cost. 
 We call this model square-loss exemplar machine, or SLEM. 
-We also introduce a kernelized SLEM which can be implemented efficiently through a low-rank matrix decomposition and
+We then introduce a kernelized SLEM which can be implemented efficiently through a low-rank matrix decomposition and
 %benefits from the same computational advantages but 
@@
 %% \subsection*{Acknowledgments} This work was supported by the ERC grant VideoWorld.
-
+%\newpage
 {\small

### 2016-11-15 sec_SLEM.tex
version J-1

@@
-\section{Square loss exemplar machine}\label{lsesvm}
+\section{The square-loss exemplar machine}\label{lsesvm}
 In this section, we revisit the exemplar SVM model proposed in \cite{Efros11} as an instance of a more general family of classifiers.
@@
 \begin{equation}
-\big(\omega^\star, \nu^\star\big) = \underset{(\omega,\nu)\in\RR^d\times\RR}{\argmin} \ J(\omega, \nu).
-\footnote{Depending on the loss function $l$, $\nu^\star(x_0,X)$ may be not unique.
-\PP{But, what about $w^{\star}$?}}
+\big(\omega^\star, \nu^\star\big) = \underset{(\omega,\nu)\in\RR^d\times\RR}{\argmin} \ J(\omega, \nu). \footnote{Depending on the loss function $l$, $\nu^\star(x_0,X)$ may be not unique.}
 %\big(\omega^\star(x_0, X), \nu^\star(x_0, X)\big) = \underset{(\omega,\nu)\in\RR^d\times\RR}{\argmin} \ J(\omega, \nu). 
@@
 %Applications of ESVM use the hinge loss function, which guarantees that $J$ is a convex function. 
-The solution of (\ref{omega:first}) can be thus found by stochastic gradient descent~\cite{bottou10} individually for each positive sample.
-The next subsection show how to calculate all exemplar classifiers simultaneously by changing the loss function.
+The solution of Eq.
+(\ref{omega:first}) can thus be found by stochastic gradient descent~\cite{bottou10} individually for each positive sample.
+The next section shows how to calculate all exemplar classifiers simultaneously by changing the loss function.
 
-\subsection{Square loss function}\label{slem:intro}
-Now, let us study the same learning problem for the square loss function $l(y,\hat{y}) = \frac{1}{2}(y-\hat{y})^2$. As in the case of the hinge loss, Equation (\ref{eq:first}) is a convex problem. 
+\subsection{The square-loss}\label{slem:intro}
+Now, let us study the same learning problem for the square-loss function $l(y,\hat{y}) = \frac{1}{2}(y-\hat{y})^2$.
+As in the case of the hinge loss, the minimization of Eq.
+(\ref{eq:first}) is a convex problem.
 However it is now a ridge regression problem, whose unique solution can be found in closed form as
@@
 \end{align}
-Equation (\ref{omega:solution}) shows how to solve (\ref{eq:first}) simultaneously for all individual exemplars.
-\cite{lssvm}
 %Replacing the hinge loss by the square loss offers a more compact solution, but not necessarily more efficient.
@@
 We can thus set $\theta=\frac{1}{n}$ for the remaining of this work.
+
+%Equation (\ref{Wood:omega}) also shows how to solve (\ref{eq:first}) in closed form and, hence, simultaneously for all individual exemplars by solving a linear system~\cite{lssvm}.
+If $X_p$ is the matrix where every column represents one positive sample, we obtain their respective exemplar classifier by solving the system $A\Omega = \bar{X}_p$ in $\Omega$.
+Here, $\bar{X}_p$ denotes the $X_p$ centralized w.r.t.
+the negatives' mean.
+It
 %The normalized exemplar classifier $\omega^\star$ is therefore a linear transformation of $x_0-\mu$.
@@
 
-It is interesting to note the relationship between SLEM and the classical linear discriminant analisys (LDA).
+It is interesting to note the relationship between SLEM and the classical linear discriminant analysis (LDA).
 Let us return to Eq.
@@
 It can be shown that in this case, the corresponding linear classifier of Eq.
-(\ref{eq:first}) for the square loss is also given by
+(\ref{eq:first}) for the square-loss is also given by
 (\ref{omega:solution}), where $x_0$ denotes this time the center of mass
@@
 \displaystyle \nu_{LDA}=-\frac{1}{2}(x_0+\mu)^T \omega_{LDA}.
-\end{array}\right.
+\end{array}\right
 \label{eq:lda}
@@
 
-This shows that SLEM is a generalized version of LDA: Indeed, when $\lambda=0$ in SLEM (\ie, no regularization), $A = \Sigma$, the vectors $\omega_{LDA}$ of Eq.
+This shows that SLEM is a generalized version of LDA: Indeed, taking $\lambda=0$ (\ie no regularization), $A = \Sigma$ and the vectors $\omega_{LDA}$ of Eq.
 (\ref{eq:lda}) and $\omega^\star$ of Eq.
-(\ref{Wood:omega}) have the same direction.
+(\ref{Wood:omega}) have the same direction, reducing SLEM to LDA.
 %This observation has been equally made by \cite{Koba15}.

### 2016-11-15 sec_eff_imp.tex
version J-1

--- 2016-11-13 sec_eff_imp.tex
@@
 \input{figure_LR.tex}
-When compared to the linear square-loss classifier of Section \ref{slem:intro}, one drawback of a kernelized approach is that the dimension of our problem grows with the size $n$ of the negative samples.
+When compared to the linear square-loss classifier of Section \ref{slem:intro}, one drawback of the kernelized approach is that the dimension of our problem grows with the size $n$ of the negative samples.
 The offline factorization $B$ of $K$ demands $O(nr)$ storage and at best $O(nr^2)$ time.
-This factorization can be obtained in different ways, depending on the aspect we wish to optimize.
-Efficiency in time and efficiency in storage are the two most common of these aspects.
-In this section we propose three different decompositions of $K$, to be applied depending if we want to minimize time or storage.
+This factorization can be obtained in two fundamental ways: full-rank and low-rank decomposition.
+In this section we propose three different decompositions of $K$, to be applied depending on the compromise between accuracy and storage we went to achieve.
 %on which of these two efficiencies we hope to achieve.
@@
 
-\subsection{Time-efficient implementation}
-The Cholesky decomposition is the most used factorization of positive-semidefinite matrices in kernel-based learning \cite{BaJo02,BaJo05,FiSc01}. 
-The computation of the factor $B$ depends on a permutation $\pi(n) =\{i_1,i_2,\dots,i_n\}$ of $\{1,2,\dots, n\}$, called \textit{pivots}. The $t$-th column of $B$ is calculated iteratively as:
+\subsection{Full-rank decomposition}
+The complete Cholesky decomposition (CCD) is the most used factorization of positive-definite matrices in kernel-based learning due to its time efficiency \cite{BaJo02}.
+We use it as our default decomposition.
+We make sure $K$ is positive-definite by adding $\epsilon$ to its diagonal, where $\epsilon=\min(0,-\lambda_{min})$ and $\lambda_{min}$ is the smallest eigenvalue of $K$.
+Therefore, from the equation $BB^T=K+\epsilon\mathrm{Id}_n$, $B$ has rank $n$ and can be obtained by CCD.
 
-\begin{align}
-\begin{cases}
-\vspace{3 mm}
-B(i_t, t) = \left(K(i_t,i_t)-\sum_{m=1}^{t-1} B(i_{m},m)  \right)^{\frac{1}{2}},\\
-\vspace{3 mm}
-B(i_j,t) = 0,\\
-B(i_j, t) = \frac{1}{B(i_t,t)}(K(i_j,i_t)  -\sum_{m=1}^{t-1}B(i_j,m)B(i_t,m)), \end{cases}\label{icd:algo}
-\end{align}
-for all $t<j\le n$. There are two possible choices of Cholesky decomposition. The incomplete Cholesky decomposition (ICD) stops its algorithm after $r$ steps and has complexity $O(nr^2)$. 
-To minimize the residue $tr(K-BB^T)$, the $t$-th pivot is calculated at the end of the $t$-th interaction. 
-In the other hand, the complete Cholesky decomposition (CCD) consists of a full-rank decomposition, \textit{i.e.} the rank $r$ of $B$ is equal to $n$ and it has complexity $O(n^3)$. 
-In this case, the pivots are irrelevant to the final residue (which is zero) and can be assumed to be $i_t=t$. %The computation of pivots makes CCD more efficient for similar residues.
-We compare CCD and ICD performances in Fig.
+\subsection{Low-rank decomposition}
+Let us now turn to the problem of improving storage efficiency. One of the major requisites of a global image descriptor for large scale retrieval is to minimize the storage.
+As discussed in Section \ref{simi_score}, for each positive exemplar we store its base feature plus a $2r$ vector. Hence, we aim to decompose $K$ at a small rank $r$.
+We consider the problem of finding a factor $B$ of fixed rank $r$ such that the residue $\mathrm{tr}(K-BB^T)/\mathrm{tr}(K)$ is small.
+\hlc{I wanted to write "factor $B$ of fixed rank $r$ that minimizes the residue $\mathrm{tr}(K-BB^T)/\mathrm{tr}(K)$", but I don't want to propose it as an optimization problem.
+The "such that the residue is small" is not perfect, but it is the best I could find.
+Any proposition?}
+
+The incomplete Cholesky decomposition (ICD) is widely used in machine learning~\cite{BaJo02,BaJo05,FiSc01}.
+Its algorithm is similar to CCD, and it greedily chooses which column of $K$ to add to the decomposition based on the gain in approximation error for all non-added columns.
+This process is called \emph{pivot selection}~\cite{BaJo05}.
+We stop the algorithm $r$ steps, obtaining the factor $B$ in time $O(nr^2)$.
+
+The kernel PCA (KPCA) proposes that, instead of a greedy addition of columns, we choose the $r$ first eigenvectors. This is done by doing a truncated singular value decomposition of $K$:
+\begin{equation}
+    K = WDW^T; \ b_i = \sqrt{d_{ii}}w_i, \label{svd}
+\end{equation}
+where $w_i$ denotes the $i$-th column of $W$ and $d_{ii}$ the $i$-th diagonal element of $D$. Decomposing $K$ from Equation~(\ref{svd}) has complexity $O(n^2r)$ in time.~\cite{golub2012matrix}
+
+We plot in Fig.
+\ref{fig:residue} the ratio $\mathrm{tr}(K-BB^T)/\mathrm{tr}(K)$ when we vary the number of columns $r$ of $B$ and we see a faster convergence for the kernel PCA decomposition when $r$ goes to $n$.
+\hlc{I could not finish the discussion in this section, I need some sleep...}
+
+%The incomplete Cholesky decomposition (ICD) is the most used factorization of positive-semidefinite matrices in kernel-based learning due to its time efficiency \cite{BaJo02,BaJo05,FiSc01}.
+The computation of the factor $B$ depends on a permutation $\pi(n) =\{i_1,i_2,\dots,i_n\}$ of $\{1,2,\dots, n\}$, called \textit{pivots}.
+The $t$-th column of $B$ is calculated iteratively as:
+
+%\begin{align}
+%\begin{cases}
+%\vspace{3 mm}
+%B(i_t, t) = \left(K(i_t,i_t)-\sum_{m=1}^{t-1} B(i_{m},m)  \right)^{\frac{1}{2}},\\
+%\vspace{3 mm}
+%B(i_j,t) = 0,\\
+%B(i_j, t) = \frac{1}{B(i_t,t)}(K(i_j,i_t)  -\sum_{m=1}^{t-1}B(i_j,m)B(i_t,m)), \end{cases}\label{icd:algo}
+%\end{align}
+%for all $t<j\le n$.
+
+%There are two possible choices of Cholesky decomposition. The incomplete Cholesky decomposition (ICD) stops its algorithm after $r$ steps and has complexity $O(nr^2)$. 
+%To minimize the residue $\mathrm{tr}(K-BB^T)$, the $t$-th pivot is calculated at the end of the $t$-th interaction. 
+%In the other hand, the complete Cholesky decomposition (CCD) consists of a full-rank decomposition, \textit{i.e.} the rank $r$ of $B$ is equal to $n$ and it has complexity $O(n^3)$. 
+%In this case, the pivots are irrelevant to the final residue (which is zero) and can be assumed to be $i_t=t$. %The computation of pivots makes CCD more efficient for similar residues.
+%We compare CCD and ICD performances in Fig.
 \ref{fig:residue}.
@@
 
-We make sure $K$ is positive-definite by adding $\epsilon$ to its diagonal, where $\epsilon=\min(0,-\lambda_{min})$ and $\lambda_{min}$ is the smallest eigenvalue of $K$.
+%We make sure $K$ is positive-definite by adding $\epsilon$ to its diagonal, where $\epsilon=\min(0,-\lambda_{min})$ and $\lambda_{min}$ is the smallest eigenvalue of $K$.
 Therefore, from the equation $BB^T=K+\epsilon\mathrm{Id}_n$, $B$ has rank $n$ and we can apply CCD.
 
-\subsection{Storage-efficient implementation}\label{low-rank} %/Low-rank decomposition} 
-Let us now consider instead the problem of improving storage efficiency.
-As discussed in section \ref{simi_score}, for each positive exemplar we store its original encoder plus a $2r$ vector, so we aim to decompose $K$ at a small rank $r$.
-For small values of $r$, a kernel PCA (KPCA) factorization guarantees a better approximation of $K$ than the Cholesky decomposition at a higher time cost.
+%\subsection{Storage-efficient implementation}\label{low-rank} %/Low-rank decomposition} 
+%Let us now consider instead the problem of improving storage efficiency.
+As discussed in section \ref{simi_score}, for each positive exemplar we store its original encoding plus a $2r$ vector, so we aim to decompose $K$ at a small rank $r$.
+%For small values of $r$, a kernel PCA (KPCA) factorization guarantees a better approximation of $K$ than the Cholesky decomposition at a higher time cost.
 Indeed, as illustrated in Figure \ref{fig:residue}, a kernel PCA gives smaller residues than ICD at small rank.
-We plot the ratio $tr(K-BB^T)/tr(K)$ when we vary the number of columns $r$ of $B$, and we see a faster convergence for the kernel PCA decomposition when $r$ goes to $n$.
-However, at these small values of $r$, KPCA is more time consuming than both ICD and CCD.
-This time cost comes from applying a singular values decomposition of K to obtain $B$:
-\begin{equation}
-    K = WDW^T; \ b_i = \sqrt{d_{ii}}w_i, \label{svd}
-\end{equation}
-where $w_i$ denotes the $i$-th column of $W$ and $d_{ii}$ the $i$-th diagonal element of $D$.
-Decomposing $K$ from Equation~(\ref{svd}) has complexity $O(n^2r)$ in time.
-\hlc{Francis, I need a reference for this complexity.}
-
-ICD may be the fastest decomposition for small ranks, but it is outperformed by KPCA; for higher ranks, ICD is the slowest alternative and its performance approximates CCD's performance.
-We therefore use KPCA for our storage-efficient.
-%\hlc{RR: However, ICD is still the only $O(nr^2)$ in time, which could be useful for bigger values of $n$.
-I'm considering showing the experiments done in the ECCV supmat to show why ICD is obsolete.}
-
-%\begin{figure}
-%    \centering
-%    \includegraphics[width=0.5\textwidth]{trace_residual.jpg}
-%    \caption{Residual error of decompositions varying by its rank.}
-%    \label{fig:residue}
-%\end{figure}

### 2016-11-15 sec_eval_raf.tex
version J-1

--- 2016-11-13 sec_eval_raf.tex
@@
 \subsection{Datasets and evaluation protocol} \label{eval:protocol}
-We perform our experiments on three standard datasets for image retrieval. 
+We perform our experiments on three standard datasets for image retrieval:
 \begin{itemize}
-    \item The INRIA \emph{Holidays} dataset \cite{holidays} consists of 1491 images divided in 500 groups of matching images.
-We manually rotate by 90 degrees some images that are not in their natural orientation to compensate for the fact that CNN features are not rotation invariant~\cite{babenko14,Arandjelovic15,RaToCh16}
+    \item The INRIA \emph{Holidays} dataset \cite{holidays}, which consists of 1491 images divided in 500 groups of matching images.
+We manually rotate by 90 degrees some images that are not in their natural orientation to compensate for the fact that CNN features are not rotation invariant~\cite{Arandjelovic15,babenko14,RaToCh16}
     \item The \emph{Oxford5k} dataset \cite{oxford}, which consists of 5063 images separated in 55 groups of matching images, each group associated to a landmark of Oxford.
-We use the "full" crop, ignoring the region of interest of each image.
-    \item The \emph{Oxford105k} dataset \cite{oxford}, large-scale dataset containing the same images and queries from Oxford5k plus \emph{Flickr100k}, a collection of $10^5$ distractor images Flickr.
+We use the ``full'' crop, ignoring the region of interest of each image.
+    \item The \emph{Oxford105k} dataset \cite{oxford}, a large-scale dataset containing the same images and queries from Oxford5k plus \emph{Flickr100k}, a collection of $10^5$ distractor Flickr images.
 \end{itemize}
@@
 For low rank decomposition, we use all 100000 images.
-
 At evaluation time, for a dataset that consists of $p$ images and $q$ query images, we calculate the $p\times q$ \emph{similarity matrix} $S$, where each of its $q$ columns is the matching scores of the query image with all the $p$ images.
-\hlc{RR: maybe take out this paragraph completely.}
 
@@
 Baseline            & 72.7 & 76.5 & 68.2  &  85.4 & 46.3 & 54.4 & 40.6 & 67.5 & 50.1 & - \\% & \\
-%Whitening           & -    & -    & -    & -    & -\\
 PCAW                & 75.5 & 81.7 & 69.2 & 88.3 & 50.9  & 63.7 & 45.0 & 69.1 & 55.5 & - \\% & \\
 LDA                 & 54.7 & 82.2 & 64.1 & 74.3 & 29.6  & 62.2 & 42.5 & 72.7 & 52.4 & - \\
-ESVM~\cite{ZePe15}  & $77.5^3$ & $84.0^3$ & 71.3 & $91.4^2$ & 57.2  & 62.1 & 43.9 & 72.5 & 56.5 & - \\% & \\ 
-Linear SLEM         & $78.0^2$ & 82.3 & 72.1 & $91.3^3$ & \ul{59.3}  & 64.1 & 46.2 & 72.9 & 56.7 & - \\% & \\ /42.2
-Gaussian SLEM (16)  & 76.8 & 80.3 & 71.2 & $91.4^2$ & 52.8 & 63.0 & 43.5 & 71.9 & 55.8 & - \\
-Gaussian SLEM (32)  & 77.4 & 81.7 & $72.0^3$ & $91.4^2$ & 54.9 & 63.1 & 44.0 & 71.1 & 56.0 & - \\
-Gaussian SLEM (fr)  & \bf{78.1} & $86.2^2$ & \bf{72.9} & \bf{91.7} & 59.0   & \ul{64.9} & 47.0 & \ul{74.4} & 59.5 & - \\% & \\ 
-Poly SLEM (16)      & 76.9 & 82.3 & 71.4 & $91.3^3$ & 53.0 & 63.6 & 43.6 & 71.4 & 56.1 & - \\
-Poly SLEM (32)      & 77.3 & 82.4 & $72.1^2$ & \bf{91.7} & 54.9 & 63.6 & 44.1 & 71.6 & 56.3 & - \\
-Poly SLEM   (fr)    & \bf{78.1} & \bf{86.3}  & \bf{72.9} & \bf{91.7} & \ul{59.3}  & 64.8 & \ul{47.3} & \ul{74.1} & 62.5 & - \\
+ESVM~\cite{ZePe15}  & 77.5^3 & 84.0^3 & 71.3 & 91.4^2 & 57.2^3 & 62.1 & 43.9 & 72.5 & 56.5 & - \\% & \\ 
+Linear SLEM         & 78.0^2 & 82.3 & 72.1 & 91.3^3 & \bf{59.3}  & 64.1^3 & 46.2^3 & 72.9^3 & 56.7^3 & - \\% & \\ /42.2
+Gaussian SLEM (16)  & 76.8 & 80.3 & 71.2 & 91.4^2 & 52.8 & 63.0 & 43.5 & 71.9 & 55.8 & - \\
+Gaussian SLEM (32)  & 77.4 & 81.7 & 72.0^3 & 91.4^2 & 54.9 & 63.1 & 44.0 & 71.1 & 56.0 & - \\
+Gaussian SLEM (fr)  & \bf{78.1} & 86.2^2 & \bf{72.9} & \bf{91.7} & 59.0^2 & \bf{64.9} & 47.0^2 & \bf{74.4} & 59.5^2 & - \\% & \\ 
+Poly SLEM (16)      & 76.9 & 82.3 & 71.4 & 91.3^3 & 53.0 & 63.6 & 43.6 & 71.4 & 56.1 & - \\
+Poly SLEM (32)      & 77.3 & 82.4 & 72.1^2 & \bf{91.7} & 54.9 & 63.6 & 44.1 & 71.6 & 56.3 & - \\
+Poly SLEM   (fr)    & \bf{78.1} & \bf{86.3}  & \bf{72.9} & \bf{91.7} & \bf{59.3}  & 64.8^2 & \bf{47.3} & 74.1^2 & \bf{62.5} & - \\
 \hline
@@
 In this table, we present our results for VLAD \cite{Delhumeau2013}, sum-pooling of convolutional features (SPoC) \cite{babenko15}, activation coefficients from the previous-to-last CNN layer (AlexNet) \cite{Krizhevsky2012} and activation of NetVLAD layer~\cite{Arandjelovic15}.
-In parenthesis, the rank of he decomposition ('fr' for full rank decomposition)}
+In parenthesis, the rank of he decomposition ('fr' for full rank decomposition).
+For each column, we \textbf{bold} the best results and index the second and third best.}
+\label{fullrank}
 \end{center}
-\label{fullrank}
 \end{table*}
@@
 \subsection{Which kernel to choose?}
-For the experiment of SLEM as a feature encoder, we tested two different kernels -- Gaussian and polynomial -- each with a scalar parameter $\gamma$:
+We have tested two different kernels, each with a scalar parameter $\gamma$:
+%For the experiment of SLEM as a feature encoder, we tested two different kernels -- Gaussian and polynomial -- each with a scalar parameter $\gamma$:
 \begin{align}
@@
 well known reproducing kernel, used for classification with support vector machines.\\
+\begin{align}
+    k_1(x,y) = e^{-\gamma\|x-y \|^2};
+\end{align}
 \textbf{Poly SLEM.} The polynomial kernel $k_2$ is a reproducing kernel often used in natural language processing.\\
+\begin{align}
+    k_2(x,y) = x^Ty+\gamma(x^Ty)^2.
+\end{align}
 %Note the non-kernel version of SLEM is equivalent to a SLEM for the linear kernel.
@@
 
-The first CNN-based features we use consist of the activation coefficients of the previous-to-last layer of the AlexNet architecture \cite{Krizhevsky2012}, based on a publicly available pre-trained model \cite{jia2014caffe}.
+The first CNN features we use consist of the activation coefficients of the previous-to-last layer of the AlexNet architecture \cite{Krizhevsky2012}, based on a publicly available pre-trained model \cite{jia2014caffe}.
 These are also the features used in \cite{ZePe15}.
@@
 We use the base features of the previous subsection as baseline.
-Since \textit{Babenko et al.}~\cite{babenko15} and \textit{Arandjelovi\'c et al.}~\cite{Arandjelovic15} have improved retrieval results by applying PCA followed by whitening to their features, we also apply this post-processing to our base features as second baseline (PCAW), compressing its its dimension to half of the original.
-We then compare the baselines with the original ESVM, LDA and several variants of our approach (SLEM), since all the those methods are based on similar \hlc{???
-(I can't read the last word you wrote.} The results are presented in Table \ref{fullrank}.
+Since Babenko and Lemptisky~\cite{babenko15} and Arandjelovi\'c \etal~\cite{Arandjelovic15} have improved retrieval results by applying PCA followed by whitening to their features, we also apply this post-processing to our base features as a second baseline (PCAW), compressing base feature dimension to half of the original.
+We then compare the baselines with the original ESVM, LDA and several variants of our approach (SLEM), since all the those methods are based on similar ideas.
+The results are presented in Table \ref{fullrank}.
+For the large-scale dataset that is Oxford105k, we limit our experiments to our best performing base features, SPoC and NetVLAD.
 
-Linear SLEM performs similarly to ESVM despite being much more time efficient (see discussion in section \ref{time-scale}). 
-The fact that a hinge-loss classifier does not outperform a square-loss classifier can seem counter-intuitive, but both were shown to be equivalent for binary classification under mild constraints~\cite{YeXi07}.
+Linear SLEM performs similarly to ESVM despite being much more time efficient (Fig. \ref{fullrank:results}). 
+The fact that a hinge-loss classifier does not outperform a square-loss classifier can seem counter-intuitive, but both have been shown to be equivalent for binary classification under mild constraints~\cite{YeXi07}.
 
-We perform both Gaussian SLEM and Polynomial SLEM with two decompositions: one full-rank CCD decomposition signaled by (fr) and two low-rank KPCA decompositions signaled by the rank of the decomposition.
+We use both Gaussian SLEM and Polynomial SLEM with two decompositions: one full-rank CCD decomposition indicated by (fr) and two low-rank KPCA decompositions indicated by the rank of the decomposition.
 We train our exemplar classifiers for between $6000$ and $15000$ negative samplers.
-For all the experiments we calibrate the regularization cost $\lambda$, as well as the parameter $\gamma$.
+For all the experiments we calibrate the regularization cost $\lambda$, as well as the parameter $\gamma$ similarly to the calibration in~\cite{ZePe15}.
 
 The full-rank variant outperform all methods for all base features, although the gains when compared to linear SLEM are not always significant (\emph{e.g.} for VLAD features).
+We notice significant improvement for SPoC in both Holidays and Oxford and for AlexNet and NetVLAD in Oxford.
 It is interesting to notice that LDA, ESVM and Linear SLEM do not seem to outperform one another or the baselines consistently for all features and datasets.
@@
 \small
-\begin{tabular}{l|c|ll}
+\begin{tabular}{l|c|c|ll}
 \toprule
-Features & dim & \textbf{Hol.} & \textbf{Ox5k} \\%& \textbf{Oxford 105k}\\
+Features & rank & dim & \textbf{Hol.} & \textbf{Ox5k} \\%& \textbf{Oxford 105k}\\
 \midrule
-Babenko \textit{et al.}\cite{babenko15}  & 256 & 80.2 & 58.9 \\ %& 57.8\\
-Radenovi\'c \textit{et al.} \cite{RaToCh16}   & 256 & 81.5 & \un{77.4} \\
-Arandjelovi\'c \textit{et al.} \cite{Arandjelovic15}& 256 & 86.0 & 62.5 \\ %& - \\
-Kalantidis  \textit{et al.} \cite{KaMeOs16}   & 256 & 83.1 & 65.4 \\
-SPoC + Linear SLEM & 256 & 81.5 & 64.7 \\
-SPoC + Poly SLEM & 288 & 80.1 & 63.6 \\
-SPoC + Poly SLEM & 320 & 81.8 & 63.6 \\
-NetVLAD + Linear SLEM & 256 & \un{88.5} & 65.9 \\
-NetVLAD + Poly SLEM & 288 & 87.7 & 65.5 \\
-NetVLAD + Poly SLEM & 320 & 88.3 & 65.6 \\
+Babenko \textit{et al.}\cite{babenko15}         & - & 256 & 80.2 & 58.9 \\ 
+Radenovi\'c \textit{et al.} \cite{RaToCh16} & - & 256 & 81.5 & \un{77.4} \\
+Arandjelovi\'c \textit{et al.} \cite{Arandjelovic15} & - & 256 & 86.0 & 62.5 \\ 
+Kalantidis  \textit{et al.} \cite{KaMeOs16}   & - & 256 & 83.1 & 65.4 \\
+SPoC + Linear SLEM & - & 256 & 81.5 & 64.7 \\
+SPoC + Poly SLEM & 16 & 288 & 80.1 & 63.6 \\
+SPoC + Poly SLEM & 32 & 320 & 81.8 & 63.6 \\
+NetVLAD + Linear SLEM & - & 256 & \un{88.5} & 65.9 \\
+NetVLAD + Poly SLEM & 16 & 288 & 87.7 & 65.5 \\
+NetVLAD + Poly SLEM & 32 & 320 & 88.3 & 65.6 \\
 \midrule
-Radenovi\'c \textit{et al.} \cite{RaToCh16}   & 512 & 82.5 & 79.7 \\
-Arandjelovi\'c \textit{et al.} \cite{Arandjelovic15}& 512 & 86.7 & 65.6 \\
-Kalantidis \textit{et al.} \cite{KaMeOs16}   & 512 & 84.9 & 68.2 \\
-Gordo \textit{et al.} \cite{GoAlReLa16} & 512 & 89.1^{\dag} & \bf{83.1}^{\dag} \\
-SPoC + Linear SLEM & 512 & 82.3 & 64.1 \\
-SPoC + Poly SLEM & 544 & 82.3 & 63.0 \\
-SPoC + Poly SLEM & 576 & 82.4 & 63.1 \\
-NetVLAD + Linear SLEM & 512 & 89.3 & 72.3 \\
-NetVLAD + Poly SLEM & 544 & \un{89.9} & 71.9 \\
-NetVLAD + Poly SLEM & 576 & \un{89.9} & 72.3 \\
+Radenovi\'c \textit{et al.} \cite{RaToCh16}   & - & 512 & 82.5 & 79.7 \\
+Arandjelovi\'c \textit{et al.} \cite{Arandjelovic15}& - & 512 & 86.7 & 65.6 \\
+Kalantidis \textit{et al.} \cite{KaMeOs16}   & - & 512 & 84.9 & 68.2 \\
+Gordo \textit{et al.} \cite{GoAlReLa16} & - & 512 & 89.1^{\dag} & \bf{83.1}^{\dag} \\
+SPoC + Linear SLEM & - & 512 & 82.3 & 64.1 \\
+SPoC + Poly SLEM   & 16 & 544 & 82.3 & 63.0 \\
+SPoC + Poly SLEM   & 32 & 576 & 82.4 & 63.1 \\
+NetVLAD + Linear SLEM & - & 512 & 89.3 & 72.3 \\
+NetVLAD + Poly SLEM & 16 & 544 & \un{89.9} & 71.9 \\
+NetVLAD + Poly SLEM & 32 & 576 & \un{89.9} & 72.3 \\
 \midrule
-Arandjelovi\'c \textit{et al.} \cite{Arandjelovic15}& 4096 & 88.3 & 69.1 \\
+Arandjelovi\'c \textit{et al.} \cite{Arandjelovic15}& - & 4096 & 88.3 & 69.1 \\
 %Li \textit{et al.} \cite{LiLaHa15}& - & 89.2^{\dag} & 73.7 \\
-NetVLAD + Linear SLEM & 4096 & 91.3 & 72.9 \\
-NetVLAD + Poly SLEM & 4128 & 91.3 & 71.2 \\
-NetVLAD + Poly SLEM & 4160 & \bf{91.7} & 71.7 \\
+NetVLAD + Linear SLEM & - & 4096 & 91.3 & 72.9 \\
+NetVLAD + Poly SLEM & 16 & 4128 & 91.3 & 71.2 \\
+NetVLAD + Poly SLEM & 32 & 4160 & \bf{91.7} & 71.7 \\
 %Zepeda \textit{et al.} \cite{ZePe15} & \underline{78.3} & - & 71.8 & - & 57.5 & - & 44.6 & - & - \\

### 2016-11-15 sec_intro.tex
version J-1

--- 2016-11-13 sec_intro.tex
@@
 
-An effective image 
-representation is a crucial component of image retrieval system
-where, given some query image,  images must be retrieved from a large 
-database, then ranked according to their degree of similarity with the 
-query.
+This paper addresses the problem of designing an image representation suitable for content-based retrieval, in particular supporting effective (discriminative) and
+efficient (fast) comparisons between a query picture and images stored in some large database.
 %Robust image representations are a crucial component of a vast number of computer vision applications. 
 %Within these, the \emph{image retrieval} application is an important example wherein a query image is provided to the system, which must in turn find all matching images from within a large, unannotated database.
-The matching process is done entirely based on the pixel content of the query and database images, and the search must be robust to large image variations due to camera pose, color differences and scene illumination, amongst others.
+These comparisons are done entirely based on the pixel content of the query and database images, and the search must be robust to large image variations due to camera pose, color differences and scene illumination, amongst others.
 
-The advances in image representation enabled the success of new systems in this challenging task.
+%The advances in image representation enabled the success of new systems in this challenging task.
 %An image representation function commonly maps a given input image to a vector in a fixed dimensional space called the image \emph{feature vector}.
-Indeed, many successful feature representations for image retrieval rely on unsupervised models such as $K$-means \cite{Delhumeau2013} or Gaussian mixture models \cite{Perronnin2010} and, before the neural networks \textit{renaissance}, these representations would outperform methods that exploit supervised learning of image features directly~\cite{Bilen2015,Rana}.
-Today, with the popularization of convolutional architectures, global image descriptors are often obtained by aggregation and/or pooling the last convolutional layers \cite{babenko15,GoAlReLa16,KaMeOs16} or addition of new differentiable layers to an existing architecture~\cite{Arandjelovic15}.
+Many successful image representations for retrieval rely on unsupervised models such as $K$-means \cite{Delhumeau2013} or Gaussian mixture models \cite{Perronnin2010,PeSan10} and, before the neural networks \textit{renaissance}, these representations would outperform methods that exploit supervised learning of image features directly~\cite{Bilen2015,Rana}.
+Today, with the success of convolutional architectures, global image descriptors are often obtained by aggregating and/or pooling the last convolutional layers \cite{babenko15,GoAlReLa16,KaMeOs16} or by addition of new differentiable layers to an existing architecture~\cite{Arandjelovic15}.
 %One of the main reasons for this is the lack of adequately large and varied supervised datasets that are expensive to collect.
@@
 It uses a large generic pool of images as a set of negative examples, while using a single image (the \emph{exemplar}) as a positive example.
-Given these training sets, a classifier is learned that can  generalize well, despite the drastically limited size of the set of positive examples.
+Given these training sets, a SVM classifier is learned that can generalize well, despite the drastically limited size of the set of positive examples.
 
-Zepeda \emph{et al.} \cite{ZePe15} propose to instead treat the weights of the resulting classifier as a new feature vector for image retrieval.
+Zepeda and P\'erez \cite{ZePe15} propose to treat instead the weights of the resulting classifier as a new feature vector for image retrieval.
 %This new ESVM feature is derived from \emph{base} feature representations ({\it e.g.,} CNN activation features) of the exemplar and the negative pool.
 %An extension \cite{ZePe15} of the ESVM formulation discussed above instead treats the resulting  classifier as an \emph{enhanced} representation of the image feature representing the exemplar image (the \emph{base} feature representation).
-An ESVM feature is extracted from each database image, as well as from the query image, by treating each image as an exemplar while keeping a fixed pool of generic negative images.
+An ESVM feature is extracted from each database and query image, by treating it as the only positive sample (an \emph{exemplar}) while keeping a fixed pool of generic negative images.
 Searching amounts to computing distances between the query and the database ESVM features.
@@
 
-One important drawback of the ESVM feature encoding approach is that computing the  classifier requires solving an optimization problem for each positive example.
+One drawback of the ESVM feature encoding approach is that computing the  classifier requires solving an optimization problem for each positive example (\ie, each query and each database image).
 This can be time consuming for the large negative pool sizes required for good ESVM feature performance.
-In this work we propose using the square loss instead of the hinge loss, in effect converting the ESVM problem into a ridge regression that can be solved in closed form.
-The parameters of the corresponding classifier, which we call the squared loss exemplar machine (SLEM), had been shown to both generalize LDA \cite{Koba15} and, for the general case of multiple positive samples (LS-SVM)~\cite{lssvm}, form features with comparable binary classification performance while drastically reducing the computational cost.
+In this work, we propose using the square-loss instead of the hinge loss, in effect converting the ESVM problem into a ridge regression that can be solved in closed form.
+We dub the corresponding classifier a {\em square-loss exemplar machine} (or {\em SLEM}).
+The square-loss has been used before to replace the hinge loss in classification tasks (e.g.,~\cite{lssvm,YeXi07}), and the SLEM has also been used to compare ESVMs to classical classifiers such as the linear discriminant analysis (or {\em LDA})~\cite{Koba15}.
+We focus here on the role of SLEMs as feature encoder for image retrieval instead of categorization.
+%The parameters of the corresponding classifier, which we call the squared loss exemplar machine (SLEM), had been shown to both generalize LDA \cite{Koba15} and, for the general case of multiple positive samples (LS-SVM)~\cite{lssvm}, form features with comparable binary classification performance while drastically reducing the computational cost.
 
 Since computing the SLEM features requires inverting a large matrix related to the training set's covariance matrix, we propose an efficient way to compute this inverse that exploits the fact that only a single (positive) example changes in the training set when computing SLEM features for different images.
-We show experimentally that our representation matches and even improves upon the performance of ESVM features on two standard datasets using a wide range of base features.
+We show experimentally that our representation matches and even improves upon the performance of ESVM features on three standard datasets using a wide range of base features.
 
-We also introduce a kernelized variant of SLEM that enjoys similar computational advantages but improves retrieval performances.
-Further computational efficiency is obtained using a kernel principal components analysis (KPCA) decomposition of the negative samples.
+We also introduce a kernelized variant of SLEM that enjoys similar computational advantages and improves retrieval performances. 
+This kernel approach allows the use of any global image feature (\eg HoG~\cite{hog}, BoW~\cite{BoW04}, spatial pyramids~\cite{spk}) as base representation, independent of its use on retrieval.
+Further computational efficiency is obtained using low-rank factorization methods, such as the incomplete Cholesky decomposition (ICD) or kernel principal components analysis (KPCA), to decompose the kernel matrix of negative samples.
 %low rank incomplete Cholesky decomposition of the negative pool kernel matrix.

### 2016-11-15 sec_kernel_methods.tex
version J-1

@@
 where $b_i^T$ denotes the $i$-th row of $B$ and $r$ is the number of columns of $B$. 
-If $(\beta^\star, \nu^\star)$ is the solution of (\ref{beta:first}), the corresponding vector $\alpha^\star$ (or, more correctly, \textit{a} corresponding vector of dimension $n\geq r$) 
+If $(\beta^\star, \nu^\star)$ is the solution of Eq. (\ref{beta:first}), the corresponding vector $\alpha^\star$ (or, more correctly, \textit{a} corresponding vector of dimension $n\geq r$) 
 can be computed by $\alpha^\star=P\beta^\star$, where $P$ is the pseudoinverse of $B^T$.
@@
 Note that Eq.
-(\ref{beta:first}) allows us to write the kernel learning problem (\ref{ker:aff}) as an instance of Eq.
+(\ref{beta:first}) is written as the usual form of a linear classifier.
+In particular, it allows us to write the kernel learning problem (\ref{ker:aff}) as an instance of Eq.
 (\ref{eq:first}) by setting $y_i=-1$ for all but one training sample.
@@
 
-\subsection{Offline preprocess of negative samples}\label{offline}
+\subsection{Offline preprocessing of negative samples}\label{offline}
+Let us now return to the (kernelized) SLEM with $l$ as the square-loss.
 In order to calculate offline all operations that are dependent only on negative samples, let us suppose that the positive training sample is unknown and $K$ is the kernel matrix of the negative samples $X$.
@@
 (\ref{Wood:omega}): $\mu = \frac{1}{n}\sum_{i=1}^n b_i^T$ and $A = \frac{1}{n}B^TB-\mu\mu^T +\lambda\mathrm{Id}_r.$
-These operations are done offline and their results are stored. In the processing of positive samples, \emph{i.e.} the online phase, these results are loaded.
+These operations are done offline and their results are stored. 
+%In the processing of positive samples, \emph{i.e.} the online phase, these results are loaded.
 %Equation (\ref{beta:first}) has the same structure of Eq. (\ref{eq:first}) if we were to consider $\theta=0$ and all labels $y_i=-1$.
@@
 \begin{align}
- \dfrac{1}{n}l(1, b_0'^T\beta+\nu) +\dfrac{1}{n}\sumi l(-1,b_i'^T\beta+\nu)
+ \dfrac{1}{n}(b_0'^T\beta+\nu - 1)^2 +\dfrac{1}{n}\sumi (b_i'^T\beta+\nu + 1)^2
 +\dfrac{\lambda}{2}\|\beta\|^2,\label{beta:final}
@@
 In particular, $b_0'=[u; \ v]$ and, for $i>0$, $b_i'=[0; \ b_i]$.
-The solution $(\beta^\star,\nu^\star)$ in $\RR^{r+1}\times\RR$ can be computed just as before by Equation (\ref{omega:solution}), replacing $x_0$ by $b_0'$, $\mu$ by $\mu' = \frac{1}{n}\sum_{i=1}^n b_i'$ and $X$ by the $(r+1)\times n$ matrix $Q$ of columns $b_1', b_2',...,b_n'$.
-Solution $\alpha^\star$ is now calculated as
+The solution $(\beta^\star,\nu^\star)$ in $\RR^{r+1}\times\RR$ can be computed just as before by Eq.
+(\ref{omega:solution}), replacing $x_0$ by $b_0'$, $\mu$ by $\mu' = \frac{1}{n}\sum_{i=1}^n b_i'$ and $X$ by the $(r+1)\times n$ matrix $Q$ of columns $b_1', b_2',...,b_n'$.
+The solution $\alpha^\star$ is now calculated as
 $\alpha^\star=P'\beta^\star$, where $P' = [u^{-1} \ 0^T; -u^{-1}Pv \ P]$ is the pseudoinverse of $B'^T$. 
@@
 \subsection{Similarity score}\label{simi_score}
+%It is important to note that the exemplar classifier $h$ is never explicitly calculated, but we are able to compute the similarity between a pair of exemplar classifiers regardless.
 Once the optimal parameters $(\beta, \nu)$ from (\ref{beta:final}) and the coordinates $u$, $v$ of $b_0'$ from (\ref{eq:lemma1}) have been found\footnote{We drop the ``$\star$'' in this subsection to avoid cluttering the notation.}, they can be used directly for measuring similarity between matching images.
 
-Indeed, suppose two image descriptors $x_0$ and $x_0'$ are given and we wish to calculate the similarity score between theirs SLEM representations $h$ and $h'$, denoted by $s(h,h')$. 
+Suppose two image descriptors $x_0$ and $x_0'$ are given and we wish to calculate the similarity score between theirs SLEM representations $h$ and $h'$, denoted by $s(h,h')$. 
 We write $h'=\alpha_0'\varphi(x_0')+\sum_{i=1}^n \alpha_i'\varphi (x_i)+\nu'$. 
@@
 
-For a given image whose descriptor is $x_0$, we need to store $x_0$, $\hat{\beta}$ and $v$ to calculate its matching score to whichever other image for SLEM.
-Each image depends therefore on a vector of $p+2r$ dimensions \PP{Remind what are $p$ and $r$, I lost the former!}.
+For a given image whose descriptor is $x_0$, we need to store $x_0$, $\hat{\beta}$ and $v$ to calculate its similarity score to whichever other image for SLEM.
+Since we assume the base feature $x_0$ has dimension $p$ and $\hat{\beta}$ and $v$ each have dimension $r$, we store a vector of dimension $p+2r$ for each image.
 

### 2016-11-15 sec_eff_imp.tex
version J-1

@@
 We make sure $K$ is positive-definite by adding $\epsilon$ to its diagonal, where $\epsilon=\min(0,-\lambda_{min})$ and $\lambda_{min}$ is the smallest eigenvalue of $K$.
-Therefore, from the equation $BB^T=K+\epsilon\mathrm{Id}_n$, $B$ has rank $n$ and can be obtained by CCD.
+Therefore, from the equation $BB^T=K+\epsilon\mathrm{Id}_n$, $B$ has rank $n$ and can be calculated by CCD.
 
@@
 Let us now turn to the problem of improving storage efficiency. One of the major requisites of a global image descriptor for large scale retrieval is to minimize the storage.
-As discussed in Section \ref{simi_score}, for each positive exemplar we store its base feature plus a $2r$ vector. Hence, we aim to decompose $K$ at a small rank $r$.
-We consider the problem of finding a factor $B$ of fixed rank $r$ such that the residue $\mathrm{tr}(K-BB^T)/\mathrm{tr}(K)$ is small.
+As discussed in Section \ref{simi_score}, for each positive exemplar we store its base representation plus a $2r$ vector. Hence, we aim to decompose $K$ at a small rank $r$.
+We consider the problem of finding a factor $B$ of fixed rank $r$ such that the residue.
+We propose two methods of decomposition.
+$\mathrm{tr}(K-BB^T)/\mathrm{tr}(K)$ is small.
 \hlc{I wanted to write "factor $B$ of fixed rank $r$ that minimizes the residue $\mathrm{tr}(K-BB^T)/\mathrm{tr}(K)$", but I don't want to propose it as an optimization problem.
@@
 
-The incomplete Cholesky decomposition (ICD) is widely used in machine learning~\cite{BaJo02,BaJo05,FiSc01}.
+\textbf{ICD} The incomplete Cholesky decomposition (ICD) is widely used in machine learning~\cite{BaJo02,BaJo05,FiSc01}.
 Its algorithm is similar to CCD, and it greedily chooses which column of $K$ to add to the decomposition based on the gain in approximation error for all non-added columns.
@@
 
-The kernel PCA (KPCA) proposes that, instead of a greedy addition of columns, we choose the $r$ first eigenvectors. This is done by doing a truncated singular value decomposition of $K$:
+\textbf{KPCA} The kernel PCA (KPCA) proposes that, instead of a greedy addition of columns, we choose the $r$ first eigenvectors to form the $r$ columns of $B$.
+This is done by doing a truncated singular value decomposition of $K$:
 \begin{equation}
@@
 \end{equation}
-where $w_i$ denotes the $i$-th column of $W$ and $d_{ii}$ the $i$-th diagonal element of $D$. Decomposing $K$ from Equation~(\ref{svd}) has complexity $O(n^2r)$ in time.~\cite{golub2012matrix}
+where $w_i$ denotes the $i$-th column of $W$ (\ie the $i$-th eigenvector of $K$) and $d_{ii}$ the $i$-th diagonal element of $D$ (\ie the $i$-th eigenvalue of $K$).
+Decomposing $K$ from Eq.~(\ref{svd}) has complexity $O(n^2r)$ in time.~\cite{golub2012matrix}
 
 We plot in Fig.
-\ref{fig:residue} the ratio $\mathrm{tr}(K-BB^T)/\mathrm{tr}(K)$ when we vary the number of columns $r$ of $B$ and we see a faster convergence for the kernel PCA decomposition when $r$ goes to $n$.
+\ref{fig:residue} the residue $\mathrm{tr}(K-BB^T)/\mathrm{tr}(K)$ when we vary the number of columns $r$ of $B$ and we see a faster convergence for the kernel PCA decomposition when $r$ goes to $n$.
+Figure \ref{kpca:icd} shows an accuracy comparison between the two decompositions, confirming that KPCA gives a better approximation of the full-rank decomposition.
+
+\input{fig_KPCAvsICD.tex}
+
+We also compare the time of calculation in Fig. \ref{fig:residue}, which shows that KPCA is slower than ICD for small values of $r$ and faster for values of $r$ such that the residue is small.
 \hlc{I could not finish the discussion in this section, I need some sleep...}

### 2017-01-30 rebuttal.tex
Create rebuttal.tex

\documentclass[10pt,twocolumn,letterpaper]{article}

\usepackage{cvpr}
\usepackage{times}
\usepackage{epsfig}
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{amssymb}

\usepackage{graphicx}
\usepackage{caption}
\usepackage{subfigure}
\usepackage{subcaption}
\captionsetup{compatibility=false}
\usepackage{amssymb}
\usepackage{amsthm}
%\usepackage{ruler}
\usepackage{color}
\usepackage{booktabs}
%\usepackage[width=122mm,left=12mm,paperwidth=146mm,height=193mm,top=12mm,paperheight=217mm]{geometry}
\newtheorem{proposition}{Proposition}
\newtheorem{lemma}{Lemma}
%\newtheorem{proof}{Proof}
\newcommand{\RR}{\mathbb R}
\newcommand{\NNN}{\mathcal N}
\newcommand{\sumi}{\displaystyle{\sum_{i=1}^n}}
\DeclareMathOperator*{\argmin}{argmin}

% Highlighting
\usepackage{soul}
%\usepackage[usenames,dvipsnames]{xcolor}
\newcommand{\hlc}[2][yellow]{{\sethlcolor{#1}\hl{#2}}}
\newcommand{\RAF}[1]{\hlc[yellow]{(RR:) #1}}
\newcommand{\JZ}[1]{\hlc[pink]{(JZ:) #1}}
\newcommand{\PP}[1]{\hlc[green]{PP: #1}}
\newcommand{\un}[1]{\underline{#1}}
\def\dul#1{\underline{\underline{#1}}}

\usepackage{tikz}
\usepackage{pgfplots,pgfplotstable}
\usetikzlibrary{pgfplots.groupplots}
\pgfplotsset{grid=major,height=2in, width=\columnwidth}
\input{tikz_styles.tex}

% Include other packages here, before hyperref.

% If you comment hyperref and then uncomment it, you should delete
% egpaper.aux before re-running latex.  (Or just hit 'q' on the first latex
% run, let it finish, and you should be clear).
\usepackage[pagebackref=true,breaklinks=true,letterpaper=true,colorlinks,bookmarks=false]{hyperref}

% \cvprfinalcopy % *** Uncomment this line for the final submission

\def\cvprPaperID{882} % *** Enter the CVPR Paper ID here
\def\httilde{\mbox{\tt\raisebox{-.5ex}{\symbol{126}}}}

% Pages are numbered in submission mode, and unnumbered in camera-ready
\ifcvprfinal\pagestyle{empty}\fi
\begin{document}

%%%%%%%%% TITLE
\title{Kernel Square-Loss Exemplar Machines for Image Retrieval: Rebuttal}

\author{Rafael Rezende\\
INRIA Paris\\
Institution1 address\\
{\tt\small rafael.sampaio_de_rezende@inria.fr}
% For a paper whose authors are all at the same institution,
% omit the following lines up until the closing ``}''.
% Additional authors and addresses can be added with ``\and'',
% just like the second author.
% To save space, use either the email address or home page, not both
\and
Second Author\\
Institution2\\
First line of institution2 address\\
{\tt\small secondauthor@i2.org}
}
\maketitle

%\date{}

We thank the reviewers for their constructive comments and we address their concerns below.

\section{R1} The comments are mostly positive.

\noindent{\em - ESVM and linear SLEM.} We wish to note that the original ESVM method for retrieval only outperforms the linear version of the proposed SLEM algorithm for SPOC features and the Holidays datasets. For all other features and datasets the linear SLEM does as well or better, and is of course much faster than the ESVM.

\noindent{\em - Computational cost of kernel SLEMs.} The high computational cost of kernelized SLEMs is decreased by the low-rank decompositions we presented. Table 2 shown around 0.5$\%$ improvement (significant for this performance zone) between the linear SLEM and 32-rank decomposition for Poly SLEM with NetVLAD-512 and NetVLAD-4096 on Holidays.

\section{R3} We answer below the questions of R3.

\noindent{\em - Question 1: main novelty.} We believe that the main novelty of the paper is two-fold: (1) using the square loss, which avoids retraining for each additional positive training example ; (2) kernelizing the method while keeping a reasonable memory footprint through the use of low-rank approximations. Similar ideas have of course been used in other contexts in machine learning \cite{BaJo02,FiSc01}, as already mentioned in the paper.  As pointed out by the reviewer, processing the data offline before adding one extra sample has also been used before for efficient cross-validation of residual error (we will add a reference if the paper is accepted). Our work is, however, to our knowledge, the first to apply these ideas to examplar-based classifiers, in particular in the context of image retrieval, and this results in a high-accuracy, efficient method for image retrieval. An additional (if minor) advantage of our approach compared to the original ESVM approach  to image retrieval of \cite{ZePe15}  is that it does not require the calibration of the $\theta$ parameter.

\noindent{\em - Question 2.} We do not have a well-justified explanation (yet) as to why our linear SLEM performs significantly better than the NetVLAD baseline in our experiments, although we suspect, like the reviewer, that it is a mixture of loss function and training data from the negative set. The code for the cited features is, as far as we know, not available publicly, which has prevented us from running experiments with them.

\noindent{\em - Question 3: why use kernel SLEM in practice.}  It is clearly true that linear SLEMs perform as well as kernelized ones for some features and datasets (e.g., VLAD, AlexNet and NetVLAD for Holidays, or VLAD and SPoC for Oxford5K), where it makes more sense to use linear SLEM. For some features and datasets, there is, however, a significant difference in performance, for example SPoC for Holidays (4$\%$ difference between linear and full-rank Gaussian or Poly SLEM), or SPoC ans NetVLAD features for Oxford105K (respectively almost 6$\%$ and over 2$\%$ between linear and Poly SLEMs). As shown in Table 1 of the supplementary material, increasing the rank of the approximation to 1024 improves accuracy within 1$\%$ of less of the full-rank method for NetVLAD and Poly SLEM for both Holidays and Oxford5K while keeping a reasonable memory footprint. To confirm this was true more generally, we have run additional experiments with Oxford105K, achieving performance within 1.5$\%$ of the full-rank method with a rank of 1024.

\section{R4} We address below the main concerns of R4.

\noindent{\em - Clarify claims.} We will indeed improve the presentation of the paper
and its claims, following our response to Question 1 of R3 above and reorganizing the Introduction.

\noindent{\em - Remove KPCA vs ICD discussion.} We agree these results are well know, in particular in the machine learnng community. We will remove the figure and
only keep a brief discussion of the issue since we believe it may be useful to
computer vision researchers.

\noindent{\em - Typos.} We will correct all typos in Lemma 1 and Eq. (16) and move Eq. (15) to section 3.3.

\noindent{\em - "LDA, ESVM and Linear SLEM".} We will correct the sentence, which should have said "PCAW, LDA and ESVM".

\bibliographystyle{plain}
\bibliography{sup,jz}
\end{document}
